### 2020-06-03 docs/index.Rmd
Create index.Rmd

---
title: "Keyrun project"
output: html_document
---

 <!-- ADD LINKS BELOW, note, using the RMD link syntax will open the presentation in the same tab, which is annoying. Use the html with the "_blank" to fix that behavior -->

These pages document decisions made regarding datasets and modeling for multispecies model keyruns for Georges Bank, USA. Methods documentation and eventually draft manuscripts will be found here too.

### 2020-06-03 docs/DatasetDimensions.Rmd
initialize decision documentation

---
title: "MS Keyrun Dataset Dimensions"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## To be filled in meeting week of June 15 2020

### Years (start-end)

### Seasons (fall/spring surveys? fishery seasons?)

### Spatial footprint (Georges Bank defined how)

### 2020-06-03 docs/index.Rmd
initialize decision documentation

@@
 ---
-title: "Keyrun project"
+title: "NEFSC multispecies model keyrun project"
 output: html_document
@@
 
-These pages document decisions made regarding datasets and modeling for multispecies model keyruns for Georges Bank, USA. Methods documentation and eventually draft manuscripts will be found here too.
+These pages document decisions made regarding datasets and modeling for multispecies model keyruns for Georges Bank, USA.
+Methods documentation and eventually draft manuscripts will be found here too.
+The target date for model keyrun reviews is October 2021, with all modeling results and documentation submitted by 1 September 2021.
+
+Model "keyruns" are defined as "standardized model runs updated with recent data, producing agreed output." The ICES group WGSAM regularly reviews multispecies model keyruns using criteria found [here](https://ices-eg.github.io/wg_WGSAM/ReviewCriteria.html).
+
+## Dataset dimensions
+
+<a href="https://https://noaa-edab.github.io/ms-keyrun/DatasetDimensions.html" target="_blank"> Decisions with meeting minutes - 15 June 2020 </a>
+
+## Dataset source and methods documentation
+
+## Dataset descriptions and links
+
+## Model documentation
+
+## Model results
+
+## Overall keyrun results

### 2020-06-04 docs/DataNeeds.Rmd
add data needs discussion doc

---
title: "MS Keyrun Data Needs"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Decisions from meeting June 3 2020

Notes were taken on the [KEYRUN Kanban board](https://www.st.nmfs.noaa.gov/jira/secure/RapidBoard.jspa?rapidView=384&projectKey=KEYRUN) under the epic [KEYRUN-4](https://www.st.nmfs.noaa.gov/jira/browse/KEYRUN-4) (a password is required to access these sites, but current information is copied below).

## Objective: Develop a common input dataset for use in all keyrun models

The purpose of the June 3 meeting was to describe the decisions and components required to develop the common input dataset for the project.

### Epic description

(This text copied from KEYRUN-4 Description and lightly edited)

Below is a list of potential stories under this Epic with links to other Epics. Numbers indicate order tasks need to be done (priority) established at the June 3 2020 meeting.

TODO: Add each story. Add done condition for each story.

1. Document decisions on all stories below (maybe rmd) and document as we go for all datasets, Jira not searchable so establish decision document and documentation elsewhere (pull comment from done column? plugins may be available by license)

1. Establish temporal and spatial dimensions for data (years, seasons, areas).

  define Georges Bank for survey and fisheries which will not match, could use habitat model

2. Establish species lists (Rpath KEYRUN-2 larger list than multispecies models KEYRUN-1 and KEYRUN-3)

2. Define environmental data needed for each model

2. Establish dataset standards (See e.g. condition work)

3. Biomass index data from surveys for all models:

4. Establish common definitions of fleets: applies to all catch data, discard data, and composition data.

5. Catch (landings) data for all models:

   Rpath in KEYRUN-2

   Hydra and MSCAA in KEYRUN-1

   All other tested models in KEYRUN-3

5. Discard data for all models, needs methods discussion/decision (check: Andy)

6. Length composition data from surveys and fisheries for Hydra in KEYRUN-1 (check: Andy)

6. Age composition data from surveys and fisheries for MSCAA in KEYRUN-1

6. Define which diet and or consumption data from surveys and fisheries for all models

6. Define which life history data from surveys and/or fisheries for all models (e.g. maturity, weight at age, maximum age for mortality estimation, etc)

### Additional meeting notes

Progress on catch data:

Notes on mscatch and comlands. comlands in a package now. dataset from comland is input to mscatch which combines with observer data length samples, for each quarter/yr for species, gear, can assign length samples to each of those landings bins. done and tested. Discards not yet accounted for, difficult and needs decisions. mostly documented, decision process for filling blanks documented in log file. Canadian landings for GB may be needed.

there is a comdisc as well, computes discard to kept ratio. add to subtasks under discard story

Identify done criteria, meet them, then run by key stock assessment folks.

### Progress

Stories have been added to the board as of `r format(Sys.time(), '%d %B, %Y')`:

```{r, echo=FALSE}
htmltools::includeHTML("NMFS JIRA 2020-06-04T10_45_07-0400.html")
```

### 2020-06-04 docs/index.Rmd
correct link in index
add data needs discussion doc

@@
 
+# Data needs
+
+<a href="https://noaa-edab.github.io/ms-keyrun/DataNeeds.html" target="_blank"> Decisions with meeting notes - 3 June 2020 </a>
+
 ## Dataset dimensions
 
-<a href="https://https://noaa-edab.github.io/ms-keyrun/DatasetDimensions.html" target="_blank"> Decisions with meeting minutes - 15 June 2020 </a>
+<a href="https://noaa-edab.github.io/ms-keyrun/DatasetDimensions.html" target="_blank"> Decisions with meeting minutes - 15 June 2020 </a>
 
@@
 
+# Models
+
 ## Model documentation
@@
 
+# Results
+
+## Individual model keyrun results
+
 ## Overall keyrun results

### 2020-06-29 docs/DatasetDimensions.Rmd
pull in results of google forms survey

@@
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: html_document
+output:
+  html_document:
+    code_fold: hide
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, 
+                      message = FALSE,
+                      warning = FALSE)
+
+library(googledrive)
+library(pacman)
+p_load(magrittr, gsubfn, stringr, tidyverse)
+
 ```
 
-## To be filled in meeting week of June 15 2020
+## Proposed Dimensions
+
+### Years: 1967-2019
+
+All surveyed agreed to end with the complete year of data 2019 to be as up to date as possible. Start year is the earliest survey year where both spring and fall are available.
+
+### Seasons:
+
+Pull annual data for fisheries, pull annual survey maintaining seasons as "spring", "fall" etc.
+for each year.
+Many models use annual catch so further resolution is unnecessary; modelers can choose to fit seasonal survey indices separately or combine to a single annual value from seasonal (documenting methods).
+
+### Spatial:
+
+Define using fishery statistical areas and match survey data as much as possible to these. Easier to map survey strata to larger fishery areas than the other way around.
+
+## Full survey results
+
+### Survey conducted week of June 15 2020
+
+```{r getsurvey}
+
+resultfile <- drive_find(pattern = "MS-Keyrun", type = "spreadsheet")
+
+responses <- drive_download(resultfile, type = "csv", overwrite = TRUE) %>%
+  {read.csv(.$local_path)}
+
+names(responses)[3] <- "StartYear"
+names(responses)[5] <- "EndYear"
+names(responses)[7] <- "SurvSeason"
+names(responses)[9] <- "FishSeason"
+names(responses)[11] <- "SpatialGB"
+
+```
+
+We had `r nrow(responses)` responses.
 
@@
 
+Responses:
+```{r}
+startend <- data.frame(Start = responses$StartYear,
+                       End = responses$EndYear)
+knitr::kable(startend)
+```
+
+**Start year rationale:**
+`r knitr::kable(responses$Rationale.for.start.time, col.names = NULL)`
+
+**End year rationale:**
+`r knitr::kable(responses$Rationale.for.end.time, col.names = NULL)`
+
 ### Seasons (fall/spring surveys? fishery seasons?)
 
+Responses:
+```{r seasons}
+
+seasons <- responses %>%
+  select(SurvSeason, FishSeason) %>%
+  gather(key = SurvFish, value = "Season") %>%
+  arrange(Season)
+
+ggplot(seasons, aes(Season, fill=SurvFish)) + 
+  geom_bar() +
+  scale_x_discrete(labels = function(x) str_wrap(x, width = 15))#+ 
+  #theme(axis.text.x = element_text(angle = 90))
+
+```
+
+**Survey season rationale:**
+`r knitr::kable(responses$Rationale.for.survey.temporal.resolution, col.names = NULL)`
+
+**Fishery season rationale:**
+`r knitr::kable(responses$Rationale.for.fishery.temporal.resolution, col.names = NULL)`
+
 ### Spatial footprint (Georges Bank defined how)
+
+Responses:
+```{r space}
+
+ggplot(responses, aes(SpatialGB)) + 
+  geom_bar() +
+  scale_x_discrete(labels = function(x) str_wrap(x, width = 35))#+
+
+```
+
+**Spatial rationale:**
+`r knitr::kable(responses$Rationale.for.spatial.definition, col.names = NULL)`

### 2020-06-29 docs/DatasetDimensions.Rmd
fix tables update index

@@
 
-### Seasons:
+### Seasons: Annual fisheries, Seasonal surveys
 
@@
 
-### Spatial:
+### Spatial: Fishery stat areas define Georges Bank
 
-Define using fishery statistical areas and match survey data as much as possible to these. Easier to map survey strata to larger fishery areas than the other way around.
+Define using fishery statistical areas (521-522, 525-526, 551-552, 561-562) and match survey data as much as possible to these.
+Easier to map survey strata to larger fishery areas than the other way around.
 
@@
                        End = responses$EndYear)
-knitr::kable(startend)
+knitr::kable(startend, booktabs=TRUE)
 ```
 
-**Start year rationale:**
-`r knitr::kable(responses$Rationale.for.start.time, col.names = NULL)`
+```{r} 
+knitr::kable(responses$Rationale.for.start.time, col.names = "Start year rationale", booktabs=TRUE)
+```
 
-**End year rationale:**
-`r knitr::kable(responses$Rationale.for.end.time, col.names = NULL)`
+```{r}
+knitr::kable(responses$Rationale.for.end.time, col.names = "End year rationale", booktabs=TRUE)
+```
 
@@
 
-**Survey season rationale:**
-`r knitr::kable(responses$Rationale.for.survey.temporal.resolution, col.names = NULL)`
+```{r} 
+knitr::kable(responses$Rationale.for.survey.temporal.resolution, col.names = "Survey season rationale", booktabs=TRUE)
+```
 
-**Fishery season rationale:**
-`r knitr::kable(responses$Rationale.for.fishery.temporal.resolution, col.names = NULL)`
+```{r} 
+knitr::kable(responses$Rationale.for.fishery.temporal.resolution, col.names = "Fishery season rationale", booktabs=TRUE)
+```
 
@@
 
-**Spatial rationale:**
-`r knitr::kable(responses$Rationale.for.spatial.definition, col.names = NULL)`
+```{r} 
+knitr::kable(responses$Rationale.for.spatial.definition, col.names = "Spatial Rationale", booktabs=TRUE)
+```

### 2020-07-08 docs/DatasetDimensions.Rmd
add link to survey form
add one more survey response
update dimensions from meeting discussion

@@
 
-### Years: 1967-2019
+### Years: 1968-2019
 
@@
 
-### Spatial: Fishery stat areas define Georges Bank
+### Spatial: 2 Approaches for Georges Bank
 
-Define using fishery statistical areas (521-522, 525-526, 551-552, 561-562) and match survey data as much as possible to these.
-Easier to map survey strata to larger fishery areas than the other way around.
+Ideal: Define using Georges Bank EPU, scale fisheries catch for each species from stat areas to GB EPU footprint using ratios in spatially explicit VTR subset of catch data.
+Work towards this ideal, comparing coverage of VTR data by species and different methods for scaling as part of data documentation.
+Consult with ADIOS team and assessment experts to see if results look reasonable.
+
+The ideal is time-consuming and cannot be started until mid-late September 2020 at the earliest.
+
+Interim: Define using fishery statistical areas (521-522, 525-526, 551-552, 561-562) and match survey data as much as possible to these.
+It is easier to map survey strata to larger fishery areas than the other way around; Sean can share code which may be modified by Gavin's student to achieve this.
+
+The interim represents the upper bound of what catch would be by species for comparison to the ideal approach and can be implemented fairly quickly.
+
+### Other notes
+
+By either spatial definition, inshore survey strata should not be relevant to Georges Bank, so the beginning and end years remain as above, and there should not be a need to combine inshore and offshore surveys using Conn method or VAST for Bigelow years.
+
+The group decided that the base dataset would pull survey data separately for the Albatross and Bigelow years without applying any conversion factors.
+These separate series can be used as indices in the structured models (MSCAA and Hydra).
+
+Methods for making a single survey time series from the two vessel survey time series (and for making annual indices from the seasonal surveys) would need to be discussed for the other models, and methods clearly documented.
+For example, the Georges Bank Rpath was calibrated to Bigelow data from 2013-2015; therefore if a single survey time series is needed, it makes sense to convert Albatross survey catches to Bigelow units using established conversion factors.
+We are putting off this discussion for now.
 
@@
 
-### Survey conducted week of June 15 2020
+### Survey conducted week of June 15 2020, extended to 3 July
+
+[This is the survey form.](https://forms.gle/Pq78yqwa5ueFi8Kq8)
 
 ```{r getsurvey}
+
+#survey link https://forms.gle/Pq78yqwa5ueFi8Kq8
+#results https://docs.google.com/forms/d/1y78AdOeRp0ohGSDU5QP2d2UBCz--eP_5b5Tu79QO-N8/edit#responses
 
@@
 ```
+
+## Meeting minutes
+
+[Click here!](https://docs.google.com/document/d/19FpJ8w2a-ZZ8ZY4ZmsnjlIilLeN1X9_Tz1dM-KsdOOw/edit#heading=h.kvnvc0pxyvqq)

### 2020-07-08 docs/SpeciesEnvDat.Rmd
start on species list with full rpath table

---
title: "MS Keyrun Species, Environment, Data Formats"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE,
                      warning = FALSE)

library(googledrive)
library(pacman)
p_load(repmis, RCurl, DT, magrittr, gsubfn, stringr, tidyverse)

```

## Species lists

### Multispecies models (focal species)

Multispecies models focus on a subset of interacting species. They will estimate population parameters based on fits to biomass, catch, and (if appropriate) size or age information, establish reference points, and evaluate status for the interacting species.

### Food web model (focal + rest of system)

The food web model is intended to examine wider ecosystem responses to management measures applied to a subset of interacting species. It will estimate predator-prey interaction and other parameters based on fits to biomass, catch, and diet information. This model should track general trends (if not interannual variability) for focal species and produce reasonable reactions by their predators and prey.

Current species groups in the GB Rpath model:
```{r rpathspp, message=FALSE}

# Sean's file from https://github.com/NOAA-EDAB/GBRpath/blob/master/data/SOE_species_list.RData
# loads the RData as object "species"

source_data("https://github.com/NOAA-EDAB/GBRpath/blob/master/data/SOE_species_list.RData?raw=True")

# used to get a list to paste into google form
#write.table(file = "spplist.txt", sort(unique(species$RPATH)))

rpathspp <- species %>%
  select(RPATH, COMNAME, SCINAME, SVSPP) %>%
  filter(RPATH != "") %>%
  distinct()

aggs <- rpathspp %>%
  count(RPATH, sort = TRUE, name = "Nspp") %>%
  filter(Nspp > 1)

singlespp <- anti_join(rpathspp, aggs) %>%
  mutate(Nspp = 1)

rpathlist <- full_join(singlespp, aggs)

datatable(rpathlist, rownames = FALSE, 
          options = list(pageLength = 25, 
                         order = list(list(0, 'asc'))
          ))

```

## Envrionmental data

### Multispecies models

### Food web model

## Programming frameworks and data formats

## Full survey results

### Survey conducted July 13-24 2020

[This is the survey form.](https://forms.gle/gfA3tKFKRBxf3Mu38)

```{r getsurvey, eval=FALSE}

#survey link https://forms.gle/gfA3tKFKRBxf3Mu38
#results

resultfile <- drive_find(pattern = "MS-Keyrun Species", type = "spreadsheet")

responses <- drive_download(resultfile, type = "csv", overwrite = TRUE) %>%
  {read.csv(.$local_path)}

names(responses)[3] <- "StartYear"
names(responses)[5] <- "EndYear"
names(responses)[7] <- "SurvSeason"
names(responses)[9] <- "FishSeason"
names(responses)[11] <- "SpatialGB"

```

## Meeting minutes (old link)

[Click here!](https://docs.google.com/document/d/19FpJ8w2a-ZZ8ZY4ZmsnjlIilLeN1X9_Tz1dM-KsdOOw/edit#heading=h.kvnvc0pxyvqq)

### 2020-07-08 docs/SpeciesEnvDat2.Rmd
start on species list with full rpath table

--- 2020-07-08 docs/SpeciesEnvDat.Rmd
@@
 Current species groups in the GB Rpath model:
-```{r rpathspp, message=FALSE}
+```{r rpathspp}
 
@@
   select(RPATH, COMNAME, SCINAME, SVSPP) %>%
-  filter(RPATH != "") %>%
   distinct()
 
-aggs <- rpathspp %>%
-  count(RPATH, sort = TRUE, name = "Nspp") %>%
-  filter(Nspp > 1)
+  #count(SVSPP, RPATH, sort = TRUE, name = "Nspp")
 
-singlespp <- anti_join(rpathspp, aggs) %>%
-  mutate(Nspp = 1)
-
-rpathlist <- full_join(singlespp, aggs)
-
-datatable(rpathlist, rownames = FALSE, 
-          options = list(pageLength = 25, 
-                         order = list(list(0, 'asc'))
-          ))
+datatable(rpathspp, rownames = FALSE, options = list(pageLength = 25))
 
@@
 [This is the survey form.](https://forms.gle/gfA3tKFKRBxf3Mu38)
+
+### REVISE BELOW HERE FOR UPDATED SURVEY
 
@@
 
-## Meeting minutes (old link)
+We had `r nrow(responses)` responses.
+
+### Years (start-end)
+
+Responses:
+```{r}
+startend <- data.frame(Start = responses$StartYear,
+                       End = responses$EndYear)
+knitr::kable(startend, booktabs=TRUE)
+```
+
+```{r} 
+knitr::kable(responses$Rationale.for.start.time, col.names = "Start year rationale", booktabs=TRUE)
+```
+
+```{r}
+knitr::kable(responses$Rationale.for.end.time, col.names = "End year rationale", booktabs=TRUE)
+```
+
+### Seasons (fall/spring surveys? fishery seasons?)
+
+Responses:
+```{r seasons}
+
+seasons <- responses %>%
+  select(SurvSeason, FishSeason) %>%
+  gather(key = SurvFish, value = "Season") %>%
+  arrange(Season)
+
+ggplot(seasons, aes(Season, fill=SurvFish)) + 
+  geom_bar() +
+  scale_x_discrete(labels = function(x) str_wrap(x, width = 15))#+ 
+  #theme(axis.text.x = element_text(angle = 90))
+
+```
+
+```{r} 
+knitr::kable(responses$Rationale.for.survey.temporal.resolution, col.names = "Survey season rationale", booktabs=TRUE)
+```
+
+```{r} 
+knitr::kable(responses$Rationale.for.fishery.temporal.resolution, col.names = "Fishery season rationale", booktabs=TRUE)
+```
+
+### Spatial footprint (Georges Bank defined how)
+
+Responses:
+```{r space}
+
+ggplot(responses, aes(SpatialGB)) + 
+  geom_bar() +
+  scale_x_discrete(labels = function(x) str_wrap(x, width = 35))#+
+
+```
+
+```{r} 
+knitr::kable(responses$Rationale.for.spatial.definition, col.names = "Spatial Rationale", booktabs=TRUE)
+```
+
+## Meeting minutes
 

### 2020-07-08 docs/index.Rmd
fix tables update index
add one more survey response
update dimensions from meeting discussion
add species list to index

@@
 
-<a href="https://noaa-edab.github.io/ms-keyrun/DatasetDimensions.html" target="_blank"> Decisions with meeting minutes - 15 June 2020 </a>
+<a href="https://noaa-edab.github.io/ms-keyrun/DatasetDimensions.html" target="_blank"> Year, season, area decisions - 8 July 2020 </a>
+
+<a href="https://noaa-edab.github.io/ms-keyrun/SpeciesEnvDat.html" target="_blank"> Species table from Rpath - 8 July 2020 </a>
 

### 2020-07-09 docs/SpeciesEnvDat.Rmd
add blank spreadsheet

@@
 
-### Multispecies models
+The `ecodata` R package has up to date information at the Georges Bank EPU scale as listed here: https://noaa-edab.github.io/ecodata/landing_page.
 
-### Food web model
+Current time series in ecodata:
+```{r ecodata}
+
+#remotes::install_github("noaa-edab/ecodata",build_vignettes=TRUE)
+library(ecodata)
+
+ecodsets <- data(package = "ecodata")$result[, c("Item", "Title")]
+
+datatable(ecodsets, rownames = FALSE, 
+          options = list(pageLength = 25)
+)
+
+```
 
 ## Programming frameworks and data formats
+
+How many models are implemented in R?
+See [discussion](https://docs.google.com/document/d/19FpJ8w2a-ZZ8ZY4ZmsnjlIilLeN1X9_Tz1dM-KsdOOw/edit#heading=h.kvnvc0pxyvqq) of pros and cons of using R data package for standardizing input datasets and making them available.
 
@@
 
-```{r getsurvey, eval=FALSE}
+```{r getsurvey}
 
 #survey link https://forms.gle/gfA3tKFKRBxf3Mu38
-#results
+#results link https://docs.google.com/spreadsheets/d/1sckY46jdyBa1Fhydw2syBawkjGPcHX4YwQmDCDBXcGU/edit#gid=284425233
 
@@
 
-names(responses)[3] <- "StartYear"
-names(responses)[5] <- "EndYear"
-names(responses)[7] <- "SurvSeason"
-names(responses)[9] <- "FishSeason"
-names(responses)[11] <- "SpatialGB"
+names(responses)[3] <- "Species"
+names(responses)[5] <- "Environment"
+names(responses)[7] <- "ModelInR"
+names(responses)[8] <- "ModelUseR"
+names(responses)[9] <- "RData"
+
+```
+
+We had `r nrow(responses)` responses.
+
+### Species
+
+Responses:
+```{r}
+# each row is a response with a comma delimited list of species
+# string split each row to species
+# tally by species
+# histogram
+
+#msspecies <- responses$Species
+
+```
+
+```{r} 
+knitr::kable(responses$Rationale.for.species.included, col.names = "Species rationale", booktabs=TRUE)
+```
+
+### Environment
+
+Responses:
+```{r}
+# each row is a response with a comma delimited list of environmental data
+# string split each row to dataset name
+# tally by dataset
+# histogram
+
+#msspecies <- responses$Environment
+
+```
+
+```{r} 
+knitr::kable(responses$Rationale.for.environmental.time.series.included, col.names = "Environment rationale", booktabs=TRUE)
+```
+
+### Model programming frameworks
+
+Responses:
+```{r}
+# these are multiple choice yes/no so just plot
+
+#useR <- data.frame(ModelinR = responses$ModelinR,
 

### 2020-07-30 docs/SpeciesEnvDat.Rmd
update with survey results

@@
 Responses:
-```{r}
+```{r, fig.height=6}
 # each row is a response with a comma delimited list of species
@@
 
-#msspecies <- responses$Species
+msspecies <- data.frame(Species = responses$Species,
+                        Model = c("MS", "MS", "NA", "MS", "Rpath", "MS")) %>%
+  filter(Model != "NA") %>%
+  separate_rows(Species, sep = ",") %>%
+  mutate(Species = str_trim(Species, side = "both"))
+
+ggplot(msspecies, aes(Species)) + 
+  geom_bar() +
+  facet_wrap(~Model, nrow = 2, scales="free") +
+  theme(axis.text.x = element_text(angle = 90))
 
@@
 Responses:
-```{r}
+```{r, fig.height=10}
 # each row is a response with a comma delimited list of environmental data
@@
 
-#msspecies <- responses$Environment
+msenv <- data.frame(Environment = responses$Environment,
+                        Model = c("MS", "MS", "NA", "MS", "Rpath", "MS")) %>%
+  filter(Model != "NA") %>%
+  separate_rows(Environment, sep = ",") %>%
+  mutate(Environment = str_trim(Environment, side = "both"))
+
+ggplot(msenv, aes(Environment)) + 
+  geom_bar() +
+  facet_wrap(~Model, nrow = 2, scales="free") +
+  theme(axis.text.x = element_text(angle = 90))
 
@@
 
+Many but not all models interface with R and can use R datasets. Even for those that do not, all agreed on using an R data package for standardizing input datasets and making them available.
+
 Responses:
@@
 
-#useR <- data.frame(ModelinR = responses$ModelinR,
+useR <- data.frame(ModelInR = responses$ModelInR,
+                   ModelUseR = responses$ModelUseR,
+                   RData = responses$RData) %>%
+  gather(question, answer, ModelInR:RData) %>%
+  mutate(shortans = ifelse(is.na(word(answer, 1, 2)), 
+                           answer, 
+                           word(answer, 1, 2)))
+
+ggplot(useR, aes(question)) +
+  geom_bar(aes(fill=shortans)) +
+  labs(x = "Question", 
+       fill = "Short Answer") +
+  scale_x_discrete(labels = function(x) str_wrap(x, width = 35))
 
 ```
+Short answer key:
+"No but" = `r unique(useR$answer)[2]`  
+"Probably, we" = `r unique(useR$answer)[4]`
 
-## Meeting minutes (old link)
+## Meeting minutes, 5 August 2020
 
-[Click here!](https://docs.google.com/document/d/19FpJ8w2a-ZZ8ZY4ZmsnjlIilLeN1X9_Tz1dM-KsdOOw/edit#heading=h.kvnvc0pxyvqq)
+[Click here!](https://docs.google.com/document/d/19FpJ8w2a-ZZ8ZY4ZmsnjlIilLeN1X9_Tz1dM-KsdOOw/edit#heading=h.m573izndd2nh)

### 2020-07-31 docs/SpeciesEnvDat.Rmd
update with proposed decsions and date

@@
 
-### Multispecies models (focal species)
+(We first pull species lists from the Rpath model and load the survey results.)
 
-Multispecies models focus on a subset of interacting species.
-They will estimate population parameters based on fits to biomass, catch, and (if appropriate) size or age information, establish reference points, and evaluate status for the interacting species.
-
-### Food web model (focal + rest of system)
-
-The food web model is intended to examine wider ecosystem responses to management measures applied to a subset of interacting species.
-It will estimate predator-prey interaction and other parameters based on fits to biomass, catch, and diet information.
-This model should track general trends (if not interannual variability) for focal species and produce reasonable reactions by their predators and prey.
-
-Current species groups in the GB Rpath model:
-```{r rpathspp, message=FALSE}
+```{r getrpathspp, message=FALSE}
 
@@
 
-source_data("https://github.com/NOAA-EDAB/GBRpath/blob/master/data/SOE_species_list.RData?raw=True")
+invisible(source_data("https://github.com/NOAA-EDAB/GBRpath/blob/master/data/SOE_species_list.RData?raw=True"))
 
@@
 
+```
+
+```{r getsurvey, message=FALSE}
+
+#survey link https://forms.gle/gfA3tKFKRBxf3Mu38
+#results link https://docs.google.com/spreadsheets/d/1sckY46jdyBa1Fhydw2syBawkjGPcHX4YwQmDCDBXcGU/edit#gid=284425233
+
+resultfile <- drive_find(pattern = "MS-Keyrun Species", type = "spreadsheet")
+
+responses <- drive_download(resultfile, type = "csv", overwrite = TRUE) %>%
+  {read.csv(.$local_path)}
+
+names(responses)[3] <- "Species"
+names(responses)[5] <- "Environment"
+names(responses)[7] <- "ModelInR"
+names(responses)[8] <- "ModelUseR"
+names(responses)[9] <- "RData"
+
+```
+
+### Multispecies models (focal species)
+
+Multispecies models focus on a subset of interacting species.
+They will estimate population parameters based on fits to biomass, catch, and (if appropriate) size or age information, establish reference points, and evaluate status for the interacting species.
+
+These species received >1 vote for inclusion in MS-Keyrun multispecies models:
+
+```{r msmodspp}
+# each row is a response with a comma delimited list of species
+# string split each row to species
+# tally by species
+# histogram
+
+msspecies <- data.frame(Species = responses$Species,
+                        Model = c("MS", "MS", "NA", "MS", "Rpath", "MS")) %>%
+  filter(Model != "NA") %>%
+  separate_rows(Species, sep = ",") %>%
+  mutate(Species = str_trim(Species, side = "both"))
+
+focal <- msspecies %>%
+  filter(Model == "MS") %>%
+  count(Species) %>%
+  filter(n > 1) %>%
+  left_join(rpathspp, by = c("Species" = "RPATH")) %>%
+  select(-COMNAME) %>%
+  arrange(desc(n))
+
+knitr::kable(focal)
+
+```
+
+Species modeled by all compared multispecies models could include all of these, or could be limited to a subset.
+
+### Food web model (focal + rest of system)
+
+The food web model is intended to examine wider ecosystem responses to management measures applied to a subset of interacting species.
+It will estimate predator-prey interaction and other parameters based on fits to biomass, catch, and diet information.
+This model should track general trends (if not interannual variability) for focal species and produce reasonable reactions by their predators and prey.
+
+Current species groups in the GB Rpath model (all retained for MS-Keyrun):
+```{r rpathspp, message=FALSE}
+
 datatable(rpathlist, rownames = FALSE, 
@@
 ## Envrionmental data
+
+Hydra, the length-based multispecies model, already incorporates a time series of Georges Bank bottom temperature as a covariate in simulation mode to force changes in growth, maturity, and recruitment.
+For MS-Keyrun, we can explore fitting the parameters governing the influence of this covariate.
+
+Rpath can be directly forced by a primary production time series, and temperature-based mediation functions can be included (but it is unclear whether parameters can be fit for these functions).
+
+Two mulispecies models are not currently set up to incorporate environmental data:
+MSCAA, MSSPM/Kraken
+
+The WHAM model can potentially incorporate any environmental time series that can influence recruitment or mortality.
+
+These environmental time series received >1 vote for inclusion in MS-Keyrun multispecies and Rpath models:
+```{r envdat}
+# each row is a response with a comma delimited list of environmental data
+# string split each row to dataset name
+# tally by dataset
+# histogram
+
+msenv <- data.frame(Environment = responses$Environment,
+                        Model = c("MS", "MS", "NA", "MS", "Rpath", "MS")) %>%
+  filter(Model != "NA") %>%
+  separate_rows(Environment, sep = ",") %>%
+  mutate(Environment = str_trim(Environment, side = "both"))
+
+includedenv <- msenv %>%
+  count(Environment) %>%
+  filter(n > 1) %>%
+  arrange(desc(n))
+
+knitr::kable(includedenv)
+
+```
+
+We can include GB temperature and primary production time series as part of the MS-Keyrun dataset. Modelers may retrieve additional data from `ecodata`.
 
@@
 
-How many models are implemented in R?
+Most models are either implemented in R or can work directly with R datasets:
+```{r useRdata}
+# these are multiple choice yes/no so just plot
+
+useR <- data.frame(ModelInR = responses$ModelInR,
+                   ModelUseR = responses$ModelUseR,
+                   RData = responses$RData) %>%
+  gather(question, answer, ModelInR:RData) %>%
+  mutate(shortans = ifelse(is.na(word(answer, 1, 2)), 
+                           answer, 
+                           word(answer, 1, 2)))
+yesR <- useR %>%
+  group_by(question) %>%
+  count(answer=="Yes") %>%
+  mutate(tot = sum(n)) %>%
+  mutate(percent = n/tot*100)
+
+knitr::kable(yesR)
+
+```
+
+All surveyed agreed that we should maintain datasets within an R data package for the MS-Keyrun project.
+
 See [discussion](https://docs.google.com/document/d/19FpJ8w2a-ZZ8ZY4ZmsnjlIilLeN1X9_Tz1dM-KsdOOw/edit#heading=h.kvnvc0pxyvqq) of pros and cons of using R data package for standardizing input datasets and making them available.
@@
 
-### Survey conducted July 13-24 2020
+### Survey conducted July 13-29 2020
 
 [This is the survey form.](https://forms.gle/gfA3tKFKRBxf3Mu38)
-
-```{r getsurvey}
-
-#survey link https://forms.gle/gfA3tKFKRBxf3Mu38
-#results link https://docs.google.com/spreadsheets/d/1sckY46jdyBa1Fhydw2syBawkjGPcHX4YwQmDCDBXcGU/edit#gid=284425233
-
-resultfile <- drive_find(pattern = "MS-Keyrun Species", type = "spreadsheet")
-
-responses <- drive_download(resultfile, type = "csv", overwrite = TRUE) %>%
-  {read.csv(.$local_path)}
-
-names(responses)[3] <- "Species"
-names(responses)[5] <- "Environment"
-names(responses)[7] <- "ModelInR"
-names(responses)[8] <- "ModelUseR"
-names(responses)[9] <- "RData"
-
-```
 
@@
 Responses:
-```{r, fig.height=6}
-# each row is a response with a comma delimited list of species
-# string split each row to species
-# tally by species
-# histogram
-
-msspecies <- data.frame(Species = responses$Species,
-                        Model = c("MS", "MS", "NA", "MS", "Rpath", "MS")) %>%
-  filter(Model != "NA") %>%
-  separate_rows(Species, sep = ",") %>%
-  mutate(Species = str_trim(Species, side = "both"))
+```{r plotspp, fig.height=6}
 
@@
 Responses:
-```{r, fig.height=10}
-# each row is a response with a comma delimited list of environmental data
-# string split each row to dataset name
-# tally by dataset
-# histogram
-
-msenv <- data.frame(Environment = responses$Environment,
-                        Model = c("MS", "MS", "NA", "MS", "Rpath", "MS")) %>%
-  filter(Model != "NA") %>%
-  separate_rows(Environment, sep = ",") %>%
-  mutate(Environment = str_trim(Environment, side = "both"))
+```{r plotenv, fig.height=10}
 
@@
 Responses:
-```{r}
-# these are multiple choice yes/no so just plot
-
-useR <- data.frame(ModelInR = responses$ModelInR,
-                   ModelUseR = responses$ModelUseR,
-                   RData = responses$RData) %>%
-  gather(question, answer, ModelInR:RData) %>%
-  mutate(shortans = ifelse(is.na(word(answer, 1, 2)), 
-                           answer, 
-                           word(answer, 1, 2)))
+```{r plotdat}
 
@@
 ```
+
 Short answer key:

### 2020-09-03 docs/CompsDat.Rmd
update from August meeting

--- 2020-07-30 docs/SpeciesEnvDat.Rmd
@@
 ---
-title: "MS Keyrun Species, Environment, Data Formats"
+title: "MS Keyrun Age, Length, and Diet Composition Data Needs"
 author: "Sarah Gaichas"
@@
 
-## Species lists
+(We first load the survey results.)
 
-### Multispecies models (focal species)
+```{r getsurvey, message=FALSE}
 
-Multispecies models focus on a subset of interacting species.
-They will estimate population parameters based on fits to biomass, catch, and (if appropriate) size or age information, establish reference points, and evaluate status for the interacting species.
+#survey link https://forms.gle/5PkpTiCvXEgda2it6
+#results link https://docs.google.com/spreadsheets/d/1BZFqpIfmPdUW0XG3nyZbRrCBW-N-jV_r26mG7oKVdlQ/edit#gid=1759185696
 
-### Food web model (focal + rest of system)
-
-The food web model is intended to examine wider ecosystem responses to management measures applied to a subset of interacting species.
-It will estimate predator-prey interaction and other parameters based on fits to biomass, catch, and diet information.
-This model should track general trends (if not interannual variability) for focal species and produce reasonable reactions by their predators and prey.
-
-Current species groups in the GB Rpath model:
-```{r rpathspp, message=FALSE}
-
-# Sean's file from https://github.com/NOAA-EDAB/GBRpath/blob/master/data/SOE_species_list.RData
-# loads the RData as object "species"
-
-source_data("https://github.com/NOAA-EDAB/GBRpath/blob/master/data/SOE_species_list.RData?raw=True")
-
-# used to get a list to paste into google form
-#write.table(file = "spplist.txt", sort(unique(species$RPATH)))
-
-rpathspp <- species %>%
-  select(RPATH, COMNAME, SCINAME, SVSPP) %>%
-  filter(RPATH != "") %>%
-  distinct()
-
-aggs <- rpathspp %>%
-  count(RPATH, sort = TRUE, name = "Nspp") %>%
-  filter(Nspp > 1)
-
-singlespp <- anti_join(rpathspp, aggs) %>%
-  mutate(Nspp = 1)
-
-rpathlist <- full_join(singlespp, aggs)
-
-datatable(rpathlist, rownames = FALSE, 
-          options = list(pageLength = 25, 
-                         order = list(list(0, 'asc'))
-          ))
-
-```
-
-## Envrionmental data
-
-The `ecodata` R package has up to date information at the Georges Bank EPU scale as listed here: https://noaa-edab.github.io/ecodata/landing_page.
-
-Current time series in ecodata:
-```{r ecodata}
-
-#remotes::install_github("noaa-edab/ecodata",build_vignettes=TRUE)
-library(ecodata)
-
-ecodsets <- data(package = "ecodata")$result[, c("Item", "Title")]
-
-datatable(ecodsets, rownames = FALSE, 
-          options = list(pageLength = 25)
-)
-
-```
-
-## Programming frameworks and data formats
-
-How many models are implemented in R?
-See [discussion](https://docs.google.com/document/d/19FpJ8w2a-ZZ8ZY4ZmsnjlIilLeN1X9_Tz1dM-KsdOOw/edit#heading=h.kvnvc0pxyvqq) of pros and cons of using R data package for standardizing input datasets and making them available.
-
-## Full survey results
-
-### Survey conducted July 13-24 2020
-
-[This is the survey form.](https://forms.gle/gfA3tKFKRBxf3Mu38)
-
-```{r getsurvey}
-
-#survey link https://forms.gle/gfA3tKFKRBxf3Mu38
-#results link https://docs.google.com/spreadsheets/d/1sckY46jdyBa1Fhydw2syBawkjGPcHX4YwQmDCDBXcGU/edit#gid=284425233
-
-resultfile <- drive_find(pattern = "MS-Keyrun Species", type = "spreadsheet")
+resultfile <- drive_find(pattern = "MS-Keyrun Compositional", type = "spreadsheet")
 
@@
 
-names(responses)[3] <- "Species"
-names(responses)[5] <- "Environment"
-names(responses)[7] <- "ModelInR"
-names(responses)[8] <- "ModelUseR"
-names(responses)[9] <- "RData"
+names(responses)[3] <- "ModType"
+names(responses)[4] <- "ModName"
+names(responses)[5] <- "AgeLength"
+names(responses)[6] <- "AgeNeeds"
+names(responses)[7] <- "LengthNeeds"
+names(responses)[8] <- "Diet"
+names(responses)[9] <- "DietNeeds"
+names(responses)[10] <- "Comments"
 
 ```
+
+## Model types
+
+Multispecies models will focus on 10 species: Atlantic herring, Atlantic cod, goosefish, haddock, spiny dogfish, winter founder, yellowtail flounder, Atlantic mackerel, silver hake, and winter skate.
+
+Models will estimate population parameters based on fits to biomass, catch, and (if appropriate) size or age information, establish reference points, and evaluate status for the interacting species.
+Data needs by model type are summarized below.
+
+```{r modtypedat}
+
+compdat <- data.frame(ModType = responses$ModType,
+                      ModName = responses$ModName,
+                      AgeLength = responses$AgeLength) %>%
+  filter(ModType != "NA") %>%
+  separate_rows(AgeLength, sep = ",")
+
+dietdat <- data.frame(ModType = responses$ModType,
+                      ModName = responses$ModName,
+                      Diet = responses$Diet) %>%
+  filter(ModType != "NA") %>%
+  separate_rows(Diet, sep = ",")
+
+# rename AgeLength and Diet columns to CompDat and merge
+
+```
+
+## Age and length composition data needed
+
+## Diet composition data needed
+
+## Full survey results
+
+### Survey conducted August 27-September 2 2020
+
+[This is the survey form.](https://forms.gle/5PkpTiCvXEgda2it6)
 
@@
 ```{r} 
-knitr::kable(responses$Rationale.for.species.included, col.names = "Species rationale", booktabs=TRUE)
+knitr::kable(responses$AgeNeeds, col.names = "Age information needs", booktabs=TRUE)
+```
+
+```{r} 
+knitr::kable(responses$LengthNeeds, col.names = "Length information needs", booktabs=TRUE)
 ```

### 2020-09-03 docs/SpeciesEnvDat.Rmd
update from August meeting

@@
 They will estimate population parameters based on fits to biomass, catch, and (if appropriate) size or age information, establish reference points, and evaluate status for the interacting species.
+
+Decision: Use the top ranked 10 species in the table below: Atlantic herring, Atlantic cod, goosefish, haddock, spiny dogfish, winter founder, yellowtail flounder, Atlantic mackerel, silver hake, and winter skate.
+Rationale: These species occupy Georges Bank while pollock is found more in the neighboring Gulf of Maine.
 
@@
 
-We can include GB temperature and primary production time series as part of the MS-Keyrun dataset. Modelers may retrieve additional data from `ecodata`.
+Decision: We will include GB temperature and primary production time series as part of the MS-Keyrun dataset. Modelers may retrieve additional data from `ecodata`.
 
@@
 
-All surveyed agreed that we should maintain datasets within an R data package for the MS-Keyrun project.
+Decision: All surveyed agreed that we should maintain datasets within an R data package for the MS-Keyrun project.
 

### 2020-09-03 docs/index.Rmd
update index
update with proposed decsions and date
update description
update from August meeting

@@
 
-<a href="https://noaa-edab.github.io/ms-keyrun/SpeciesEnvDat.html" target="_blank"> Species table from Rpath - 8 July 2020 </a>
+<a href="https://noaa-edab.github.io/ms-keyrun/SpeciesEnvDat.html" target="_blank"> Species, environment, data package decisions - 5 August 2020 </a>
+
+<a href="https://noaa-edab.github.io/ms-keyrun/CompsDat.html" target="_blank"> Age, length, and diet composition decisions - 3 September 2020 </a>
 

### 2020-11-13 docs/GBSurveySet.Rmd
first draft of testing different spatial footprints for GB

---
title: "MS Keyrun - GB survey footprint"
author: "Sean Lucey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE,
                      warning = FALSE)

library(data.table); library(here); library(rgdal); library(Survdat)

```

## Georges Bank - Fisheries Survey Footprint

### Spatial: 2 Approaches for Georges Bank

During the June/July survey <a href="https://noaa-edab.github.io/ms-keyrun/DatasetDimensions.html" target="_blank"> (Year, season, area decisions - 8 July 2020) </a> there were two approaches proposed for the spatial footprint of Georges Bank.  The ideal footprint would be to define using Georges Bank EPU, scale fisheries catch for each species from stat areas to GB EPU footprint using ratios in spatially explicit VTR subset of catch data.

While work is ongoing towards segregating the fisheries dependent data, there is a decision to make about how to use the fisheries independent survey.  The current definition of the EPU is based on ten-minute squares which does not fit with the survey design.  This leaves several options of how to use the survey data.  The first set of options is to use the survey design and assign survey strata to Georges Bank.  The second set of options is to ignore the survey design and either post-stratify or take a simple average within the EPU.

### Option 1 - use survey design
The NEFSC bottom trawl survey uses a stratified random design.  The rationale behind a stratified random design is to minimize variance. Figure \@ref(fig:Survey_map) shows the full range of strata with the Georges Bank EPU overlaid in red. As you will notice, some of the survey strata extend beyond the EPU boundary.  There are several ways that this can be accounted for.  The simplest option is to include all strata that are partially within the EPU boundary.

```{r shapefile load}
coast <- readOGR(here('gis'), 'NES_LME_coast', verbose = F)
coast <- spTransform(coast, CRS('+init=epsg:4326'))

strata <- readOGR(here('gis'), 'strata', verbose = F)
strata <- spTransform(strata, CRS('+init=epsg:4326'))

epu <- readOGR(here('gis'), 'EPU_extended', verbose = F)
epu <- spTransform(epu, CRS('+init=epsg:4326'))

GB.strata <- readOGR(here('gis'), 'EPU_strata', verbose = F)
GB.strata <- spTransform(GB.strata, CRS('+init=epsg:4326'))

MAB <- c(1010:1080, 1100:1120, 1600:1750, 3010:3450, 3470, 3500, 3510)
GB  <- c(1090, 1130:1210, 1230, 1250, 3460, 3480, 3490, 3520:3550)
GOM <- c(1220, 1240, 1260:1290, 1360:1400, 3560:3830)
SS  <- c(1300:1352, 3840:3990)
```

```{r Survey_map}
opar <- par(mar = c(3, 3.5, 2, 2))
plot(0, 0, xlim = c(-77, -65), ylim = c(35, 45), xlab = '', ylab = '', axes = F,
     typ = 'n')
box()
axis(1, at = c(-75, -70, -65), labels = c(expression('75'* degree * ' W'), 
                                          expression('70'* degree * ' W'),
                                          expression('65'* degree * ' W')))

axis(2, at = c(35, 40, 45), labels = c(expression(35 * degree * ' N'),
                                       expression(40 * degree * ' N'),
                                       expression(45 * degree * ' N')), las = T)

plot(strata, add = T)
plot(epu[which(epu@data$EPU == 'GB'), ], add = T, border = 'red', lwd = 2)

plot(coast, col = 'grey', add = T)

legend('bottomright', legend = c('Survey Strata', 'Georges Bank EPU'), 
       col = c('black', 'red'), lty = 1, lwd = 3)

```

Another option is to include only those strata that are at least 50% within the EPU boundary by area.  This was the preferred method of the Northeast IEA program for their State of the Ecosystem report.  The map below shows the strata that were included (Figure \@ref{fig:SOE}).  The medium blue color represents Georges Bank.

```{r SOE}
opar <- par(mar = c(3, 3.5, 2, 2))
plot(0, 0, xlim = c(-77, -65), ylim = c(35, 45), xlab = '', ylab = '', axes = F,
     typ = 'n')
box()
axis(1, at = c(-75, -70, -65), labels = c(expression('75'* degree * ' W'), 
                                          expression('70'* degree * ' W'),
                                          expression('65'* degree * ' W')))

axis(2, at = c(35, 40, 45), labels = c(expression(35 * degree * ' N'),
                                       expression(40 * degree * ' N'),
                                       expression(45 * degree * ' N')), las = T)

plot(strata[which(strata@data$STRATA %in% MAB), ], add = T, col = '#0093D0')
plot(strata[which(strata@data$STRATA %in% GB),  ], add = T, col = '#0055A4')
plot(strata[which(strata@data$STRATA %in% GOM), ], add = T, col = '#00467F')
plot(strata[which(strata@data$STRATA %in% SS),  ], add = T, col = '#007078')
plot(epu[which(epu@data$EPU == 'GB'), ], add = T, border = '#D65f00', lwd = 2)

plot(coast, col = 'grey', add = T)

legend('bottomright', legend = c('Scotian Shelf', 'Gulf of Maine', 'Georges Bank',
                                 'Mid-Atlantic', 'EPU'), 
       fill = c('#007078', '#00467F', '#0055A4', '#0093D0', NA), 
       border = c(rep('black', 4), NA), 
       col = c(rep(0, 4), '#D65f00'), lty = c(rep(NA, 4), 1), lwd = 3, 
       x.intersp = c(rep(-0.5, 4), 1))

```

The difference between these two options would be the inclusion of strata 01100, 01110, and 01120 in the southwest corner of the EPU.

### Option 2 - ignore survey design
Ignoring the survey design leaves us with two options: a straight mean of the EPU or a post-stratified design.  The first option creates large variances which make it difficult to conduct any kind of statistical testing.  Other previous work by EDAB stratified Georges Bank into large strata based on whether the area was considered open access or withing a closed area, inside or outside of US waters, and the shelf break.

```{r GB}
opar <- par(mar = c(3, 3.5, 2, 2))
plot(0, 0, xlim = c(-72, -65), ylim = c(38, 43), xlab = '', ylab = '', axes = F,
     typ = 'n')
box()
axis(1, at = c(-75, -70, -65), labels = c(expression('75'* degree * ' W'), 
                                          expression('70'* degree * ' W'),
                                          expression('65'* degree * ' W')))

axis(2, at = c(35, 40, 45), labels = c(expression(35 * degree * ' N'),
                                       expression(40 * degree * ' N'),
                                       expression(45 * degree * ' N')), las = T)

plot(GB.strata, add = T, border = '#D65f00', lwd = 2)

plot(coast, col = 'grey', add = T)
```

### 2020-11-13 docs/GBSurveySet.Rmd
added plot

@@
 The rationale behind a stratified random design is to minimize variance.
-Figure \@ref(fig:Survey_map) shows the full range of strata with the Georges Bank EPU overlaid in red.
+The figure below shows the full range of strata with the Georges Bank EPU overlaid in red.
 As you will notice, some of the survey strata extend beyond the EPU boundary.
@@
 This was the preferred method of the Northeast IEA program for their State of the Ecosystem report.
-The map below shows the strata that were included (Figure \@ref{fig:SOE}).
+The map below shows the strata that were included.
 The medium blue color represents Georges Bank.
@@
 ```
+
+## Testing the options
+To see if there was an noticeable difference between the three options outlined above (All - Any strata that is at least partially within the GB EPU boundary, SOE - strata set used by the State of the Ecosystem report, and EPU - post-stratified EPU) biomass indices were calculated for the top 11 species that will be included in the MS Keyrun project.
+
+```{r test}
+library(data.table); library(here); library(rgdal); library(Survdat);
+library(ggplot2)
+
+#source(here('Survdat.r'))
+load(here('data_raw/survdat.RData'))
+
+#Top 10(11) species
+top10 <- c(15, 23, 32, 72, 73, 74, 75, 105, 106, 121, 197)
+
+#fix numbers as character
+survdat <- survdat[, STRATUM := as.numeric(STRATUM)]
+survdat <- survdat[, SVSPP   := as.numeric(SVSPP)]
+
+#First with survey strata
+strata <- readOGR(here('gis'), 'strata', verbose = F)
+
+#Generate area table
+strat.area <- getarea(strata, 'STRATA')
+setnames(strat.area, 'STRATA', 'STRATUM')
+
+#Run 1 - All strata for GB ()
+GB.soe   <- c(1090, 1130:1210, 1230, 1250, 3460, 3480, 3490, 3520:3550)
+GB.extra <- 1100:1120
+
+#Subset by season/ strata set
+fall.all <- survdat[SEASON == 'FALL' & STRATUM %in% c(GB.soe, GB.extra), ]
+
+#Run stratification prep
+all.prep <- stratprep(fall.all, strat.area, strat.col = 'STRATUM', area.col = 'Area')
+
+#Calculate stratified means
+all.mean <- stratmean(all.prep, groups = top10, group.col = 'SVSPP', 
+                       strat.col = 'STRATUM')
+all.mean <- all.mean[, list(YEAR, SVSPP, strat.biomass)]
+all.mean[, subset := 'All']
+
+#Run 2 - SOE strata for GB ()
+#Subset by season/ strata set
+fall.soe <- survdat[SEASON == 'FALL' & STRATUM %in% GB.soe, ]
+
+#Run stratification prep
+soe.prep <- stratprep(fall.soe, strat.area, strat.col = 'STRATUM', area.col = 'Area')
+
+#Calculate stratified means
+soe.mean <- stratmean(soe.prep, groups = top10, group.col = 'SVSPP', 
+                      strat.col = 'STRATUM')
+soe.mean <- soe.mean[, list(YEAR, SVSPP, strat.biomass)]
+soe.mean[, subset := 'SOE']
+
+#Run 3 - Use EPU strata
+#Grab strata
+epu <- readOGR(here('gis'), 'EPU_strata', verbose = F)
+
+#Generate area table
+epu.area <- getarea(epu, 'Strata')
+setnames(epu.area, 'Strata', 'EPU')
+
+#Post stratify data if necessary
+survdat.epu <- poststrat(survdat, epu, strata.col = 'Strata')
+setnames(survdat.epu, 'newstrata', 'EPU')
+
+#Subset by season/ strata set
+fall.epu <- survdat.epu[SEASON == 'FALL', ]
+
+#Run stratification prep
+epu.prep <- stratprep(fall.epu, epu.area, strat.col = 'EPU', area.col = 'Area')
+
+#Calculate stratified means
+epu.mean <- stratmean(epu.prep, groups = top10, group.col = 'SVSPP', 
+                      strat.col = 'EPU', poststrat = T)
+epu.mean <- epu.mean[, list(YEAR, SVSPP, strat.biomass)]
+epu.mean[, subset := 'EPU']
+
+#Combine data
+combo.data <- rbindlist(list(all.mean, soe.mean, epu.mean))
+
+survey.plot <- ggplot(combo.data,
+                      aes(x = YEAR, y = strat.biomass)) +
+    geom_point(aes(color = subset)) +
+    facet_wrap(~SVSPP, ncol = 2, nrow = 6, scales = 'free_y')
+
+plot(survey.plot)
+
+```
+
+As you can see, all three spatial footprints give similar trends with only minimal differences in the magnitudes.
+Therefore it was decided at a working group meeting on 11/4/2020 to use the survey strata set from the State of the Ecosystem report.
+This accomplishes two goals.
+First it stays consistent with the survey design of the survey.
+Second it is consistent with other ecosystem products coming out of the Center.

### 2020-11-25 data-raw/plot_shapefiles.R
changed data_raw foldere to data-raw

@@ -8,9 +8,9 @@ ggplot2::ggplot(data=EPU_strata) +
   ggplot2::geom_sf(ggplot2::aes(fill=Strata))
 
 # combines EPU and statistical area
-EPU_stat <- sf::st_read(here::here("gis","GB_EPU_stat.shp"))
+stat_Areas <- sf::st_read(here::here("gis","Statistical_Areas_2010.shp"))
 
-ggplot2::ggplot(data=EPU_stat) +
+ggplot2::ggplot(data=stat_Areas) +
   ggplot2::geom_sf(ggplot2::aes(fill=Id))
 
 # all survey strata
@@ -25,10 +25,14 @@ EPU10 <- sf::st_read(here::here("gis","EPU_extended.shp"))
 ggplot2::ggplot(data=EPU10) +
   ggplot2::geom_sf(ggplot2::aes(fill=EPU))
 
-# NEW GB strata  min square definitions
+# NEW GB EPU def based on survey strata
 GB_strata <- sf::st_read(here::here("gis","GB_SOE_strata.shp"))
 
 ggplot2::ggplot(data=GB_strata) +
   ggplot2::geom_sf(ggplot2::aes(fill=EPU))
 
+# GB strata min square definitions
+GB_EPU_stat <- sf::st_read(here::here("gis","GB_SOE_strata_stat.shp"))
 
+ggplot2::ggplot(data=GB_EPU_stat) +
+  ggplot2::geom_sf(ggplot2::aes(fill=Id))

### 2020-11-25 data-raw/calc_protected_data.R
code to summarize sensitive fishing data - summarized dat added

@@ -1,0 +1,55 @@
+#'Reads in output from VTR/raster/DMIS (Dennis Corvi offshoreWind package)
+#'
+#'Calculates the proportion of landings over time for species in Statareas/EPU intersection
+#'
+#'
+#'@section Process to obtain data
+#'
+#'offshoreWind package must be installed and run on the server (mars)
+#'1. Copy shapefiles "GB_SOE_strata_stat" onto the server
+#'2. install  remotes::install_github("dcorvi/offshoreWind",auth_token="createOne") - it's a private repo
+#'3. offshoreWind::getCallAreaEffort(shapefile_f = "/net/home5/abeet/shapefiles/GB_stat_EPU",
+#'                                  output_f = "/net/home5/abeet/offshoreWind/out",
+#'                                  start = 1994, end = 2018)
+#'4. offshoreWind::get_revenuefile(project_name="Landings_VTR_Geret_Data",
+#'                                output_folder_name = "/net/home5/abeet/offshoreWind/out",
+#'                                control_data=FALSE,
+#'                                save_files = "r",
+#'                                area_data_path="/net/home5/abeet/offshoreWind/out")
+#' see get_landings_vtr_geret.r
+library(magrittr)
+
+calc_protected_data <- function() {
+
+  # lookup table for species name, SVSPP, NESPP3 etc
+  nespp3codes <- mscatch::speciesLookupTable %>%
+    dplyr::distinct(NESPP3) %>% 
+    dplyr::mutate(NESPP3 = as.integer(NESPP3)) %>%
+    dplyr::pull()
+  
+  speciesNames <- mscatch::speciesLookupTable %>%
+    dplyr::select(NESPP3,COMMON_NAME.y) %>%
+    dplyr::mutate(NESPP3 = as.integer(NESPP3))
+  
+  # REVENUEFILE is a df where each record contains landings for a subtrip
+  REVENUEFILE <- readRDS(here::here("data-raw","Landings_VTR_Geret_Data.rds")) # This stores a variable called REVENUEFILE
+  
+  # split the are column to define inside and out, sum landings by year, spp, area and 
+  # select just 11 species that make up 90% of landings
+  data <- REVENUEFILE %>%
[…]

### 2020-11-25 data-raw/create_shapefile_mskeyrun.R
r code to create shape file needed for allocating fishing to GB/stat areas

@@ -1,0 +1,45 @@
+#' create new shapefile which has statareas and GB intersections as seperate region
+#' 
+
+library(magrittr)
+
+create_shapefile_mskeyrun <- function() {
+
+  # read in Statareas and select GB stat areas
+  GBStatAreas <- c(cfdbs::EPUs$data$GB,537)
+  statAreas <- sf::st_read(dsn=here::here("gis","Statistical_Areas_2010.shp"),quiet=T) %>%
+    dplyr::filter(Id %in% GBStatAreas)
+  
+  # read in GB EPU
+  
+  GB <- sf::st_read(dsn=here::here("gis","GB_SOE_strata.shp"),quiet=T) %>%
+    dplyr::filter(EPU == "GB")
+  
+  GB <- sf::st_transform(GB,crs=sf::st_crs(statAreas))
+  
+  # intersection of GB with stat areas
+  GBStat <- sf::st_intersection(statAreas,GB) %>%
+    dplyr::mutate(Id=paste0(Id,"_in")) %>% 
+    dplyr::select(Id,EPU)
+  # non intersections of GB with stat areas
+  GBNotStat <- sf::st_difference(statAreas,GB) %>%
+    dplyr::mutate(Id=paste0(Id,"_out")) %>% 
+    dplyr::select(Id,EPU)
+  # combine shape files 
+  GBShape <- rbind(GBStat,GBNotStat)
+  
+  # write to file
+  sf::write_sf(GBShape,dsn=here::here("gis","GB_SOE_strata_stat.shp"))
+  
+  # plots combined shapefile
+  p1 <- ggplot2::ggplot(data=GBShape) +
+    ggplot2::geom_sf(mapping=ggplot2::aes(fill=Id))
+  print(p1)
+  
+  # plots nicer, overlays statareas on GB
+  p2 <- ggplot2::ggplot(data=statAreas) +
+    ggplot2::geom_sf(mapping=ggplot2::aes(fill=Id),alpha =0.3,fill ="grey") +
+    ggplot2::geom_sf(data=GB,mapping=ggplot2::aes(fill=GB),alpha =0.3,fill ="blue")
+  
+  print(p2)
+}

### 2020-11-25 data-raw/plot_landings_proportions.R
plotting functions to create statarea/species grid plots - proportion plot - total landings plot

@@ -1,0 +1,66 @@
+#'Reads in output from VTR/raster/DMIS (Dennis Corvi offshoreWind package)
+#'
+#'Calculates the proportion of landings over time for species in Statareas/EPU intersection
+#'
+#'
+#'@section Process to obtain data
+#'
+#'offshoreWind package must be installed and run on the server (mars)
+#'1. Copy shapefiles "GB_SOE_strata_stat" onto the server
+#'2. install  remotes::install_github("dcorvi/offshoreWind",auth_token="createOne") - it's a private repo
+#'3. offshoreWind::getCallAreaEffort(shapefile_f = "/net/home5/abeet/shapefiles/GB_stat_EPU",
+#'                                  output_f = "/net/home5/abeet/offshoreWind/out",
+#'                                  start = 1994, end = 2018)
+#'4. offshoreWind::get_revenuefile(project_name="Landings_VTR_Geret_Data",
+#'                                output_folder_name = "/net/home5/abeet/offshoreWind/out",
+#'                                control_data=FALSE,
+#'                                save_files = "r",
+#'                                area_data_path="/net/home5/abeet/offshoreWind/out")
+#' see get_landings_vtr_geret.r
+library(magrittr)
+
+calc_prop <- function(landings,inout) {
+  prop <- sum(landings*(inout=="in"))/sum(landings)
+  return(prop)
+}
+
+plot_landings_proportions <- function(species="cols") {
+   # REVENUEFILE is a df where each record contains landings for a subtrip
+  REVENUEFILE <- readRDS(here::here("data-raw","Landings_VTR_Geret_Data_summarized.rds"))
+  
+  # calculate proportion of landings inside GB section of stat area
+  data <- REVENUEFILE %>% 
+    dplyr::group_by(YEAR,AREA,NESPP3,COMMON_NAME) %>% 
+    dplyr::summarise(PROP_GB = calc_prop(TOTALLANDINGS,InOut))
+
+  if(species=="cols") {
+    # plot grid area by species
+    p <- ggplot2::ggplot(data=data) +
+      ggplot2::geom_line(mapping = ggplot2::aes(x=YEAR,y=PROP_GB)) +
+      ggplot2::facet_grid(cols = ggplot2::vars(COMMON_NAME), rows= ggplot2::vars(AREA)) +
[…]

### 2020-11-25 docs/allocateLandingsEPU.Rmd
allocation of landings to GB documentation

---
title: "Allocate landings to Georges Bank EPU"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
options(warn=-1)
```

## Spatial: Approaches for Georges Bank [Option 1](https://noaa-edab.github.io/ms-keyrun/GBSurveySet.html#opt1)

The spatial footprint for the Georges Bank EPU (Ecological Production Unit) is defined as the set of survey strata that are at 50% within the 10 min square definition of the EPU.

```{r GBfoot, echo=F, eval=T}
crs <- 4326
# NEW GB EPU def based on survey strata
coast <- sf::st_read(here::here("gis", 'NES_LME_coast.shp'), quiet = T) %>%
  sf::st_transform(.,crs=crs)
GB_strata <- sf::st_read(here::here("gis","GB_SOE_strata.shp"),quiet=T) %>%
    sf::st_transform(.,crs=crs)

ggplot2::ggplot(data=coast) +
  ggplot2::geom_sf() +
  ggplot2::geom_sf(data=GB_strata,col="black",fill="grey") + 
  ggplot2::coord_sf(xlim = c(-76,-65), ylim = c(38,44))

```

Overlay the surrounding statistical areas:

```{r GBfootstat,echo=F, eval=T}
crs <- 4326
# read in coastline and GB shape file
coast <- sf::st_read(here::here("gis", 'NES_LME_coast.shp'), quiet = T) %>%
  sf::st_transform(.,crs=crs)
GB_strata <- sf::st_read(here::here("gis","GB_SOE_strata_stat.shp"),quiet=T) %>%
    sf::st_transform(.,crs=crs)

# separate polygons inside GB from those outside. Calculate centroid of polygon for label
GB_in <- GB_strata %>%
  dplyr::filter(grepl("in",Id) )
GB_out <- GB_strata %>%
  dplyr::filter(grepl("out",Id) ) %>% 
  dplyr::mutate(Id=stringr::str_split_fixed(Id,"_",2)[,1])
centroids <- sf::st_coordinates(sf::st_centroid(GB_out))
GB_out <- cbind(GB_out,centroids)

statAreas <- GB_out %>%
  sf::st_drop_geometry() %>%
  dplyr::distinct(Id) %>%
  dplyr::pull()

#plot map
ggplot2::ggplot(data=coast) +
  ggplot2::geom_sf() +
  ggplot2::geom_sf(data=GB_out,fill="grey") + 
  ggplot2::geom_text(data=GB_out,ggplot2::aes(x=X,y=Y,label=Id),size=2)    +
  ggplot2::geom_sf(data=GB_in,col="black") + 
  ggplot2::coord_sf(xlim = c(-76,-65), ylim = c(38,44))

```

Fishing data are available by statistical area. The statistical areas surrounding Georges Bank EPU (`r statAreas`) contain catch from Georges Bank and its neighboring EPUs. This data needs to be partitioned and catch allocated to the Georges Bank EPU.

## Method for partioning catch data

VTR data is used ... (DePiper, 20xx) and combined with DMIS data. (Need to expand on this). These methods are available within the [`offshoreWind`](https://github.com/dcorvi/offshoreWind) R package. It is necessary to run the package from the server (the `mars` server is recommended).

The [top ranked species](https://noaa-edab.github.io/ms-keyrun/SpeciesEnvDat.html#focalSpecies) of interest are shown below. Each panel depicts the proportion of landings (for each species) in Georges Bank relative to the statistical area.

```{r calc_props_sp, echo = F, eval=T, out.width="100%"}
source(here::here("data-raw","plot_landings_proportions.r"))
plot_landings_proportions(species="cols")

```

The same data is represented in the plot below but the species are plotted as rows instead of columns
```{r calc_props_st, echo = F, eval=T, out.width="100%"}
plot_landings_proportions(species="rows")

```

The total landings of each species by stat area is shown below. This should ideally be compared to landings from [comlandr](https://github.com/NOAA-EDAB/comlandr). We should also determine how landings on the Canadian side of the Hague line should be handled

```{r calc_lands, echo = F, eval=T, out.width="100%"}
source(here::here("data-raw","plot_landings.r"))
plot_landings()

```

The majority of landings (for a species) occurs in statistical areas where the proportion allocated to Georges Bank is fairly constant.

### 2020-11-27 docs/DatasetDimensions.Rmd
hyperlink added to allocateLandingsEPU page

@@
 
-Ideal: Define using Georges Bank EPU, scale fisheries catch for each species from stat areas to GB EPU footprint using ratios in spatially explicit VTR subset of catch data.
+Ideal: Define using Georges Bank EPU, scale fisheries catch for each species from stat areas to GB EPU footprint using [ratios in spatially explicit VTR subset of catch data](allocateLandingsEPU.html).
 Work towards this ideal, comparing coverage of VTR data by species and different methods for scaling as part of data documentation.

### 2020-11-27 docs/index.Rmd
Added GB survey footprint discussion
update index
hyperlink to allocateLandingsEPU added

@@
 
+*   <a href="https://noaa-edab.github.io/ms-keyrun/allocateLandingsEPU.html" target="_blank"> Allocate landings to EPU - 25 Nov 2020 </a>
+
 <a href="https://noaa-edab.github.io/ms-keyrun/SpeciesEnvDat.html" target="_blank"> Species, environment, data package decisions - 5 August 2020 </a>
@@
 ## Dataset source and methods documentation
+<a href="https://noaa-edab.github.io/ms-keyrun/GBSurveySets.html" target="_blank"> Which Georges Bank survey footprint - 4 November 2020 </a>
 

### 2020-11-27 docs/index.Rmd

--- 2020-09-03 docs/index.Rmd
@@
 ## Dataset source and methods documentation
+<a href="https://noaa-edab.github.io/ms-keyrun/GBSurveySet.html" target="_blank"> Which Georges Bank survey footprint - 4 November 2020 </a>
 

### 2020-12-02 docs/allocateLandingsEPU.Rmd
add citations to allocateLandingsEPU.rmd

@@
 title: "Allocate landings to Georges Bank EPU"
-output: html_document
+author: Andy Beet
+date: "`r format(Sys.time(), '%d %B, %Y')`"
+output: 
+  html_document:
+    code_fold: hide
+csl: apa-annotated-bibliography.csl
+bibliography: references.bib
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, 
+                      message = FALSE,
+                      warning = FALSE)
 library(magrittr)
@@
 
-```{r GBfoot, echo=F, eval=T}
+```{r GBfoot, eval=T}
 crs <- 4326
@@
 
-```{r GBfootstat,echo=F, eval=T}
+```{r GBfootstat, eval=T}
 crs <- 4326
@@
 
-## Method for partioning catch data
+## Method for partitioning catch data
 
-VTR data is used ...
-(DePiper, 20xx) and combined with DMIS data.
-(Need to expand on this).
-These methods are available within the [`offshoreWind`](https://github.com/dcorvi/offshoreWind) R package.
+```{r vtrdmisdata, eval=T}
+data <- readRDS(here::here("data-raw","Landings_VTR_Geret_Data_summarized.rds"))
+yearRange <- data %>% dplyr::distinct(YEAR) %>% range
+```
+
+Methods found in [@depiper2014; & @benjamins2018] are implemented in the [`offshoreWind`](https://github.com/dcorvi/offshoreWind) R package.
 It is necessary to run the package from the server (the `mars` server is recommended).
+
+In summary:
+
+-   VTR data are joined with observer data and a model is specified ( @depiper2014) to assign ...
+
+-   Raster files are then created for each year
+
+-   expand ..
+
+-   The landings data span the period from `r paste0(yearRange,collapse="-")`.
+
+-   Landings are assigned to each of the statistical areas (`r statAreas`)
+
+-   The landings with each statistical area (for each year) are then allocated to each portion of a statistical area (within Georges Bank and outside Georges Bank)
+
+-   The proportion of landings attributed to Georges Bank are then calculated and displayed below
 
@@
 
-```{r calc_props_sp, echo = F, eval=T, out.width="100%"}
+```{r calc_props_sp, eval=T, out.width="100%"}
 source(here::here("data-raw","plot_landings_proportions.r"))
@@
 The same data is represented in the plot below but the species are plotted as rows instead of columns
-```{r calc_props_st, echo = F, eval=T, out.width="100%"}
+
+```{r calc_props_st, eval=T, out.width="100%"}
 plot_landings_proportions(species="rows")
@@
 
-```{r calc_lands, echo = F, eval=T, out.width="100%"}
+```{r calc_lands,  eval=T, out.width="100%"}
 source(here::here("data-raw","plot_landings.r"))
@@
 The majority of landings (for a species) occurs in statistical areas where the proportion allocated to Georges Bank is fairly constant.
+
+## References

### 2020-12-04 docs/allocateLandingsEPU.Rmd
changes to bookdown::html_document2 for fig referencing

@@
 output: 
-  html_document:
+  bookdown::html_document2:
     code_fold: hide
@@
 
-```{r GBfootstat, eval=T}
+```{r GBfootstat, eval=T, fig.cap="Statistical areas intersected by Georges Bank EPU",fig.align='center'}
 crs <- 4326
@@
 
-Methods found in [@depiper2014; & @benjamins2018] are implemented in the [`offshoreWind`](https://github.com/dcorvi/offshoreWind) R package.
-It is necessary to run the package from the server (the `mars` server is recommended).
+Methods found in [@depiper2014; & @benjamins2018] and the [`offshoreWind`](https://github.com/dcorvi/offshoreWind) package are briefly summarized below.
+It is necessary to run the package from the `mars` server.
 
-In summary:
+-   VTR data are joined with observer data (since observer data is considered reliable).
+A duration model is fitted to characterize a spatial footprint (rather than a single point location) associated with a trip ( @depiper2014).
+Gear type and trip length (which includes multiple hauls) influence the spatial precision.
 
--   VTR data are joined with observer data and a model is specified ( @depiper2014) to assign ...
+-   The duration model is then broadly applied to all VTR data. The result being a set of maps (or rasters). Each map/raster depicts the spatial probability of fishing for a trip.
 
--   Raster files are then created for each year
+-   The `offshoreWind` package then uses these rasters combined with commerical fishing data (DMIS) to determine the landings assigned to any spatial footprint supplied.
+This is achieved by calculating the overlap of the rasters with the user supplied spatial footprint.
 
--   expand ..
-
--   The landings data span the period from `r paste0(yearRange,collapse="-")`.
-
--   Landings are assigned to each of the statistical areas (`r statAreas`)
-
--   The landings with each statistical area (for each year) are then allocated to each portion of a statistical area (within Georges Bank and outside Georges Bank)
-
--   The proportion of landings attributed to Georges Bank are then calculated and displayed below
+The shape file shown in Figure \@ref(fig:GBfootstat) is supplied to the `offshoreWind` package and the landings with each statistical area (for each year) are then allocated to each portion of a statistical area (within Georges Bank and outside Georges Bank).
+The proportion of landings attributed to Georges Bank are then calculated and displayed below.
 

### 2020-12-04 docs/allocateLandingsEPU.Rmd
remove chapter numbering

@@
     code_fold: hide
+    number_sections: false
 csl: apa-annotated-bibliography.csl

### 2021-02-02 docs/SimData.Rmd

---
title: "Generating Simulated Data"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(here)
library(tidyverse)  
library(atlantisom)
library(ggthemes)

```

## Simulating input data from an ecosystem model

Here we use existing [Atlantis](https://github.com/Atlantis-Ecosystem-Model/Atlantis_example_and_instructions) ecosystem model output to generate input datasets for a variety of simpler population models, so that the performance of these models can be evaluated against known (simulated) ecosystem dynamics. Atlantis models simulate a wide range of physical and ecological processes, include a full food web, and can be run using different climate forcing, fishing, and other scenarios.

We extract simulated data using the R package [atlantisom](https://github.com/r4atlantis/atlantisom). The purpose of atlantisom is to use existing Atlantis model output to generate input datasets for a variety of models, so that the performance of these models can be evaluated against known (simulated) ecosystem dynamics. Atlantis models can be run using different climate forcing, fishing, and other scenarios. Users of atlantisom will be able to specify fishery independent and fishery dependent sampling in space and time, as well as species-specific catchability, selectivty, and other observation processes for any Atlantis scenario. Internally consistent multispecies and ecosystem datasets with known observation error characteristics will be the atlantisom outputs, for use in individual model performance testing, comparing performance of alternative models, and performance testing of model ensembles against "true" Atlantis outputs.

## Species in the ms-keyrun dataset

Our initial species selection includes 11 single species groups from the Atlantis NOBA model.

```{r spp-table}

fgs <- load_fgs(here("simulated-data/atlantisoutput/NOBA_March_2020"), "nordic_groups_v04.csv")

lname <- data.frame(Latin = c("*Hippoglossoides platessoides*",
                              "*Reinhardtius hippoglossoides*",
                              "*Scomber scombrus*",
                              "*Melongrammus aeglefinus*",
                              "*Pollachius virens*",
                              "*Sebastes mentella*",
                              "*Micromesistius poutassou*",
                              "*Clupea harengus*",
                              "*Gadus morhua*",
                              "*Boreogadus saida*",
                              "*Mallotus villosus*"),
                    Code = c("LRD", "GRH", "MAC", "HAD", "SAI", "RED", 
                             "BWH", "SSH", "NCO", "PCO", "CAP")
)

sppsubset <- merge(fgs, lname, all.y = TRUE)
spptable <- sppsubset %>% 
  arrange(Index) %>%
  select(Name, Long.Name, Latin)

knitr::kable(spptable, col.names = c("Model name", "Full name", "Latin name"))

```

Below we will double check survey census outputs against the true values, and then test some survey specifications that may be reasonable for multispecies model testing. Finally we will apply survey functions to the detailed diet data.

## Generating a census dataset (February 2021)

### Configuration files specified once

`NOBA2config.R` specifies location and names of files needed for atlantisom to initialize:

```{r, code = readLines("./simulated-data/config/NOBA2Config.R"), eval=F}
```

`omdimensions.R` standardizes timesteps, etc. (this is part of atlantisom and should not need to be changed by the user):

```{r, code = readLines("./simulated-data/config/omdimensions.R"), eval=F}
```

### Change these survey and fishery config files

`mssurvey_spring.R` and `mssurvey_fall.R` configure the fishery independent surveys (in this census test, surveys sample all model polygons in all years and have efficiency of 1 for all species, with no size selectivity):

```{r, code = readLines("./simulated-data/config/mssurvey_spring.R"), eval=F}
```

```{r, code = readLines("./simulated-data/config/mssurvey_fall.R"), eval=F}
```

`msfishery.R` configures the fishery dependent data:

```{r, code = readLines("./simulated-data/config/msfishery.R"), eval=F}
```

### Run atlantisom and save outputs

This was run previously but true datasets are generated as follows, with outputs saved to the `atlantisoutput` folder (not kept on github due to size):

```{r, eval=FALSE}

NOBAom <- om_init(here("simulated-data/config/NOBA2config.R"))

NOBAom_ms <- om_species(sppsubset$Name, NOBAom)

NOBAom_ms_ind <- om_index(usersurvey = c(here("simulated-data/config/mssurvey_spring.R"), 
                                         here("simulated-data/config/mssurvey_fall.R")), 
                           userfishery = here("simulated-data/config/msfishery.R"),
                           omlist_ss = NOBAom_ms, 
                           n_reps = 1, 
                           save = TRUE)

NOBAom_ms_comp <- om_comps(usersurvey = c(here("simulated-data/config/mssurvey_spring.R"), 
                                         here("simulated-data/config/mssurvey_fall.R")), 
                           userfishery = here("simulated-data/config/msfishery.R"),
                           omlist_ss = NOBAom_ms, 
                           n_reps = 1, 
                           save = TRUE)

NOBAom_ms_diet <- om_diet(config = here("config", "NOBA2config.R"),
                          dietfile = "NOBADetDiet.gz",
                          usersurvey = c(here("config/mssurvey_spring.R"), 
                                         here("config/mssurvey_fall.R")), 
                          omlist_ss = NOBAom_ms, 
                          n_reps = 1, 
                          save = TRUE)

```

### Write to model input files

##

### 2021-02-04 docs/SimData.Rmd
rerun census surveys and write msspm bio files

@@
 
-This was run previously but true datasets are generated as follows, with outputs saved to the `atlantisoutput` folder (not kept on github due to size):
+True datasets are generated as follows, using `atlantisom` wrapper functions `om_init` to assemble initial true atlantis data, `om_species` to subset true data for desired species, `om_index` to generate survey biomass and total catch biomass indices, `om_comps` to generate age and length compositions and average weight at age from surveys and fisheries, and `om_diet` to generate diet from surveys.
+Outputs are saved to the `atlantisoutput` folder (not kept on github due to size):
 
@@
 
-NOBAom <- om_init(here("simulated-data/config/NOBA2config.R"))
+NOBAom <- om_init(here("simulated-data/config/NOBA2Config.R"))
 
 NOBAom_ms <- om_species(sppsubset$Name, NOBAom)
+
+#need to change internal call to source in atlantisom om_index om_comps and om_diet functions
+#expecting a config folder in same directory as rmd
+#this is a workaround
+
+dir.create(file.path(here("docs/config")))
+
+file.copy(here("simulated-data/config/omdimensions.R"), here("docs/config/omdimensions.R"))
 
@@
 
-NOBAom_ms_diet <- om_diet(config = here("config", "NOBA2config.R"),
+NOBAom_ms_diet <- om_diet(config = here("simulated-data/config", "NOBA2config.R"),
                           dietfile = "NOBADetDiet.gz",
-                          usersurvey = c(here("config/mssurvey_spring.R"), 
-                                         here("config/mssurvey_fall.R")), 
+                          usersurvey = c(here("simulated-data/config/mssurvey_spring.R"), 
+                                         here("simulated-data/config/mssurvey_fall.R")), 
                           omlist_ss = NOBAom_ms, 
@@
 
+unlink(here("docs/config"), recursive = TRUE)
 ```
@@
 
-##
+Example input files are in the `model-input-files` directory.
+First we write the fairly simple biomass and catch inputs for the multispecies surplus production model MSSPM from our census survey outputs.
+
+```{r write-msspm}
+
+o.name <- here("simulated-data/msinput/msspm/")
+
+#read in survey biomass data
+survObsBiom <- read_savedsurvs(d.name, 'survB') #reads in all surveys
+
+#multiple surveys named in list object, use as filename
+for(s in names(survObsBiom)){
+  #arrange into wide format: year, Species1, Species2 ... and write csv
+  svbio <- survObsBiom[[s]][[1]] %>%
+    calc_timestep2time(d.name, ., run.prm.file) %>%
+    dplyr::select(species, year, atoutput) %>%
+    tidyr::spread(species, atoutput) %>%
+    write_csv(paste0(o.name,s,".csv"))
+}
+
+#read in catch data
+read_savedfisheries <- function(dir, type){
+
+  datlook <- data.frame(dattype = c('Catch', 'catchAge', 'catchLen', 'catchWtage', 'catchAnnAge', 'catchAnnWtage'),
+                        pattern = c("*fishCatch.rds", "*fishObsAgeComp.rds", "*fishObsLenComp.rds", "*fishObsWtAtAge.rds",
+                                    "*fishObsFullAgeComp.rds", "*fishObsFullWtAtAge.rds"))
+
+  fisheries <- list.files(path=dir, pattern = as.character(datlook$pattern[datlook$dattype %in% type]), full.names = TRUE)
+
+  fishery.name <-  str_match(fisheries, paste0(scenario.name,"_\\s*(.*?)\\s",datlook$pattern[datlook$dattype==type]))[,2]
+
+  dat <- lapply(fisheries, readRDS)
+
+  names(dat) <- fishery.name
+
+  return(dat)
+}
+
+catchbio_ss <- read_savedfisheries(d.name, 'Catch')
+
+#arrange into wide format: Year, Species1, Species2 ...
+
+#write to csv
+
+#write readme for metadata
+
+```
+
+## Generating a survey dataset

### 2021-02-04 simulated-data/config/NOBA2Config.R

@@ -1,4 +1,4 @@
-d.name <- here("atlantisoutput","NOBA_march_2020")
+d.name <- here("simulated-data/atlantisoutput","NOBA_march_2020")
 functional.groups.file <- "nordic_groups_v04.csv"
 biomass.pools.file <- "nordic_biol_v23.nc"
 biol.prm.file <- "nordic_biol_incl_harv_v_007_3.prm"

### 2021-02-04 simulated-data/config/msfishery.R

@@ -1,4 +1,6 @@
 # Default fishery configuration here is a census
+fishery.name="census"
+
 # Inherits species from input omlist_ss
 fishspp <- omlist_ss$species_ss 
 

### 2021-02-04 simulated-data/config/mssurvey_fall.R

@@ -44,6 +44,9 @@ survselex <- data.frame(species=rep(names(annages), n_annages), #
                         agecl=unlist(sapply(n_annages,seq)),
                         selex=rep(1.0,sum(n_annages)))
 
+survselex.agecl <- survselex 
+
+
 # effective sample size needed for sample_fish
 # this effective N is high but not equal to total for numerous groups
 surveffN <- data.frame(species=survspp, effN=rep(100000, length(survspp)))
@@ -59,3 +62,6 @@ lenage_cv <- 0.1
 # max size bin for length estimation, function defaults to 150 cm if not supplied
 maxbin <- 200
 
+# diet sampling parameters
+alphamult <- 10000000
+unidprey <- 0

### 2021-02-04 simulated-data/config/mssurvey_spring.R

@@ -44,6 +44,9 @@ survselex <- data.frame(species=rep(names(annages), n_annages), #
                         agecl=unlist(sapply(n_annages,seq)),
                         selex=rep(1.0,sum(n_annages)))
 
+survselex.agecl <- survselex
+
+
 # effective sample size needed for sample_fish
 # this effective N is high but not equal to total for numerous groups
 surveffN <- data.frame(species=survspp, effN=rep(100000, length(survspp)))
@@ -58,3 +61,9 @@ lenage_cv <- 0.1
 
 # max size bin for length estimation, function defaults to 150 cm if not supplied
 maxbin <- 200
+
+# diet sampling parameters
+alphamult <- 10000000
+unidprey <- 0
+
+

### 2021-02-05 docs/SimData.Rmd
generating test simulated datasets and msspm inputs

@@
 
-```{r, code = readLines("./simulated-data/config/NOBA2Config.R"), eval=F}
+```{r, code = readLines(here("/simulated-data/config/NOBA2Config.R")), eval=F}
 ```
@@
 
-```{r, code = readLines("./simulated-data/config/omdimensions.R"), eval=F}
+```{r, code = readLines(here("/simulated-data/config/omdimensions.R")), eval=F}
 ```
@@
 
-```{r, code = readLines("./simulated-data/config/mssurvey_spring.R"), eval=F}
+```{r, code = readLines(here("/simulated-data/config/mssurvey_spring.R")), eval=F}
 ```
 
-```{r, code = readLines("./simulated-data/config/mssurvey_fall.R"), eval=F}
+```{r, code = readLines(here("/simulated-data/config/mssurvey_fall.R")), eval=F}
 ```
@@
 
-```{r, code = readLines("./simulated-data/config/msfishery.R"), eval=F}
+```{r, code = readLines(here("/simulated-data/config/msfishery.R")), eval=F}
 ```
@@
 
-```{r write-msspm}
+```{r write-msspm, eval=FALSE}
 
 o.name <- here("simulated-data/msinput/msspm/")
+
+stepperyr <- if(NOBAom_ms$runpar$outputstepunit=="days") 365/NOBAom_ms$runpar$toutinc
 
@@
   svbio <- survObsBiom[[s]][[1]] %>%
-    calc_timestep2time(d.name, ., run.prm.file) %>%
+    #calc_timestep2time(d.name, ., run.prm.file) %>% #this gives an arbitrary start year of 1970
+    dplyr::mutate(year = ceiling(time/stepperyr)) %>%
     dplyr::select(species, year, atoutput) %>%
@@
 
-#arrange into wide format: Year, Species1, Species2 ...
-
-#write to csv
+for(c in names(catchbio_ss)){
+  #arrange into wide format: year, Species1, Species2 ... and write csv
+  catchbio <- catchbio_ss[[c]][[1]] %>%
+    dplyr::mutate(year = time/365) %>%
+    dplyr::select(species, year, atoutput) %>%
+    tidyr::spread(species, atoutput) %>%
+    write_csv(paste0(o.name,"fishery_",c,".csv"))
+}
 
 #write readme for metadata
+meta <- c("Simulated data from Norweigan-Barents Atlantis Model by Hansen et al. \n\n",
+          paste("Atlantis outputs stored in", d.name, '\n\n'),
+          paste("Biomass output for surveys", names(survObsBiom), '\n\n'),
+          paste("Fishery catch output for fisheries", names(catchbio_ss), '\n\n'),
+          "Biomass and catch output units are annual tons.\n\n",
+          "Survey specifications:\n",
+          "----------------------\n",
+          read_file(here("/simulated-data/config/mssurvey_fall.R")),
+          "\n",
+          "----------------------\n",
+          read_file(here("/simulated-data/config/mssurvey_spring.R")),
+          "\n",
+          "----------------------\n",
+          read_file(here("/simulated-data/config/mssurvey_fall_01.R")),
+          "\n",
+          "----------------------\n",
+          read_file(here("/simulated-data/config/mssurvey_spring_01.R")),
+          "\n\n",
+          "Fishery specifications:\n",
+          "----------------------\n",
+          read_file(here("/simulated-data/config/msfishery.R")),
+          "\n",
+          "----------------------\n",
+          read_file(here("/simulated-data/config/msfishery_01.R")),
+          "\n"
+          )
+
+cat(as.character(meta), file=paste0(here("simulated-data/msinput/msspm/"),"README.txt"))
 

### 2021-02-22 data_raw/diet_data_pull.R
updated oracle table

@@ -6,7 +6,7 @@ channel <- dbutils::connect_to_database(server="sole",uid="slucey")
 allfh.qry <- "select year, season, cruise6, station, stratum, tow, declat, declon,
              svspp, pdid, pdgutw, pdgutv,
              pynam, pyamtw, pyamtv, perpyw, perpyv
-             from FHDBS.ALLFH_FEAST_17"
+             from FHDBS.ALLFH_FEAST"
 
 allfh <- data.table::as.data.table(DBI::dbGetQuery(channel, allfh.qry))
 

### 2021-02-22 docs/SimData.Rmd

@@
 
+The MSSPM needs diet compostion information as well. Here we supply two survey time series of diet data, one census and a trial survey dataset with more typical survey coverage issues.
+
+```{r, write-msspm-diet}
+
+all_diets <- read_savedsurvs(d.name, 'survDiet')
+
+# aggregate over ages (survey diet is in proportion by age)
+
+trueB_age_allbox <- NOBAom_ms$truebio_ss %>%
+  #sum over polygons
+  aggregate(dat$atoutput,list(dat$species,dat$agecl,dat$time, dat$prey),sum)
+
+```
+
 ## Generating a survey dataset

### 2021-02-22 docs/allocateLandingsEPU.Rmd

@@
 
-```{r GBfoot, eval=T}
+```{r GBfoot, eval=T, fig.cap="Georges Bank EPU", fig.align='center'}
 crs <- 4326
@@
 
--   The `offshoreWind` package then uses these rasters combined with commerical fishing data (DMIS) to determine the landings assigned to any spatial footprint supplied.
+-   The `offshoreWind` package then uses these rasters combined to determine the landings assigned to any spatial footprint supplied.
 This is achieved by calculating the overlap of the rasters with the user supplied spatial footprint.
@@
 
-```{r calc_props_sp, eval=T, out.width="100%"}
+```{r calc_props_sp, eval=T, out.width="100%",fig.align='center',fig.cap="Proportion of landings in GB relative to statistical area"}
 source(here::here("data-raw","plot_landings_proportions.r"))
@@
 
-```{r calc_props_st, eval=T, out.width="100%"}
+```{r calc_props_st, eval=T, out.width="100%",fig.align='center',fig.cap="Proportion of landings in GB relative to statistical area"}
 plot_landings_proportions(species="rows")
@@
 
-```{r calc_lands,  eval=T, out.width="100%"}
+```{r calc_lands,  eval=T, out.width="100%",fig.align='center',fig.cap="Total landings by statistical area"}
 source(here::here("data-raw","plot_landings.r"))

### 2021-02-26 R/fleetClusterAnalysis.R
add r code to get landings data by gear and perform cluster analysis

@@ -1,0 +1,88 @@
+#' Perform cluster analysis
+#' 
+#' Uses all gear Types and species contributing to x% of landings for each gear type.
+#' Reads in preprocessed data from get_landings_by_gear.r
+#' 
+#' @return object of class \code{agnes} representing the clustering. Use \code{plot} top display dendrogram
+#' 
+#' @example 
+#' \dontrun{
+#' clusterObj <- cluster_analysis_all_gears()
+#' # dendrogram plot
+#' plot(clusterObj,ask=T,which.plots=2,main="Complete gear list using 90% landing from each gear",xlab="")
+#' }
+#' 
+#' 
+
+library(magrittr)
+
+
+
+cluster_analysis_all_gears <- function(){
+
+  # read in landings data by gear/species aggregated over time
+  allGearData <- readRDS(here::here("data","gearLandingsBySpecies.rds"))
+  # read in all gear codes
+  gearCodes <- readRDS(here::here("data","gearCodeTable.rds"))
+  
+  # split long names by comma and select first name
+  gearCodes <- gearCodes$data %>% 
+    dplyr::select(NEGEAR2,GEARNM) %>%
+    dplyr::mutate(GEARNM2=unlist(lapply(stringr::str_split(gearCodes$data$GEARNM,","),`[[`,1))) %>% 
+    dplyr::select(-GEARNM) %>% 
+    dplyr::distinct(NEGEAR2,GEARNM2)
+  
+  # find unique gear codes, make single string if has multiple entries and create new table
+  uniqueCodes <- unique(gearCodes$NEGEAR2)
+  gearCodesUpdate <- data.frame(matrix(nrow = length(uniqueCodes),ncol=2))
+  names(gearCodesUpdate) <- c("NEGEAR2","GEARName")
+  for (icode in 1:length(uniqueCodes)) {
+    ind <- gearCodes$NEGEAR2 == uniqueCodes[icode]
+    if(sum(ind) > 1) {
+      newCode <- paste0(gearCodes$GEARNM2[ind],"/",collapse=" ")
+      gearCodesUpdate$NEGEAR2[icode] <- as.numeric(uniqueCodes[icode])
+      gearCodesUpdate$GEARName[icode] <-substr(newCode,1,10) # shortens name to 10 characters for ease in plotting
+    } else {
+      gearCodesUpdate$NEGEAR2[icode] <- as.numeric(uniqueCodes[icode])
[…]

### 2021-02-26 R/get_landings_by_gear.R

@@ -1,0 +1,64 @@
+#' Read in landings data by gear and save as RDS
+#' 
+#' cfdbs is accessed and landings pulled by gear type for GB over time
+#' Data then aggregated over trips for each year and saved as RDS
+#' Data also further aggregated over time and saved as RDS
+#' 
+#' @param channel an Object inherited from \link[DBI]{DBIConnection-class}. This object is used to communicate with the database engine. (see \code{\link[dbutils]{connect_to_database}})
+#' @param percentLanded Numeric scalar. Proportion [0,1] of landings to filter by
+#' @param area Numeric vector. Statistical areas defining Area from which landings data should be pulled.
+#' @param outDir Character string. Full path to directory where output should be saved. Default=here::here("data")
+#' 
+#' @return Two RDS files saved
+
+library(magrittr)
+
+get_landings_by_gear <- function(channel,percentLanded=0.90,area=c(cfdbs::EPUs$data[["GB"]],537),outDir=here::here("data")) {
+
+  mGearData <- NULL
+  mTimeGearData <- NULL
+  gearCodes <- cfdbs::fleets$codes
+  message(paste0("Pulling data for following gear Codes: ",paste0(gearCodes,collapse=",")))
+  for (gear in gearCodes){
+    message(paste0("gear Code = ",gear))
+    gearData <- cfdbs::get_landings(channel,area=area,gear=gear, year="all",tonnage = "all",species = "all")
+    
+    # sum over trips by year group by gear and species
+    aggTimeGear <- gearData$data %>% 
+      dplyr::group_by(YEAR,NESPP3) %>% 
+      dplyr::summarise(totsplandlb=sum(as.numeric(SPPLNDLB)),n=dplyr::n()) %>% 
+      dplyr::ungroup() %>%
+      dplyr::mutate(NEGEAR2=gear)
+    
+    # sum over time group by gear and species
+    aggGear <- gearData$data %>% 
+      dplyr::group_by(NESPP3) %>% 
+      dplyr::summarise(totsplandlb=sum(as.numeric(SPPLNDLB)),n=dplyr::n()) %>% 
+      dplyr::ungroup()
+    
+    # select species that account for x% of landings
+    orderSpecies <- aggGear %>% 
+      dplyr::group_by(NESPP3) %>% 
+      dplyr::summarise(sumOverYears=sum(as.numeric(totsplandlb))) %>% # total by species agg over time
[…]

### 2021-02-26 docs/GBFleetDefinitions.Rmd
GBFleetDefinitions.rmd describing clustering - update index.rmd to include this - update references.bib file - added figures folder + dendrogram png

---
title: "Fleet Definitions"
output:
  bookdown::html_document2:
    code_fold: hide
    number_sections: false
csl: apa-annotated-bibliography.csl
bibliography: references.bib
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{r getGears,  echo=F, eval=T}
gears <- readRDS(here::here("data","gearCodeTable.rds"))
gearCodes <- gears$data %>% 
  dplyr::arrange(NEGEAR,NEGEAR2) %>%
  dplyr::select(NEGEAR,NEGEAR2,GEARNM,NAFOGEAR,NAFOGRNM)
numGears <- length(unique(gearCodes$NEGEAR2))

gearData <- readRDS(file=here::here("data","gearLandingsBySpecies.rds"))
gearTypes <- unique(gearData$data$NEGEAR2)
numGearsGB <- length(gearTypes)
```
There are a total of `r numGears` unique NEGEAR2 codes (found in the cfdbs.gear table). However on Georges Bank not all of these fishing gears are used. A subset of `r numGearsGB` gear types are used. A list of these codes follow

```{r printGearTable, echo=F,eval=T}

GBCodes <- gearCodes %>% dplyr::filter(NEGEAR2 %in% sprintf("%02d",gearTypes)) %>% 
  dplyr::select(-NEGEAR,-NAFOGEAR,-NAFOGRNM)

DT::datatable(GBCodes,rownames = FALSE, 
          options = list(pageLength = 10, 
                         order = list(list(0, 'asc'))
          ))

```

## Aggregating Gear Types

Each gear type is designed to target particular species. Some gear types may catch predominantly one species (eg. Lobsters traps) where others may catch multiple species (bottom trawls). We aim to aggregate the gear types into fleets using similarities in the catch among the gear types. To achieve this we will adopt an agglomerative hierarchical, nearest neighbor clustering method (@manly2017) using euclidean distance as the measure of distance between groups. In this approach all gear types begin in separate groups. Groups that are similar (similar catch, defined by a distance mesaure) are then merged until at some distance all gear types belong to a single group. The variables used to determine the grouping are the species caught, measured by landings (in lbs)

Landings data, by gear type, are pulled from the database. 
```{r pulldata, echo=T,eval=F}
channel <- dbutils::connect_to_database("server","id")
get_landings_by_data(channel,percentLanded=0.90,area=c(cfdbs::EPUs$data[["GB"]],537),outDir=here::here("data"))
```

The landings of each species are aggregated over time and then ordered from the most abundant to least abundant. To reduce the number of variables (species) in the analysis we only use species which comprise 90% of the total landings. The resulting data are then formatted as a data frame (or matrix) where each gear type is represented in a row and each species in a column. Each cell being the landings of the species by the gear type. Note: There will be many zeros in the data frame representing gear types that do not catch a particular species.

The variables are often standardized so that all of the variables are equally important in determining distances. We standardize each variable to have zero mean and unit variance. At this point pairwise distances between all gear types are calculated using the Euclidean distance function

$$d_{i,j} = \left( \sum_{k=1}^p(x_{i,k}-x_{j,k})^2 \right) ^{1/2} $$
where $x_{i,k}$ is the landings of species $k$ by gear type $i$ and $x_{j,k}$ is the landings of species $k$ by gear type $j$. $d_{i,j}$ is the measure of distance between gear type $i$ and $j$

A dendrogram (Figure \@ref(fig:dendro1)) is then formed using the agglomerative hierarchical process. This process is carried out using the statistical package `cluster`(@cluster211) in R.

```{r cluster1, echo=T,eval=F}
clusterObj <- cluster_analysis_all_gears()
plot(clusterObj,ask=T,which.plots=2,main="Complete gear list using 90% landing from each gear",xlab="")
#' 
```
```{r dendro1,out.width="100%",fig.cap="The denrogram obtained from a nearest-neighbor, hierarchical cluster analysi on all gears used in the Georges Bank Ecological Production unit defined by Statistical areas"}
knitr::include_graphics("figures/dendrogram1.png")
```

## References

### 2021-02-26 docs/GBFleetDefinitions.Rmd
updated index.html

@@
 
-A dendrogram (Figure \@ref(fig:dendro1)) is then formed using the agglomerative hierarchical process. This process is carried out using the statistical package `cluster`(@cluster211) in R.
+A dendrogram (Figure \@ref(fig:dendro1)) is then formed using the agglomerative hierarchical process.
+This process is carried out using the statistical package `cluster`(@cluster211) in R.
+Initial inspection indicates further "cleaning" of the data is required.
 
@@
 plot(clusterObj,ask=T,which.plots=2,main="Complete gear list using 90% landing from each gear",xlab="")
-#' 
+
 ```

### 2021-02-26 docs/GBFleetDefinitions.Rmd
added author, date to GBfleet rmd

@@
 title: "Fleet Definitions"
+author: Andy Beet
+date: "`r format(Sys.time(), '%d %B, %Y')`"
 output:
@@
 ```
-```{r dendro1,out.width="100%",fig.cap="The denrogram obtained from a nearest-neighbor, hierarchical cluster analysi on all gears used in the Georges Bank Ecological Production unit defined by Statistical areas"}
+```{r dendro1,out.width="100%",fig.cap="The denrogram obtained from a nearest-neighbor, hierarchical cluster analysis on all gears used in the Georges Bank Ecological Production unit defined by statistical areas"}
 knitr::include_graphics("figures/dendrogram1.png")

### 2021-03-01 docs/GBFleetDefinitions.Rmd
bug fix: labels not matching gear types in dendrogram plot

@@
 ```
+Note: Low values of the agglomerative coefficient indicate tight clustering of gears types. Larger values indicate less well formed clusters.
 

### 2021-03-01 R/fleetClusterAnalysis.R

@@ -44,7 +44,7 @@ cluster_analysis_all_gears <- function(){
       gearCodesUpdate$GEARName[icode] <-substr(newCode,1,10) # shortens name to 10 characters for ease in plotting
     } else {
       gearCodesUpdate$NEGEAR2[icode] <- as.numeric(uniqueCodes[icode])
-      gearCodesUpdate$GEARName[icode] <- gearCodes$GEARNM2[icode]
+      gearCodesUpdate$GEARName[icode] <- gearCodes$GEARNM2[ind]
   
     }
   }

### 2021-03-01 R/cluster_analysis_all_gears.R
add clustering functions - cluster_analysis_all_gears.r - cluster_analysis_filter_gears.r - explore_landings_by_gear.r

@@ -1,0 +1,89 @@
+#' Perform cluster analysis
+#' 
+#' Uses all gear Types and species contributing to x% of landings for each gear type.
+#' Reads in preprocessed data from get_landings_by_gear.r
+#' 
+#' @return object of class \code{agnes} representing the clustering. Use \code{plot} top display dendrogram
+#' 
+#' @example 
+#' \dontrun{
+#' clusterObj <- cluster_analysis_all_gears()
+#' # dendrogram plot
+#' plot(clusterObj,ask=T,which.plots=2,main="Complete gear list using 90% landing from each gear",xlab="")
+#' }
+#' 
+#' 
+
+library(magrittr)
+
+
+
+cluster_analysis_all_gears <- function(){
+
+  # read in landings data by gear/species aggregated over time
+  allGearData <- readRDS(here::here("data","gearLandingsBySpecies.rds"))
+  # read in all gear codes
+  gearCodes <- readRDS(here::here("data","gearCodeTable.rds"))
+  
+  # split long names by comma and select first name
+  gearCodes <- gearCodes$data %>% 
+    dplyr::select(NEGEAR2,GEARNM) %>%
+    dplyr::mutate(GEARNM2=unlist(lapply(stringr::str_split(gearCodes$data$GEARNM,","),`[[`,1))) %>% 
+    dplyr::select(-GEARNM) %>% 
+    dplyr::distinct(NEGEAR2,GEARNM2)
+  
+  # find unique gear codes, make single string if has multiple entries and create new table
+  uniqueCodes <- unique(gearCodes$NEGEAR2)
+  gearCodesUpdate <- data.frame(matrix(nrow = length(uniqueCodes),ncol=2))
+  names(gearCodesUpdate) <- c("NEGEAR2","GEARName")
+  for (icode in 1:length(uniqueCodes)) {
+    ind <- gearCodes$NEGEAR2 == uniqueCodes[icode]
+    if(sum(ind) > 1) {
+      newCode <- paste0(gearCodes$GEARNM2[ind],"/",collapse=" ")
+      gearCodesUpdate$NEGEAR2[icode] <- as.numeric(uniqueCodes[icode])
+      gearCodesUpdate$GEARName[icode] <-substr(newCode,1,10) # shortens name to 10 characters for ease in plotting
+    } else {
+      gearCodesUpdate$NEGEAR2[icode] <- as.numeric(uniqueCodes[icode])
[…]

### 2021-03-01 R/cluster_analysis_filter_gears.R

@@ -1,13 +1,14 @@
 #' Perform cluster analysis
 #' 
-#' Uses all gear Types and species contributing to x% of landings for each gear type.
-#' Reads in preprocessed data from get_landings_by_gear.r
+#' Uses a filtered set of gear Types. 
+#' gear types are chosen based on how many lbs in total is landed aggregated over time. 
+#' landings are ordered by gear type an the gears contributing to the top x% of landings are retained
 #' 
 #' @return object of class \code{agnes} representing the clustering. Use \code{plot} top display dendrogram
 #' 
 #' @example 
 #' \dontrun{
-#' clusterObj <- cluster_analysis_all_gears()
+#' clusterObj <- cluster_analysis_filtered_gears()
 #' # dendrogram plot
 #' plot(clusterObj,ask=T,which.plots=2,main="Complete gear list using 90% landing from each gear",xlab="")
 #' }
@@ -16,15 +17,13 @@
 
 library(magrittr)
 
-
-
-cluster_analysis_all_gears <- function(){
+cluster_analysis_filter_gears <- function(filterByLandings=.99){
 
   # read in landings data by gear/species aggregated over time
-  allGearData <- readRDS(here::here("data","gearLandingsBySpecies.rds"))
+  #allGearData <- readRDS(here::here("data","gearLandingsBySpecies.rds"))
+  allTimeGearData <- readRDS(here::here("data","timeSeriesSpeciesByGear.rds"))
   # read in all gear codes
   gearCodes <- readRDS(here::here("data","gearCodeTable.rds"))
-  
   # split long names by comma and select first name
   gearCodes <- gearCodes$data %>% 
     dplyr::select(NEGEAR2,GEARNM) %>%
@@ -49,12 +48,40 @@ cluster_analysis_all_gears <- function(){
     }
   }
   
+  # aggregate data over time
+  allTimeGearDataAgg <- allTimeGearData$data %>%
+    dplyr::select(-n) %>% 
+    dplyr::group_by(NESPP3,NEGEAR2) %>%
+    dplyr::summarise(totsplandlb = sum(totsplandlb))
+    
+
   # now joint unique code names to main gear data table
-  gearTable <- allGearData$data %>% dplyr::left_join(gearCodesUpdate,by="NEGEAR2") %>%
[…]

### 2021-03-01 R/explore_landings_by_gear.R

@@ -1,0 +1,63 @@
+#' Explore landings 
+#' 
+#' Explores time series of species caught by gear type
+#' 
+#' @param data Data frame. landings data by YEAR, NESPP3, NEGEAR2 (long format)
+#' @param gearCode Numeric. NEGEAR2 code
+#' @param filterYrs Numeric. A proportion reflecting the proportion of years containing data for a particucar species (eg.)
+#' A value of .15 retains only the species in which there were landings in more that 15% of the years sampled
+#' 
+#' @return 
+#' 
+#' @example 
+#' \dontrun{
+#' }
+#' 
+#' 
+
+library(magrittr)
+library(ggplot2)
+gearData <- readRDS(file=paste0(here::here("data","timeSeriesSpeciesByGear.rds")))
+
+explore_landings_by_gear <- function(data=gearData$data,gearCode=5,filterYrs=0.30){
+  
+  # pull gear Specific data
+  gd <- data %>%
+    dplyr::filter(NEGEAR2==gearCode) %>%
+    dplyr::mutate(NESPP3 = as.numeric(NESPP3),YEAR=as.numeric(YEAR))
+  
+  # yr range
+  yrRange <- range(as.numeric(gd$YEAR))
+
+  # find species that are not abundant
+  filteredSpecies <- gd %>% 
+    dplyr::group_by(NESPP3) %>%
+    dplyr::summarise(n2=dplyr::n()) %>% 
+    dplyr::filter(n2 > filterYrs*diff(yrRange))
+  
+  speciesLeftOut <- sort(setdiff(unique(gd$NESPP3),filteredSpecies$NESPP3))
+  print(speciesLeftOut)
+  
+  # filter out species from data set
+  gd2 <- gd %>% 
+    dplyr::filter(NESPP3 %in% filteredSpecies$NESPP3)
+  
+  # calculate % of landings lost due to filtering out non abundant species
+  landingsLeftOut <- gd %>% dplyr::filter(NESPP3 %in% speciesLeftOut) %>%
+    dplyr::select(totsplandlb) %>% 
+    sum()
+  totalLandings <- gd %>% 
+    dplyr::select(totsplandlb) %>% 
+    sum()
+  
+  message(paste0("Percentage of landings due to omitted species = ",100*landingsLeftOut/totalLandings))
+  
+  p <- ggplot(data=gd2) +
[…]

### 2021-03-01 docs/GBFleetDefinitions.Rmd
update rmd with clustering plots

--- 2021-02-26 docs/GBFleetDefinitions.Rmd
@@
 
-gearData <- readRDS(file=here::here("data","gearLandingsBySpecies.rds"))
-gearTypes <- unique(gearData$data$NEGEAR2)
-numGearsGB <- length(gearTypes)
 ```
@@
 However on Georges Bank not all of these fishing gears are used.
-A subset of `r numGearsGB` gear types are used.
-A list of these codes follow
+A subset of 45 gear types are used.
+All gear codes:
 
@@
 
-GBCodes <- gearCodes %>% dplyr::filter(NEGEAR2 %in% sprintf("%02d",gearTypes)) %>% 
+GBCodes <- gearCodes %>% 
   dplyr::select(-NEGEAR,-NAFOGEAR,-NAFOGRNM)
@@
 ```
-```{r dendro1,out.width="100%",fig.cap="The denrogram obtained from a nearest-neighbor, hierarchical cluster analysis on all gears used in the Georges Bank Ecological Production unit defined by statistical areas"}
+```{r dendro1,out.width="100%",fig.cap="The denrogram obtained from a nearest-neighbor, hierarchical cluster analysis on all gear types used in the Georges Bank Ecological Production unit as defined by statistical areas",eval=T,echo=F}
 knitr::include_graphics("figures/dendrogram1.png")
+```
+Note: Low values of the agglomerative coefficient indicate tight clustering of gears types. Larger values indicate less well formed clusters.
+
+### Alternative approach
+
+Details to follow:
+
+```{r dendro2,out.width="100%",fig.cap="The denrogram obtained from a nearest-neighbor, hierarchical cluster analysis on a subset of gear types in the Georges Bank Ecological Production unit as defined by statistical areas.
+Gear types were selected based on their contribution to 99% of landings",eval=T,echo=F}
+knitr::include_graphics("figures/dendrogram2.png")
 ```

### 2021-03-02 docs/GBFleetDefinitions.Rmd
add figures of landings by gear type to GBFleetDefinitions.rmd

@@
 ---
-title: "Fleet Definitions"
+title: "Fleet Definitions for Georges Bank"
 author: Andy Beet
@@
 library(magrittr)
+options(warn=-1)
 ```
+
+## Spatial Extent
+
+Statistical areas are used to define the Georges Bank footprint for pulling the landings data by gear
+```{r}
+crs <- 4326
+# read in coastline and GB shape file
+coast <- sf::st_read(here::here("gis", 'NES_LME_coast.shp'), quiet = T) %>%
+  sf::st_transform(.,crs=crs)
+GB_strata <- sf::st_read(here::here("gis","Statistical_Areas_2010.shp"),quiet=T) %>%
+    sf::st_transform(.,crs=crs) %>%
+  dplyr::filter(Id %in% c(521, 522, 523, 524, 525, 526, 538, 551, 552, 561, 562, 537) )
+centroids <- sf::st_coordinates(sf::st_centroid(GB_strata))
+GB_strata <- cbind(GB_strata,centroids)
+
+#plot map
+ggplot2::ggplot(data=coast) +
+  ggplot2::geom_sf() +
+  ggplot2::geom_sf(data=GB_strata,fill="grey") + 
+  ggplot2::geom_text(data=GB_strata,ggplot2::aes(x=X,y=Y,label=Id),size=2)    +
+  ggplot2::coord_sf(xlim = c(-76,-65), ylim = c(38,44))
+```
+
+## Gear Types
 
@@
 ```
-```{r dendro1,out.width="100%",fig.cap="The denrogram obtained from a nearest-neighbor, hierarchical cluster analysis on all gear types used in the Georges Bank Ecological Production unit as defined by statistical areas",eval=T,echo=F}
+```{r dendro1,out.width="100%",fig.cap="The dendrogram obtained from a nearest-neighbor, hierarchical cluster analysis on all gear types used in the Georges Bank Ecological Production unit as defined by statistical areas",eval=T,echo=F}
 knitr::include_graphics("figures/dendrogram1.png")
@@
 
-```{r dendro2,out.width="100%",fig.cap="The denrogram obtained from a nearest-neighbor, hierarchical cluster analysis on a subset of gear types in the Georges Bank Ecological Production unit as defined by statistical areas.
+```{r dendro2,out.width="100%",fig.cap="The dendrogram obtained from a nearest-neighbor, hierarchical cluster analysis on a subset of gear types in the Georges Bank Ecological Production unit as defined by statistical areas.
 Gear types were selected based on their contribution to 99% of landings",eval=T,echo=F}
 knitr::include_graphics("figures/dendrogram2.png")
+
+```
+
+### Species landings by gear {.tabset .tabset-pills}
+
+Time series of species caught by each gear type
+
+#### 01
+
+```{r gear01,  echo=F, eval =T}
+GBCodes %>% dplyr::filter(NEGEAR2=="01")
+knitr::include_graphics("figures/gear01.png")
+```
+
+#### 02
+```{r gear02,  echo=F, eval =T}
+GBCodes %>% dplyr::filter(NEGEAR2=="02")
+knitr::include_graphics("figures/gear02.png")
+```
+
+#### 05
+
+```{r gear05,  echo=F, eval =T}
+GBCodes %>% dplyr::filter(NEGEAR2=="05")
+knitr::include_graphics("figures/gear05.png")
+```
+
+#### 10
+
+```{r gear10,  echo=F, eval =T}
+GBCodes %>% dplyr::filter(NEGEAR2=="10")
+knitr::include_graphics("figures/gear10.png")
+```
+
+#### 12
+
+```{r gear12,  echo=F, eval =T}
+GBCodes %>% dplyr::filter(NEGEAR2=="12")
+knitr::include_graphics("figures/gear12.png")
+```
+
+#### 13
+
+```{r gear13,  echo=F, eval =T}
+GBCodes %>% dplyr::filter(NEGEAR2=="13")
+knitr::include_graphics("figures/gear13.png")
+```
+
+#### 14
+
+```{r gear14,  echo=F, eval =T}
+GBCodes %>% dplyr::filter(NEGEAR2=="14")
+knitr::include_graphics("figures/gear14.png")
+```
+
+#### 17
+```{r gear17,  echo=F, eval =T}
+GBCodes %>% dplyr::filter(NEGEAR2=="17")
+knitr::include_graphics("figures/gear17.png")
+```
+
+#### 18
+```{r gear18,  echo=F, eval =T}
+GBCodes %>% dplyr::filter(NEGEAR2=="18")
+knitr::include_graphics("figures/gear18.png")
+```
+
+#### 20
+```{r gear20,  echo=F, eval =T}
+GBCodes %>% dplyr::filter(NEGEAR2=="20")
+knitr::include_graphics("figures/gear20.png")
+```
+
+#### 21
+```{r gear21,  echo=F, eval =T}
+GBCodes %>% dplyr::filter(NEGEAR2=="21")
+knitr::include_graphics("figures/gear21.png")
+```
+
+#### 30
+```{r gear30,  echo=F, eval =T}
+GBCodes %>% dplyr::filter(NEGEAR2=="30")
+knitr::include_graphics("figures/gear30.png")
+```
+
+#### 37
+```{r gear37,  echo=F, eval =T}
+GBCodes %>% dplyr::filter(NEGEAR2=="37")
+knitr::include_graphics("figures/gear37.png")
+```
+
+#### 38
+```{r gear38,  echo=F, eval =T}
+GBCodes %>% dplyr::filter(NEGEAR2=="38")
+knitr::include_graphics("figures/gear38.png")
+```
+
+#### 40
+```{r gear40,  echo=F, eval =T}
+GBCodes %>% dplyr::filter(NEGEAR2=="40")
+knitr::include_graphics("figures/gear40.png")
 ```

### 2021-03-04 R/get_landings_by_gear.R
updated dendrogram. previously had missing gear types due to bug in filtering

@@ -37,13 +37,20 @@ get_landings_by_gear <- function(channel,percentLanded=0.90,area=c(cfdbs::EPUs$d
       dplyr::ungroup()
     
     # select species that account for x% of landings
-    orderSpecies <- aggGear %>% 
+    orderedSpecies <- aggGear %>% 
       dplyr::group_by(NESPP3) %>% 
       dplyr::summarise(sumOverYears=sum(as.numeric(totsplandlb))) %>% # total by species agg over time
       dplyr::arrange(desc(sumOverYears)) %>% 
       dplyr::mutate(csum=cumsum(sumOverYears)) %>%
-      dplyr::mutate(prop = csum/sum(sumOverYears)) %>% 
+      dplyr::mutate(prop = csum/sum(sumOverYears))
+    
+    orderSpecies <- orderedSpecies %>% 
       dplyr::filter(prop <= percentLanded)
+    
+    if (dim(orderSpecies)[1]==0){
+      # dominated by one species > percentLanded
+      orderSpecies <- orderedSpecies[1,]
+    }
   
     # select these species from data pull to account for gear species interaction
     useData <- aggGear %>% 

### 2021-03-04 R/cluster_analysis_filter_gears.R
moved cluster analysis (cluster_gears) into own function and called from data prep function

@@ -4,18 +4,21 @@
 #' gear types are chosen based on how many lbs in total is landed aggregated over time. 
 #' landings are ordered by gear type an the gears contributing to the top x% of landings are retained
 #' 
+#' @param filterByLandings Numeric scalar. Proportion of landing to be captured by gears. (Default=.99)
+#' 
 #' @return object of class \code{agnes} representing the clustering. Use \code{plot} top display dendrogram
 #' 
 #' @example 
 #' \dontrun{
 #' clusterObj <- cluster_analysis_filtered_gears()
 #' # dendrogram plot
-#' plot(clusterObj,ask=T,which.plots=2,main="Complete gear list using 90% landing from each gear",xlab="")
+#' plot(clusterObj,ask=T,which.plots=2,main="Complete gear list using 99% landing from each gear",xlab="")
 #' }
 #' 
 #' 
 
 library(magrittr)
+source(here::here("R","cluster_gears.r"))
 
 cluster_analysis_filter_gears <- function(filterByLandings=.99){
 
@@ -59,9 +62,7 @@ cluster_analysis_filter_gears <- function(filterByLandings=.99){
   gearTable <- allTimeGearDataAgg %>% dplyr::left_join(gearCodesUpdate,by="NEGEAR2") %>%
     dplyr::mutate(GEARID=paste0(NEGEAR2,"-",GEARName)) %>%
     dplyr::select(-GEARName)
-  
-  print(gearTable)
-  
+ 
   # Prep data for analysis --------------------------------------------------
   # aggregate landings over time for each gear type
   gd <- allTimeGearData$data %>%
@@ -74,40 +75,33 @@ cluster_analysis_filter_gears <- function(filterByLandings=.99){
   # pick top species that meet minimum landings percentage criterion 
   ind <- c(T,gd$proportion <= filterByLandings)
   ind <- head(ind,-1)
-  gd <- gd[ind,]
+  topData <- gd[ind,]
 
-  print(gd)
+  topTable <- gearTable %>% 
+    dplyr::filter(NEGEAR2 %in% topData$NEGEAR2)
   
-  gearTable <- gearTable %>% 
-    dplyr::filter(NEGEAR2 %in% gd$NEGEAR2) %>%
+  topGears <- unique(topTable$NEGEAR2)
+  
+  topTable <- topTable %>%
     dplyr::select(-NEGEAR2)
 
+  ## cluster analysis on top x%
[…]

### 2021-03-04 R/explore_landings_by_gear.R
add code to create time series plots for gear types

@@ -2,10 +2,12 @@
 #' 
 #' Explores time series of species caught by gear type
 #' 
+#' @param channel 
 #' @param data Data frame. landings data by YEAR, NESPP3, NEGEAR2 (long format)
 #' @param gearCode Numeric. NEGEAR2 code
 #' @param filterYrs Numeric. A proportion reflecting the proportion of years containing data for a particucar species (eg.)
 #' A value of .15 retains only the species in which there were landings in more that 15% of the years sampled
+#' @param saveToFile Boolean. Save to doc/figures folder
 #' 
 #' @return 
 #' 
@@ -17,9 +19,10 @@
 
 library(magrittr)
 library(ggplot2)
+
 gearData <- readRDS(file=paste0(here::here("data","timeSeriesSpeciesByGear.rds")))
 
-explore_landings_by_gear <- function(data=gearData$data,gearCode=5,filterYrs=0.30){
+explore_landings_by_gear <- function(channel,data=gearData$data,gearCode=5,filterYrs=0.30,saveToFile=F){
   
   # pull gear Specific data
   gd <- data %>%
@@ -36,8 +39,7 @@ explore_landings_by_gear <- function(data=gearData$data,gearCode=5,filterYrs=0.3
     dplyr::filter(n2 > filterYrs*diff(yrRange))
   
   speciesLeftOut <- sort(setdiff(unique(gd$NESPP3),filteredSpecies$NESPP3))
-  print(speciesLeftOut)
-  
+
   # filter out species from data set
   gd2 <- gd %>% 
     dplyr::filter(NESPP3 %in% filteredSpecies$NESPP3)
@@ -51,13 +53,55 @@ explore_landings_by_gear <- function(data=gearData$data,gearCode=5,filterYrs=0.3
     sum()
   
   message(paste0("Percentage of landings due to omitted species = ",100*landingsLeftOut/totalLandings))
+
+  # sort in order of abundance
+  # aggregate landings over time for each gear type
+  gd3 <- data %>%
+    dplyr::filter(NEGEAR2 == gearCode) %>%
+    dplyr::group_by(NESPP3) %>%
+    dplyr::summarise(totLandingslb=sum(totsplandlb)) %>%
+    dplyr::arrange(desc(totLandingslb)) %>%
+    dplyr::mutate(cumusum = cumsum(totLandingslb)) %>%
[…]

### 2021-03-04 docs/GBFleetDefinitions.Rmd
add plot for gear types = 1% of landings (least caught)

@@
 
+### Species that make up < 1% landings
+
+```{r dendro3,out.width="100%",fig.cap="The dendrogram obtained from a nearest-neighbor, hierarchical cluster analysis on a subset of gear types in the Georges Bank Ecological Production unit as defined by statistical areas.
+Gear types were selected based on not contributing to top 99% of landings",eval=T,echo=F}
+knitr::include_graphics("figures/dendrogram3.png")
+
+```
+
 ## References

### 2021-03-12 docs/SimData.Rmd
weight diets part 1

@@
 
+source(here("simulated-data/config/NOBA2Config.R"))
+
 all_diets <- read_savedsurvs(d.name, 'survDiet')
@@
 
+NOBAom_ms <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
+
 trueB_age_allbox <- NOBAom_ms$truebio_ss %>%
   #sum over polygons
-  aggregate(dat$atoutput,list(dat$species,dat$agecl,dat$time, dat$prey),sum)
+  #aggregate(dat$atoutput,list(dat$species,dat$agecl,dat$time, dat$prey),sum)
+  group_by(species,agecl,time) %>%
+  summarise(sum(atoutput))
 

### 2021-03-12 docs/allocateLandingsEPU.Rmd

@@
 
-```{r calc_props_sp, eval=T, out.width="100%",fig.align='center',fig.cap="Proportion of landings in GB relative to statistical area"}
+```{r calc_props_st, eval=T, out.width="100%",fig.align='center',fig.height=10,,fig.cap="Proportion of landings in GB relative to statistical area"}
 source(here::here("data-raw","plot_landings_proportions.r"))
-plot_landings_proportions(species="cols")
-
-```
-
-The same data is represented in the plot below but the species are plotted as rows instead of columns
-
-```{r calc_props_st, eval=T, out.width="100%",fig.align='center',fig.cap="Proportion of landings in GB relative to statistical area"}
 plot_landings_proportions(species="rows")
@@
 
-```{r calc_lands,  eval=T, out.width="100%",fig.align='center',fig.cap="Total landings by statistical area"}
+```{r calc_lands,  eval=T, out.width="100%",fig.align='center',fig.height=10,fig.cap="Total landings by statistical area"}
 source(here::here("data-raw","plot_landings.r"))

### 2021-03-12 docs/SimData.Rmd
generate diet data for MSSPM

@@
 
-```{r, write-msspm-diet}
+Diet composition is by predator age class for each time (unit=time.days) and is a proportion (sums to 1 for each time).
+We get species diet composition by aggregating over age classes (weight diet composition at age by true biomass at age).
+
+```{r, write-msspm-diet, eval=FALSE}
 
@@
 
-# aggregate over ages (survey diet is in proportion by age)
+# divide time.days by runpar$toutinc to match time in biomass
+# aggregate over ages using true biomass at age (survey diet is in proportion by age)
 
@@
   group_by(species,agecl,time) %>%
-  summarise(sum(atoutput))
+  summarise(trueBatage = sum(atoutput))
+
+trueB <- trueB_age_allbox %>%
+  group_by(species, time) %>%
+  summarise(trueB = sum(trueBatage))
+
+trueBagecl <- merge(trueB_age_allbox, trueB)
+
+#multiple surveys named in list object, use as filename
+for(s in names(all_diets)){
+  #arrange into wide format: year, Species1, Species2 ... and write csv
+  svdiet <- all_diets[[s]][[1]] %>%
+    #calc_timestep2time(d.name, ., run.prm.file) %>% #this gives an arbitrary start year of 1970
+    dplyr::mutate(time = time.days/NOBAom_ms$runpar$toutinc) %>%
+    dplyr::left_join(., trueBagecl) %>%
+    dplyr::mutate(dietBwt = dietSamp*trueBatage) %>%
+    dplyr::group_by(species,time,prey,trueB) %>%
+    dplyr::summarise(totdietBwt = sum(dietBwt)) %>%
+    dplyr::mutate(totdietComp = totdietBwt/trueB) %>%
+    dplyr::mutate(year = ceiling(time/stepperyr)) %>%
+    dplyr::ungroup() %>%
+    dplyr::select(year, species, prey, totdietComp) %>%
+    write_csv(paste0(o.name,"diet_",s,".csv"))
+}
 

### 2021-03-12 docs/SimData.Rmd
start hydra csv files

@@
 
-```{r, write-msspm-diet, eval=FALSE}
+```{r write-msspm-diet, eval=FALSE}
 
@@
 
+For the length-based model [hydra](https://github.com/thefaylab/hydra_sim), we can write .csv files to be used as input in the [hydradata](https://github.com/thefaylab/hydradata) package, which writes .dat and .pin files.
+
+Hydra needs survey biomass time series, catch time series, survey length composition, fishery length composition, and diet composition data.
+
+Observations for biomass and catch are very similar to those created for MMSPM. Hydra inputs will need to be in the same species order for each dataset, so first we make a species file.
+
+```{r write-hydra}
+
+diettrue <- load_diet_comp(here("simulated-data/atlantisoutput/NOBA_march_2020"), 
+                           "nordic_runresults_01DietCheck.txt", fgs = fgs)
+
+ms_diet <- diettrue %>%
+  filter(species %in% sppsubset$Name) %>%
+  filter(atoutput>0)
+
+predOrPrey <- ms_diet %>%
+  filter(prey %in% unique(species)) %>%
+  select(species, agecl, time.days, prey, atoutput) %>%
+  group_by(species, prey) %>%
+  summarise(avgpreyprop = mean(atoutput))
+
+specieslist <- data.frame(species = unique(unique(survObsBiom[[2]][[1]]$species)),
+                          guild = rep("na"),
+                          guildmember = c(1:11),
+                          predOrPrey = rep(1)
+)
+
+```
+
 ## Generating a survey dataset

### 2021-03-15 docs/SimData.Rmd
foodweb and speciesList files for hydra

@@
   filter(species %in% sppsubset$Name) %>%
-  filter(atoutput>0)
+  filter(atoutput>0) %>%
+  mutate(species = as.character(species),
+         prey = as.character(prey))
 
-predOrPrey <- ms_diet %>%
+predPrey <- ms_diet %>%
   filter(prey %in% unique(species)) %>%
@@
 
+foodweb <- left_join(sppsubset, predPrey, by=c("Name" = "species")) %>%
+  select(species, prey, avgpreyprop) %>%
+  mutate(avgpreyprop = ceiling(avgpreyprop)) %>%
+  spread(species, avgpreyprop) %>%
+  filter(!is.na(prey)) %>%
+  replace(is.na(.), 0)
+
+predOrPrey <- ifelse(colSums(foodweb[-1])>0, 1, 0)
+
 specieslist <- data.frame(species = unique(unique(survObsBiom[[2]][[1]]$species)),
@@
                           guildmember = c(1:11),
-                          predOrPrey = rep(1)
+                          predOrPrey = predOrPrey
 )
+
+o.name <- here("simulated-data/msinput/hydra/")
+write_csv(foodweb, paste0(o.name,"foodweb_NOBA.csv"))
+write_csv(specieslist, paste0(o.name,"speciesList_NOBA.csv"))
 

### 2021-03-15 docs/SimData.Rmd
add survey/catch dat csvs for hydra

@@
 
-```{r write-hydra}
+```{r write-hydra, eval=FALSE}
 
@@
 
+Survey and catch datasets for hydra from NOBA (same methods as MSSPM). Make this a function that takes output directory name and output names as arguments. 
+```{r, eval=FALSE}
+o.name <- here("simulated-data/msinput/hydra/")
+
+stepperyr <- if(NOBAom_ms$runpar$outputstepunit=="days") 365/NOBAom_ms$runpar$toutinc
+
+#read in survey biomass data
+survObsBiom <- read_savedsurvs(d.name, 'survB') #reads in all surveys
+
+#multiple surveys named in list object, use as filename
+for(s in names(survObsBiom)){
+  #arrange into wide format: year, Species1, Species2 ... and write csv
+  svbio <- survObsBiom[[s]][[1]] %>%
+    #calc_timestep2time(d.name, ., run.prm.file) %>% #this gives an arbitrary start year of 1970
+    dplyr::mutate(year = ceiling(time/stepperyr)) %>%
+    dplyr::select(species, year, atoutput) %>%
+    tidyr::spread(species, atoutput) %>%
+    write_csv(paste0(o.name,"observation_biomass_NOBA_",s,".csv"))
+}
+
+#read in catch data
+read_savedfisheries <- function(dir, type){
+
+  datlook <- data.frame(dattype = c('Catch', 'catchAge', 'catchLen', 'catchWtage', 'catchAnnAge', 'catchAnnWtage'),
+                        pattern = c("*fishCatch.rds", "*fishObsAgeComp.rds", "*fishObsLenComp.rds", "*fishObsWtAtAge.rds",
+                                    "*fishObsFullAgeComp.rds", "*fishObsFullWtAtAge.rds"))
+
+  fisheries <- list.files(path=dir, pattern = as.character(datlook$pattern[datlook$dattype %in% type]), full.names = TRUE)
+
+  fishery.name <-  str_match(fisheries, paste0(scenario.name,"_\\s*(.*?)\\s",datlook$pattern[datlook$dattype==type]))[,2]
+
+  dat <- lapply(fisheries, readRDS)
+
+  names(dat) <- fishery.name
+
+  return(dat)
+}
+
+catchbio_ss <- read_savedfisheries(d.name, 'Catch')
+
+for(c in names(catchbio_ss)){
+  #arrange into wide format: year, Species1, Species2 ... and write csv
+  catchbio <- catchbio_ss[[c]][[1]] %>%
+    dplyr::mutate(year = time/365) %>%
+    dplyr::select(species, year, atoutput) %>%
+    tidyr::spread(species, atoutput) %>%
+    write_csv(paste0(o.name,"observation_catch_NOBA_",c,".csv"))
+}
+
+#write readme for metadata
+meta <- c("Simulated data from Norweigan-Barents Atlantis Model by Hansen et al. \n\n",
+          paste("Atlantis outputs stored in", d.name, '\n\n'),
+          paste("Biomass output for surveys", names(survObsBiom), '\n\n'),
+          paste("Fishery catch output for fisheries", names(catchbio_ss), '\n\n'),
+          "Biomass and catch output units are annual tons.\n\n",
+          "Survey specifications:\n",
+          "----------------------\n",
+          read_file(here("/simulated-data/config/mssurvey_fall.R")),
+          "\n",
+          "----------------------\n",
+          read_file(here("/simulated-data/config/mssurvey_spring.R")),
+          "\n",
+          "----------------------\n",
+          read_file(here("/simulated-data/config/mssurvey_fall_01.R")),
+          "\n",
+          "----------------------\n",
+          read_file(here("/simulated-data/config/mssurvey_spring_01.R")),
+          "\n\n",
+          "Fishery specifications:\n",
+          "----------------------\n",
+          read_file(here("/simulated-data/config/msfishery.R")),
+          "\n",
+          "----------------------\n",
+          read_file(here("/simulated-data/config/msfishery_01.R")),
+          "\n"
+          )
+
+cat(as.character(meta), file=paste0(here("simulated-data/msinput/hydra/"),"README.txt"))
+
+```
+
 ## Generating a survey dataset

### 2021-03-15 docs/GBFleetDefinitions.Rmd
cluster gear types - by focal species - by < 1% - add two figures - update rmd + html

@@
 In this approach all gear types begin in separate groups.
-Groups that are similar (similar catch, defined by a distance mesaure) are then merged until at some distance all gear types belong to a single group.
+Groups that are similar (similar catch, defined by a distance measure) are then merged until at some distance all gear types belong to a single group.
 The variables used to determine the grouping are the species caught, measured by landings (in lbs)
@@
 
-The landings of each species are aggregated over time and then ordered from the most abundant to least abundant.
-To reduce the number of variables (species) in the analysis we only use species which comprise 90% of the total landings.
 The resulting data are then formatted as a data frame (or matrix) where each gear type is represented in a row and each species in a column.
@@
 
-A dendrogram (Figure \@ref(fig:dendro1)) is then formed using the agglomerative hierarchical process.
-This process is carried out using the statistical package `cluster`(@cluster211) in R.
-Initial inspection indicates further "cleaning" of the data is required.
+A dendrogram is then formed using the agglomerative hierarchical process. This process is carried out using the statistical package [`cluster`]() (@cluster211) in R.
+
+### Approach: Use all gear types
+
+For each gear type the landings of each species are aggregated over time and then ordered from the most abundant to least abundant.
+To reduce the number of variables (species) in the analysis we only use species which comprise 90% of the total landings for each gear type.
+
+(Figure \@ref(fig:dendro1)) shows the result of this approach. Initial inspection indicates further "cleaning" of the data is required.
 
@@
 ```
+
 ```{r dendro1,out.width="100%",fig.cap="The dendrogram obtained from a nearest-neighbor, hierarchical cluster analysis on all gear types used in the Georges Bank Ecological Production unit as defined by statistical areas",eval=T,echo=F}
@@
 
-### Alternative approach
+### Approach: Use gears responsible for most landings
 
-Details to follow:
+Total landings, aggregated over time, are ordered by gear type. The gear types that contribute to 99% of total landings are selected and used in the cluster analysis
+
+```{r cluster2, echo=T,eval=F}
+clusterObj <- cluster_analysis_filter_gears(filterByLandings=.99)
+plot(clusterObj$top,ask=T,which.plots=2,main="Gear types contributing to 99% of total landings",xlab="")
+
+```
 
@@
 
-Time series of species caught by each gear type
+Time series of species caught by each gear type.
+To reduce the number of panels in each figure, species caught < 30% of the time are not displayed.
+For example reporting of landings started in 1964 (56 years ago).
+So any species that have reported landings in fewer that 17 years are not displayed.
+
+```{r plotGearTop, echo=T,eval=F}
+for (gear in clusteObj$topGears) {
+  explore_landings_by_gear(channel,data=gearData$data,gearCode=gear,filterYrs=0.30,saveToFile=T)
+}
+
+```
 
@@
 
-### Species that make up < 1% landings
+### Gears that make up < 1% landings
+
+The remaining gears (not contributing to 99% of the landings) were used in a cluster analysis to determine how similar these gears are to each other, since 100% of landings should be accounted for.
+
+```{r cluster3, echo=T,eval=F}
+plot(clusterObj$bottom,ask=T,which.plots=2,main="Gear types contributing < 1% of total landings",xlab="")
+
+```
 
@@
 ```
+### Approach: Select gear types that catch species of interest
+
+All gear types which catch any amount of the [Species](https://noaa-edab.github.io/ms-keyrun/SpeciesEnvDat.html#focalSpecies) of interest are used in the cluster analysis.
+```{r echo=T,eval=F}
+focalSpecies <- readRDS(here::here("data","focalspecies.rds"))
+clusterObj <- cluster_by_focal_species(focalSpecies)
+plot(clusterObj,ask=T,which.plots=2,main="All gear types that land species of interest",xlab="")
+```
+
+```{r dendro4,out.width="100%",fig.cap="The dendrogram obtained from a nearest-neighbor, hierarchical cluster analysis on a subset of gear types in the Georges Bank Ecological Production unit as defined by statistical areas.
+All gear types that landed species of interest were selected",eval=T,echo=F}
+knitr::include_graphics("figures/dendrogram4.png")
+
+```
+
+## Bottom trawls (small/large mesh)
+
+Over time (dates/references????) changes in regulations resulted in the requirement of fisherman to increase the mesh size in their nets.
+This allowed smaller fish to escape the nets.
+However there were, and still are, exemptions to this regulation allowing fishermen to fish with smaller meshed nets depending on the time of the year and the area they fish.
+However all of the associated landings are flagged as using the same gear type code.
+Since the change in regulations dictate a change in the composition of species caught over time, this gear type needs to be investigated separately.
 

### 2021-03-15 docs/SimData.Rmd
lengthweight and fixed par files for hydra

@@
 
-Observations for biomass and catch are very similar to those created for MMSPM. Hydra inputs will need to be in the same species order for each dataset, so first we make a species file.
+Observations for biomass and catch are very similar to those created for MMSPM.
+Hydra inputs will need to be in the same species order for each dataset, so first we make a species file and some of the other input data that can come straight from Atlantis.
 
@@
 
+# foodweb and species file identify predators and prey; take from true atlantis diet comp
 diettrue <- load_diet_comp(here("simulated-data/atlantisoutput/NOBA_march_2020"), 
                            "nordic_runresults_01DietCheck.txt", fgs = fgs)
-
 ms_diet <- diettrue %>%
@@
 
+# take lenwt parameters from true atlantis pars
+lenwtpar <-  left_join(sppsubset, NOBAom_ms$biol$wl, by=c("Code" = "group")) %>%
+  select(Name, a, b) %>%
+  arrange(Name)
+
+# fixed parameters:
+# intake same for all teleosts
+intakespp  <- rbind(species = sppsubset$Name,
+                    alpha = rep(0.004),
+                    beta = rep(0.11))
+colnames(intakespp) <- intakespp[1,]
+intakespp <- as.data.frame(intakespp[-1,])
+
+# predator size preference ratio same for all species
+sizepref <- rbind(species = sppsubset$Name,
+                  mu = rep(0.5),
+                  sigma = rep(2))
+colnames(sizepref) <- sizepref[1,]
+sizepref <- as.data.frame(sizepref[-1,])
+
+# will need to fit von B and maturity curve functions to back out hydra input pars
+# maturity from maturity ogive input in biol file
+# growth functions from length at age in model...
+
+# write the csvs
 o.name <- here("simulated-data/msinput/hydra/")
@@
 write_csv(specieslist, paste0(o.name,"speciesList_NOBA.csv"))
+write_csv(lenwtpar, paste0(o.name,"lengthweight_species_NOBA.csv"))
+write.csv(intakespp, paste0(o.name,"intake_species_NOBA.csv"))
+write.csv(sizepref, paste0(o.name,"mortality_M2sizePreference_NOBA.csv"))
 

### 2021-03-16 docs/SimData.Rmd
add draft singletons, needs formatting

@@
 
+# dimensions for singletons file
+singletons <- data.frame(
+  debug=3,
+  Nyrs=NOBAom_ms$runpar$nyears, #may want to take from survey time series instead
+  Nspecies=length(sppsubset$Name),
+  Nsizebins=5,
+  Nareas=1,
+  Nfleets=5,
+  wtconv=1,
+  yr1Nphase=0,
+  recphase=0,
+  avg_rec_phase=0,
+  avg_F_phase=0,
+  dev_rec_phase=0,
+  dev_F_phase=0,
+  fqphase=0,
+  sqphase=0,
+  ssig_phase=0,
+  csig_phase=0,
+  phimax=1,
+  scaleInitialN=1,
+  otherFood=0000,
+  LFI_size=60,
+  bandwidth_metric=5,
+  assessmentPeriod=3,
+  flagMSE=0,
+  flagLinearRamp=1,
+  baseline_threshold=0.2
+)
+
+# M1 same for all species, depends on n length bins
+
 # will need to fit von B and maturity curve functions to back out hydra input pars
@@
 write.csv(sizepref, paste0(o.name,"mortality_M2sizePreference_NOBA.csv"))
+write.csv(t(singletons), paste0(o.name,"singletons_NOBA.csv"))
+
+```
+
+Need to fit growth and maturity parameters from Atlantis output for input into hydra.
+Possibly also recruitment?
+Maturity parameters are needed for SSB calculation and SSB-based recruitment.
+If we fit only to average recruitment with annual deviations, and don't need SSB, maturity and recruitment inputs could be dummy parameters for now.
+
+```{r fitpars}
+
+# need growth, vonB K and Linf and simple growth function
+
+# need 2-par maturity curve based on length, atlantis maturity ogive based on age group
 

### 2021-03-17 docs/SimData.Rmd
more hydra input, decisions on species length bins part 1

--- 2021-03-15 docs/SimData.Rmd
@@
 # intake same for all teleosts
-intakespp  <- rbind(species = sppsubset$Name,
+intakespp  <- rbind(species = sort.default(sppsubset$Name),
                     alpha = rep(0.004),
@@
 # predator size preference ratio same for all species
-sizepref <- rbind(species = sppsubset$Name,
+sizepref <- rbind(species = sort.default(sppsubset$Name),
                   mu = rep(0.5),
@@
 sizepref <- as.data.frame(sizepref[-1,])
+
+# dimensions for singletons file
+singletons <- data.frame(
+  debug=3,
+  Nyrs=NOBAom_ms$runpar$nyears, #may want to take from survey time series instead
+  Nspecies=length(sppsubset$Name),
+  Nsizebins=5,
+  Nareas=1,
+  Nfleets=5,
+  Nsurveys=2,
+  wtconv=1,
+  yr1Nphase=0,
+  recphase=0,
+  avg_rec_phase=0,
+  avg_F_phase=0,
+  dev_rec_phase=0,
+  dev_F_phase=0,
+  fqphase=0,
+  sqphase=0,
+  ssig_phase=0,
+  csig_phase=0,
+  phimax=1,
+  scaleInitialN=1,
+  otherFood=0000,
+  LFI_size=60,
+  bandwidth_metric=5,
+  assessmentPeriod=3,
+  flagMSE=0,
+  flagLinearRamp=1,
+  baseline_threshold=0.2
+)
+
+# M1 same for all species, depends on n length bins
+M1mat <- matrix(0.02, singletons$Nspecies, singletons$Nsizebins, 
+                dimnames = list(c(sort.default(sppsubset$Name)),  c(paste0("sizeclass",seq(1:singletons$Nsizebins))))) 
+M1 <- as.data.frame(M1mat)
 
@@
 write.csv(sizepref, paste0(o.name,"mortality_M2sizePreference_NOBA.csv"))
+write.table(t(singletons), paste0(o.name,"singletons_NOBA.csv"), sep=",",  col.names=FALSE)
+write.csv(M1, paste0(o.name,"mortality_M1_NOBA.csv"))
+
+```
+
+Decisions on size bins for hydra: keep 5 bins per species for now, build in flexibility to allocate size data to any number of bins with any definitions.
+
+Species specific bin width definitions are an input into the model, with more refined (smaller) bin sizes during rapid growth phases.
+
+Initially, we'll directly transfer bin widths for GB and NOBA common species:  
+* Aherring = Norwegian_ssh,
+* Acod = North_atl_cod, 
+* haddock = Haddock, 
+* Amackerel = Mackerel.
+
+We will substitute bin widths from these GB species for NOBA:
+* silverhake ~ Blue_whiting
+* yellowtailfl ~ Long_rough_dab
+
+NOBA Redfish
+NOBA Saithe
+
+NOBA Polar_cod and Capelin are smaller than any GB modeled species
+
+NOBA Green_halibut are larger than any GB modeled species
+
+These decisions will carry through to other life history parameters, at least initially.
+
+Need to fit growth and maturity parameters from Atlantis output for input into hydra.
+Possibly also recruitment?
+Maturity parameters are needed for SSB calculation and SSB-based recruitment.
+If we fit only to average recruitment with annual deviations, and don't need SSB, maturity and recruitment inputs could be dummy parameters for now.
+
+```{r fitpars}
+
+# need growth, vonB t0, K, and Linf and simple growth function
+# would you really base this on 140 years of data? I used all survey years
+
+len_comp_data <- read_savedsurvs(d.name, 'survLen')
+
+#multiple surveys named in list object, use as filename
+for(s in names(len_comp_data)){
+  #arrange into wide format: year, Species1, Species2 ... and write csv
+  svlen <- len_comp_data[[s]][[1]] %>%
+    #calc_timestep2time(d.name, ., run.prm.file) %>% #this gives an arbitrary start year of 1970
+    dplyr::mutate(year = ceiling(time/stepperyr)) %>%
+    dplyr::select(species, year, atoutput) %>%
+    tidyr::spread(species, atoutput) %>%
+    write_csv(paste0(o.name,"observation_lengths_NOBA_",s,".csv"))
+}
+
+# need 2-par maturity curve based on length, atlantis maturity ogive based on age group
 

### 2021-03-17 docs/SimData.Rmd
add NOBA hydra sizebins

@@
 Species specific bin width definitions are an input into the model, with more refined (smaller) bin sizes during rapid growth phases.
+*Consider a function to do this using species max size and the desired number of bins for more consistency and to easily change Nsizebins.*
 
 Initially, we'll directly transfer bin widths for GB and NOBA common species:
-* Aherring = Norwegian_ssh,
-* Acod = North_atl_cod, 
-* haddock = Haddock, 
-* Amackerel = Mackerel.
 
-We will substitute bin widths from these GB species for NOBA:
-* silverhake ~ Blue_whiting
-* yellowtailfl ~ Long_rough_dab
+ *  Aherring = [Norwegian_ssh](https://www.hi.no/en/hi/temasider/species/herring) 
+ *  Acod = [North_atl_cod](https://www.hi.no/en/hi/temasider/species/cod--northeast-arctic) (*but max size could be bigger in simulated data, may need to increase last bin*)
+ *  haddock = [Haddock](https://www.hi.no/en/hi/temasider/species/haddock) (*but max size could be bigger in simulated data, may need to increase last bin*)
+ *  Amackerel = [Mackerel](https://www.hi.no/en/hi/temasider/species/mackerel). (*note, should mackerel have smaller bin widths for fast early growth?*)
 
-NOBA Redfish
-NOBA Saithe
+We will substitute bin widths from these similar taxa/size GB species for NOBA:
 
-NOBA Polar_cod and Capelin are smaller than any GB modeled species
+ *  silverhake ~ [Blue_whiting](https://www.hi.no/en/hi/temasider/species/blue-whiting)  
+ *  yellowtailfl ~ [Long_rough_dab](https://www.sciencedirect.com/science/article/abs/pii/S1385110196907972) a.k.a. American plaice
 
-NOBA Green_halibut are larger than any GB modeled species
+We will modify or subsitute bin widths for the rest of the NOBA species based on size:
 
-These decisions will carry through to other life history parameters, at least initially.
+ *  NOBA [Redfish](https://www.hi.no/en/hi/temasider/species/redfish/beaked-redfish-in-the-barents-and-norwegian-seas) have a maximum size of 47 cm, use 5 x 10 cm bins  
+ *  NOBA [Saithe](https://www.hi.no/en/hi/temasider/species/northeast-arctic-saithe) have a maximum size of 130 cm; use goosefish bins
+ *  NOBA [Polar_cod](https://www.hi.no/en/hi/temasider/species/polar-cod) have a maximum size of 25 cm; use 5 x 5 cm bins  
+ *  NOBA [Capelin](https://www.hi.no/en/hi/temasider/species/capelin-in-the-barents-sea) are smaller than any GB modeled species, max size 21 cm; use 5 x 5 cm bins  
+ *  NOBA [Green_halibut](https://www.hi.no/en/hi/temasider/species/greenland-halibut) are large, max size 120 cm; use Acod bins
+
+Code to make sizebins csv for input; hardcoded based on these decisions.
+
+```{r sizebins, eval=FALSE}
+# for reference, GB hydra_sim size bins
+# sizeclass1,sizeclass2,sizeclass3,sizeclass4,sizeclass5,commentField-notused
+# 20,20,20,20,30,spinydog
+# 20,20,20,20,40,winterskate
+# 5,5,5,5,20,Aherring
+# 20,20,20,40,50,Acod
+# 10,10,20,20,20,haddock
+# 10,10,10,10,20,yellowtailfl
+# 10,10,10,10,10,winterfl
+# 10,10,10,10,10,Amackerel
+# 10,10,10,10,30,silverhake
+# 20,20,20,30,40,goosefish
+
+sizebinmat <- matrix(0, singletons$Nspecies, singletons$Nsizebins, 
+                dimnames = list(c(sort.default(sppsubset$Name)),  c(paste0("sizeclass",seq(1:singletons$Nsizebins)))))
+
+sizebinmat["Blue_whiting",] <- c(10,10,10,10,30)
+sizebinmat["Capelin",] <- c(5,5,5,5,5)
+sizebinmat["Green_halibut",] <- c(20,20,20,40,50)
+sizebinmat["Haddock",] <- c(10,10,20,20,20)
+sizebinmat["Long_rough_dab",] <- c(10,10,10,10,20)
+sizebinmat["Mackerel",] <- c(10,10,10,10,10)
+sizebinmat["North_atl_cod",] <- c(20,20,20,40,50)
+sizebinmat["Norwegian_ssh",] <- c(5,5,5,5,20)
+sizebinmat["Polar_cod",] <- c(5,5,5,5,5)
+sizebinmat["Redfish",] <- c(10,10,10,10,10)
+sizebinmat["Saithe",] <- c(20,20,20,30,40)
+
+sizebins <- as.data.frame(sizebinmat) %>%
+  mutate('commentField-notused' = rownames(sizebinmat))
+
+write_csv(sizebins, paste0(o.name,"length_sizebins_NOBA.csv"))
+
+```
+
+These decisions on matching species can carry through to other life history parameters, at least initially.
 
@@
 
-```{r fitpars}
+```{r fitpars, eval=FALSE}
 

### 2021-03-17 docs/SimData.Rmd
outline binwidth function

@@
 Species specific bin width definitions are an input into the model, with more refined (smaller) bin sizes during rapid growth phases.
-*Consider a function to do this using species max size and the desired number of bins for more consistency and to easily change Nsizebins.*
+
+*Consider a function to do this using species max size, the vonB k and the desired number of bins for more consistency and to easily change Nsizebins.*
+```{r lbinfunction, eval=FALSE}
+# see e.g., https://www.sciencedirect.com/science/article/abs/pii/S0165783619303297
+# estimated optimal bin size for VBGF estimation is 0.23 x Lmax ^0.6
+
+binwidths <- function(Lmax, vonBk, Nsizebins){
+  #optimal bin size is lower end
+  binopt <- 0.23*Lmax^0.6
+  #vonBk tells you how many fine bins for fast growth
+
+  #aggregate bins in multiples of optimal for bigger fish/slower growth to get Nsizebins
+
+}
+
+```
 
@@
   svlen <- len_comp_data[[s]][[1]] %>%
-    #calc_timestep2time(d.name, ., run.prm.file) %>% #this gives an arbitrary start year of 1970
     dplyr::mutate(year = ceiling(time/stepperyr)) %>%
-    dplyr::select(species, year, atoutput) %>%
+    dplyr::select(species, agecl, year, upper.bins, atoutput) %>%
+
     tidyr::spread(species, atoutput) %>%

### 2021-03-22 docs/GBFleetDefinitions.Rmd

@@
 All gear types which catch any amount of the [Species](https://noaa-edab.github.io/ms-keyrun/SpeciesEnvDat.html#focalSpecies) of interest are used in the cluster analysis.
+The hope was the set of gear types in which to begin the cluster analysis would be substantially less.
+The number of gears eliminated = 10.
+It is apparent that species of interest are caught across a broad class of gear types.
+
 ```{r echo=T,eval=F}
@@
 
+## Decision
+
+This exploratory study resulted in the following designation of [Fleets](https://noaa-edab.github.io/ms-keyrun/GBFleetDecision.html)
+
 ## References

### 2021-03-22 docs/index.Rmd
add survey/catch dat csvs for hydra
add NOBA hydra sizebins

@@
 
+<a href="https://noaa-edab.github.io/ms-keyrun/GBFleetDefinitions.html" target="_blank"> Fleet Definitions - 26 Feb 2021 </a>
+
+* <a href="https://noaa-edab.github.io/ms-keyrun/GBFleetDecisions.html" target="_blank"> Fleet Decisions </a>
+
 <a href="https://noaa-edab.github.io/ms-keyrun/SpeciesEnvDat.html" target="_blank"> Species, environment, data package decisions - 5 August 2020 </a>
@@
 ## Dataset descriptions and links
+
+<a href="https://noaa-edab.github.io/ms-keyrun/SimData.html" target="_blank"> Initial simulated data for testing - 17 March 2021 </a>
 

### 2021-03-23 docs/SimData.Rmd

@@
 
+# placeholders for unused covariates dimenstioned to the number of simulated years
+dummycovar <- data.frame(dummy1 = rep(1, NOBAom_ms$runpar$nyears))
+
 # will need to fit von B and maturity curve functions to back out hydra input pars
@@
 write.csv(M1, paste0(o.name,"mortality_M1_NOBA.csv"))
+write_csv(dummycovar, paste0(o.name,"recruitment_covariates_NOBA.csv"))
+write_csv(dummycovar, paste0(o.name,"maturity_covariates_NOBA.csv"))
+write_csv(dummycovar, paste0(o.name,"growth_covariates_NOBA.csv"))
 
@@
 
+Output temperature data from surveys using "Temp" variable in .nc file and the `atlantisom::load_nc_physics` function.
+A survey (timestep, polygons, possibly observation error) can then be applied to this output.
+
+We can get surface and bottom temp from atlatis using the layers from each polygon, with layer 0 (?) being the surface and the maximum layer in the polygon being the bottom. 
+```{r}
+
+# taken from existing load_nc
+
+file.nc <- file.path(d.name, "outputnordic_runresults_01.nc")
+
+  # Load ATLANTIS output!
+  at_out <- RNetCDF::open.nc(con = file.nc)
+
+  # Get info from netcdf file! (Filestructure and all variable names)
+  var_names_ncdf <- sapply(seq_len(RNetCDF::file.inq.nc(at_out)$nvars - 1),
+    function(x) RNetCDF::var.inq.nc(at_out, x)$name)
+  n_timesteps <- RNetCDF::dim.inq.nc(at_out, 0)$length
+  n_boxes     <- RNetCDF::dim.inq.nc(at_out, 1)$length
+  n_layers    <- RNetCDF::dim.inq.nc(at_out, 2)$length
+
+RNetCDF::close.nc(at_out)
+
+#'@physic_variables A character value spefifying which variables
+#'   should be loaded. Only one variable can be loaded at a time.
+#'   Currently, only the following variables are allowed:
+#'   "salt", "NO3", "NH3", "Temp", "Oxygen", "Si", "Det_Si", "DON",
+#'    "Chl_a", "Denitrifiction", "Nitrification".
+
+truetemp <- atlantisom::load_nc_physics(dir = d.name,
+                  file_nc = "outputnordic_runresults_01.nc",
+                  physic_variables = "Temp",
+                  aggregate_layers = FALSE,
+                  bboxes = atlantisom::get_boundary(NOBAom_ms$boxpars))
+  #if(verbose) message("Temperature read in.")
+
+truetempagg <- atlantisom::load_nc_physics(dir = d.name,
+                  file_nc = "outputnordic_runresults_01.nc",
+                  physic_variables = "Temp",
+                  aggregate_layers = TRUE,
+                  bboxes = atlantisom::get_boundary(NOBAom_ms$boxpars))
+
+```
+
 ## Generating a survey dataset

### 2021-03-23 docs/GBFleetDecisions.Rmd

---
title: "Fleet Definitions for Georges Bank"
author: Andy Beet
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    code_fold: hide
    number_sections: false
csl: apa-annotated-bibliography.csl
bibliography: references.bib
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
options(warn=-1)
```

The fleets designations were made after an exploratory [cluster analysis](https://noaa-edab.github.io/ms-keyrun/GBFleetDefinitions.html) study

Note:

<sup>1</sup> indicates a gear code < 1% of total landings

<sup>1,2</sup> indicates a gear code < 0.1% of total landings

All other gear codes make up the majority (99%) of the landings. For these gear codes each species contribution to the total landings (aggregated over time) are shown as percentages

Species of interest are **bolded**

```{r ,eval=F, echo=T}
gearData <- readRDS(file=paste0(here::here("data","timeSeriesSpeciesByGear.rds")))
gearData$data %>% dplyr::filter(NEGEAR2==gearCode)

```

## Demersal

05 - OTTER TRAWL, BOTTOM (**cod** (16%), **yellowtail fl** (15%), **haddock** (12%), **silver hake** (10%), **winter fl** (8%), loligo (6%) + 33% other bottom) - Large and small mesh

<sup>1</sup> 35 - BEAM TRAWL (summer flounder, **winter flounder**, **yellowtail flounder**)

<sup>1</sup> 36 - SCOTTISH SEINE (**silver hake**, **cod**, plaice, witch fl, **yellowtail flounder**, **winter fl**, red hake)

<sup>1</sup> 16 - DANISH SEINE (**cod**, witch fl, **yellowtail flounder**, **winter fl**, plaice, **silver hake**)

<sup>1</sup>  32 - FYKE NET (**winter fl**, windowpane fl, **dogfish**, **skates**)

## Pelagic

10 - GILL NET, FIXED (**dogfish** (22%), **cod**(22%), **goosefish** (20%))

01 - LONGLINE, BOTTOM (**cod** (40%), **dogfish** (21%), tilefish (20%), **haddock** (9%) + pollock , white hake)

02 - HANDLINE (**dogfish**(22%), **cod** (22%), **goosefish** (20%), **skates**(20%), pollock (7%))

<sup>1</sup>  52 - GILL NET (skate, angler, **haddock**, swordfish, **mackerel**, **cod** )

<sup>1</sup>  50 - GILL NET (angler, bluefish, **skate**, **dogfish**)

## Squid

14 - POUND NET (loligo (35%) , scup (20%), **mackerel** (20%), many other < 5%)

<sup>1</sup> 8 - FLOATING TRAP (**mackerel**, scup, loligo, tuna)

<sup>1</sup>  26 - WEIR (loligo, **mackerel**, scup)

## Forage Fish

17 - MIDWATER OTTER TRAWL (**Mackerel** (7%), **Herring** (91%))

37 - MIDWATER OTTER TRAWL (**Mackerel** (11%) , **Herring** (88%))

12 - PURSE SEINE (1970's menhaden, modern Tuna, **Herring**)

## Shellfish

38 - DREDGE (mussel(40%), scallop(30%), clam+quahog (17%) + bottom fish)

40 - DREDGE (surfclam (56%), ocean quahog(43%) )

13 - DREDGE, SCALLOP (scallop (88%) (**Goosefish** , **yellowtail**, bottom fish))

<sup>1</sup> 22 - HOES/FORK (soft clam, oysters, bay scallop)

<sup>1</sup> 41 - TONGS AND GRABS (quahog, soft clam, surf clam)

<sup>1</sup> 25 - RAKES (quahog, soft clam, mussels, crab, oysters)

## Lobster/Crab

21 - POT/TRAP (Lobster (88%), crab (10%) + bottom fish)

20 - POT/TRAP (lobster(53%) + crab(46%), bottom fish < 1%)

30 - POTS + TRAPS (crab (>99%))

18 - POTS + TRAPS (crab (20%), lobster (20%), whelk (10%), snails (10%), hagfish (20%) + bottom fish (20%))

<sup>1</sup> 15 - TRAWL (lobster, 1972,1974)

<sup>1</sup> 53 - TRAMMEL NET (lobster , 2012)

<sup>1</sup> 19 - POTS + TRAPS (rock crab, lobster, 2004)

<sup>1</sup> 23 - BY HAND (soft clam, crab, quahog, oysters, clam)

<sup>1</sup> 33 - DIVING GEAR (lobster, sea urch, squid)

## HMS

<sup>1</sup> 3 - HARPOONS (Tuna, swordfish)

<sup>1</sup>  11 - GILL NET, DRIFT (swordfish, tuna, shark )

<sup>1</sup>  6 - TROLL LINE (Tuna, sea herring, striped bass, bluefish)

<sup>1</sup>  4 - LONGLINE (swordfish, tunas, tilefish)

## Other gear

<sup>1</sup> 28 - DREDGE (southern flounder, 1989-1997)

<sup>1</sup> 27 - BOX TRAP (cod, angler, lobster, 2004)

<sup>1</sup> 99 - UNKNOWN (silver hake, oysters, lobster, quahog, sea herring, angler, mussels )

<sup>1,2</sup> 44 - WEEDWHACKER, SEAWEED (whelks, 2016)

<sup>1,2</sup> 31 - LIFT NET (lobster, 2002)

<sup>1,2</sup> 34 - CAST NETS (quahog, surf clam, skate, plaice )

<sup>1,2</sup> 81 - SUCTION PUMP ()

<sup>1,2</sup> 45 - TRAWL, SHRIMP  (crustacean,  2017)

<sup>1,2</sup> 9 - DIP NET (alewife, yellow flounder, scallop, soft clam)

<sup>1,2</sup> 7 - HAUL SEINE (perch, alewife, cod, winter flounder, lobster)

### 2021-03-23 docs/SimData.Rmd

@@
 
-We can get surface and bottom temp from atlatis using the layers from each polygon, with layer 0 (?) being the surface and the maximum layer in the polygon being the bottom. 
-```{r}
+We can get surface and bottom temp from Atlantis using the layers from each polygon, with layer 0 being the bottom and the maximum layer in the polygon being the surface (according to the [Atlantis wiki/Atlantis output files/Atlantis NetCDF structure](https://confluence.csiro.au/display/Atlantis/Atlantis+NetCDF+Structure).
+```{r surveytemp, eval=FALSE}
 
-# taken from existing load_nc
-
-file.nc <- file.path(d.name, "outputnordic_runresults_01.nc")
-
-  # Load ATLANTIS output!
-  at_out <- RNetCDF::open.nc(con = file.nc)
-
-  # Get info from netcdf file! (Filestructure and all variable names)
-  var_names_ncdf <- sapply(seq_len(RNetCDF::file.inq.nc(at_out)$nvars - 1),
-    function(x) RNetCDF::var.inq.nc(at_out, x)$name)
-  n_timesteps <- RNetCDF::dim.inq.nc(at_out, 0)$length
-  n_boxes     <- RNetCDF::dim.inq.nc(at_out, 1)$length
-  n_layers    <- RNetCDF::dim.inq.nc(at_out, 2)$length
-
-RNetCDF::close.nc(at_out)
+# taken from existing load_nc to explore
+# file.nc <- file.path(d.name, "outputnordic_runresults_01.nc")
+# 
+#   # Load ATLANTIS output!
+#   at_out <- RNetCDF::open.nc(con = file.nc)
+# 
+#   # Get info from netcdf file! (Filestructure and all variable names)
+#   var_names_ncdf <- sapply(seq_len(RNetCDF::file.inq.nc(at_out)$nvars - 1),
+#     function(x) RNetCDF::var.inq.nc(at_out, x)$name)
+#   n_timesteps <- RNetCDF::dim.inq.nc(at_out, 0)$length
+#   n_boxes     <- RNetCDF::dim.inq.nc(at_out, 1)$length
+#   n_layers    <- RNetCDF::dim.inq.nc(at_out, 2)$length
+# 
+# RNetCDF::close.nc(at_out)
 
@@
 
-truetempagg <- atlantisom::load_nc_physics(dir = d.name,
-                  file_nc = "outputnordic_runresults_01.nc",
-                  physic_variables = "Temp",
-                  aggregate_layers = TRUE,
-                  bboxes = atlantisom::get_boundary(NOBAom_ms$boxpars))
+# this averages all depth layers, don't want this
+# truetempagg <- atlantisom::load_nc_physics(dir = d.name,
+#                   file_nc = "outputnordic_runresults_01.nc",
+#                   physic_variables = "Temp",
+#                   aggregate_layers = TRUE,
+#                   bboxes = atlantisom::get_boundary(NOBAom_ms$boxpars))
+
+truebottomtemp <- truetemp %>%
+  filter(layer == 0) # according to Atlantis wiki, layer 0 is bottom
+
+truesurfacetemp <- truetemp %>%
+  filter(layer<7) %>%  #layer 7 is always identical temp to layer 0, may be the added sediment layer?
+  group_by(polygon, time) %>%
+  filter(layer==max(layer)) %>%
+  ungroup
+
+surveyenv <- function(trueenv, 
+                      survboxes,
+                      survtimes) {
+  surveydat <- trueenv %>%
+    dplyr::filter(polygon %in% survboxes) %>%
+    dplyr::filter(time %in% survtimes)
+
+  return(surveydat)
+}
+
+omlist_ss <- NOBAom_ms
+source(here("simulated-data/config/omdimensions.R"))
+
+#need to weight by polygon area to get mean annual temp time series for each survey
+boxarea <- map_dfr(NOBAom_ms$boxpars$boxes, "area") %>%
+  pivot_longer(everything(), names_to = "polygon", values_to = "area") %>%
+  mutate(polygon = as.integer(polygon)) %>%
+  mutate(proparea = area/sum(area))
+
+source(here("simulated-data/config/mssurvey_fall.R"))
+
+fallcensus_bottomtemp <- surveyenv(truebottomtemp, survboxes = survboxes, survtimes = survtime)
+
+fallcensus_btempindex <- left_join(fallcensus_bottomtemp, boxarea) %>%
+  dplyr::group_by(time) %>%
+  dplyr::summarise(meantemp = weighted.mean(atoutput, proparea)) %>%
+  dplyr::mutate(year = ceiling(time/stepperyr)) %>%
+  dplyr::select(year, meantemp) %>%
+  write_csv(paste0(o.name,"observation_temperature_NOBA_",survey.name,".csv"))
+
+source(here("simulated-data/config/mssurvey_spring.R"))
+
+springcensus_bottomtemp <- surveyenv(truebottomtemp, survboxes = survboxes, survtimes = survtime)
+
+springcensus_btempindex <- left_join(springcensus_bottomtemp, boxarea) %>%
+  dplyr::group_by(time) %>%
+  dplyr::summarise(meantemp = weighted.mean(atoutput, proparea)) %>%
+  dplyr::mutate(year = ceiling(time/stepperyr)) %>%
+  dplyr::select(year, meantemp) %>%
+  write_csv(paste0(o.name,"observation_temperature_NOBA_",survey.name,".csv"))
+
+source(here("simulated-data/config/mssurvey_fall_01.R"))
+
+fallnearbox_bottomtemp <- surveyenv(truebottomtemp, survboxes = survboxes, survtimes = survtime)
+
+fallnearbox_btempindex <- left_join(fallnearbox_bottomtemp, boxarea) %>%
+  dplyr::group_by(time) %>%
+  dplyr::summarise(meantemp = weighted.mean(atoutput, proparea)) %>%
+  dplyr::mutate(year = ceiling(time/stepperyr)) %>%
+  dplyr::select(year, meantemp) %>%
+  write_csv(paste0(o.name,"observation_temperature_NOBA_",survey.name,".csv"))
+
+source(here("simulated-data/config/mssurvey_spring_01.R"))
+
+springnearbox_bottomtemp <- surveyenv(truebottomtemp, survboxes = survboxes, survtimes = survtime)
+
+springnearbox_btempindex <- left_join(springnearbox_bottomtemp, boxarea) %>%
+  dplyr::group_by(time) %>%
+  dplyr::summarise(meantemp = weighted.mean(atoutput, proparea)) %>%
+  dplyr::mutate(year = ceiling(time/stepperyr)) %>%
+  dplyr::select(year, meantemp) %>%
+  write_csv(paste0(o.name,"observation_temperature_NOBA_",survey.name,".csv"))
 

### 2021-03-23 docs/SimData.Rmd
more dummy data fleet effort

@@
   Nareas=1,
-  Nfleets=5,
+  Nfleets=length(unique(NOBAom_ms$truecatchage_ss$fleet)),
   Nsurveys=2,
@@
 # M1 same for all species, depends on n length bins
+# to get annual M1 of 0.1, this matrix needs to be M1 per timestep
+# right now input works for 5 timesteps per year, but should be 
+# done in the code with input annual M1 in case timestep changes
 M1mat <- matrix(0.02, singletons$Nspecies, singletons$Nsizebins, 
@@
 
+# sexratio fixed 50:50 for all species
+# sexratio <- vector(0.5, )
+# rbind(species = sort.default(sppsubset$Name),
+#                   ratio = rep(0.5))
+# colnames(sexratio) <- sexratio[1,]
+# sexratio <- as.data.frame(sexratio[-1,])
+
 # placeholders for unused covariates dimenstioned to the number of simulated years
 dummycovar <- data.frame(dummy1 = rep(1, NOBAom_ms$runpar$nyears))
+
+# placeholder for effort data dimensioned by year and nfleets
+allfleets <- atlantisom::load_fisheries(d.name, fisheries.file)
+fleets <- unique(NOBAom_ms$truecatchage_ss$fleet)
+
+dummyeffort <- data.frame(1:singletons$Nyrs,
+                          matrix(1, singletons$Nyrs, singletons$Nfleets))
+names(dummyeffort) <- c("year", fleets)
 
@@
 write_csv(dummycovar, paste0(o.name,"growth_covariates_NOBA.csv"))
+write_csv(dummyeffort, paste0(o.name,"observation_effort_NOBA.csv"))
 

### 2021-03-23 docs/GBFleetDecisions.Rmd
heading changes. split shellfish

@@
 
-05 - OTTER TRAWL, BOTTOM (**cod** (16%), **yellowtail fl** (15%), **haddock** (12%), **silver hake** (10%), **winter fl** (8%), loligo (6%) + 33% other bottom) - Large and small mesh
+05 - OTTER TRAWL, BOTTOM (**cod** (16%), **yellowtail fl** (15%), **haddock** (12%), **silver hake** (10%), **winter fl** (8%), loligo (6%) + 33% other bottom) - Large and small mesh - 7,323,551,019 lbs
 
-<sup>1</sup> 35 - BEAM TRAWL (summer flounder, **winter flounder**, **yellowtail flounder**)
+<sup>1</sup> 16 - DANISH SEINE (**cod**, witch fl, **yellowtail flounder**, **winter fl**, plaice, **silver hake**) - 5,194,870 lbs
 
-<sup>1</sup> 36 - SCOTTISH SEINE (**silver hake**, **cod**, plaice, witch fl, **yellowtail flounder**, **winter fl**, red hake)
+<sup>1</sup> 36 - SCOTTISH SEINE (**silver hake**, **cod**, plaice, witch fl, **yellowtail flounder**, **winter fl**, red hake) - 4,521,702 lbs
 
-<sup>1</sup> 16 - DANISH SEINE (**cod**, witch fl, **yellowtail flounder**, **winter fl**, plaice, **silver hake**)
+<sup>1</sup> 35 - BEAM TRAWL (summer flounder, **winter flounder**, **yellowtail flounder**) - 927,806 lbs
 
-<sup>1</sup>  32 - FYKE NET (**winter fl**, windowpane fl, **dogfish**, **skates**)
+<sup>1</sup>  32 - FYKE NET (**winter fl**, windowpane fl, **dogfish**, **skates**) - 1,602 lb
 
-## Pelagic
+## Fixed Gear
+10 - GILL NET, FIXED (**dogfish** (22%), **cod** (22%), **goosefish** (20%), **winter skate/skates** (21%), pollock (6%)) - 417,501,765 lbs
 
-10 - GILL NET, FIXED (**dogfish** (22%), **cod**(22%), **goosefish** (20%))
+01 - LONGLINE, BOTTOM (**cod** (40%), **dogfish** (21%), tilefish (20%), **haddock** (9%) + pollock , white hake) - 234,755,227 lbs
 
-01 - LONGLINE, BOTTOM (**cod** (40%), **dogfish** (21%), tilefish (20%), **haddock** (9%) + pollock , white hake)
+02 - HANDLINE (**cod**(42%), scup (15%), tuna (13%), **spiny dogfish**(6%), **mackerel** (5%)) -72,859,422 lbs
 
-02 - HANDLINE (**dogfish**(22%), **cod** (22%), **goosefish** (20%), **skates**(20%), pollock (7%))
+<sup>1</sup>  50 - GILL NET (angler, bluefish, **skate**, **dogfish**) - 789,780 lbs
 
-<sup>1</sup>  52 - GILL NET (skate, angler, **haddock**, swordfish, **mackerel**, **cod** )
+<sup>1</sup>  52 - GILL NET (skate, angler, **haddock**, swordfish, **mackerel**, **cod** ) - 367,823 lbs
 
-<sup>1</sup>  50 - GILL NET (angler, bluefish, **skate**, **dogfish**)
+## Trap
 
-## Squid
+14 - POUND NET (loligo (35%) , scup (20%), **mackerel** (20%), many other < 5%) - 39,571,508 lbs
 
-14 - POUND NET (loligo (35%) , scup (20%), **mackerel** (20%), many other < 5%)
+<sup>1</sup>  26 - WEIR (loligo, **mackerel**, scup) - 3,094,066 lbs
 
-<sup>1</sup> 8 - FLOATING TRAP (**mackerel**, scup, loligo, tuna)
-
-<sup>1</sup>  26 - WEIR (loligo, **mackerel**, scup)
+<sup>1</sup> 8 - FLOATING TRAP (**mackerel**, scup, loligo, tuna) - 1,850,040 lbs
 
@@
 
-17 - MIDWATER OTTER TRAWL (**Mackerel** (7%), **Herring** (91%))
+17 - MIDWATER OTTER TRAWL (**Mackerel** (7%), **Herring** (91%)) -  598,176,314 lbs
 
-37 - MIDWATER OTTER TRAWL (**Mackerel** (11%) , **Herring** (88%))
+37 - MIDWATER OTTER TRAWL (**Mackerel** (11%) , **Herring** (88%)) - 367,414,353 lbs
 
-12 - PURSE SEINE (1970's menhaden, modern Tuna, **Herring**)
+12 - PURSE SEINE (1970's menhaden (40%), Tuna (34%), **Herring** (5%)) - 41,291,228 lbs
 
-## Shellfish
+## Scallop
 
-38 - DREDGE (mussel(40%), scallop(30%), clam+quahog (17%) + bottom fish)
+13 - DREDGE, SCALLOP (scallop (88%) (**Goosefish** , **yellowtail**, bottom fish)) - 733,761,209 lbs
 
-40 - DREDGE (surfclam (56%), ocean quahog(43%) )
+## Surf clam
 
-13 - DREDGE, SCALLOP (scallop (88%) (**Goosefish** , **yellowtail**, bottom fish))
+40 - DREDGE (surfclam (56%), ocean quahog(43%) ) - 425,058,521 lbs
 
-<sup>1</sup> 22 - HOES/FORK (soft clam, oysters, bay scallop)
+## Other shellfish
 
-<sup>1</sup> 41 - TONGS AND GRABS (quahog, soft clam, surf clam)
+38 - DREDGE (mussel(40%), sea scallop(30%), clam+quahog (17%) + bottom fish) - 48,683,622 lbs
 
-<sup>1</sup> 25 - RAKES (quahog, soft clam, mussels, crab, oysters)
+<sup>1</sup> 25 - RAKES (quahog, soft clam, mussels, crab, oysters) - 13,153,618 lbs
+
+<sup>1</sup> 41 - TONGS AND GRABS (quahog, soft clam, surf clam) - 232,351 lbs
+
+<sup>1</sup> 22 - HOES/FORK (soft clam, oysters, bay scallop) - 217,491 lbs
 
@@
 
-21 - POT/TRAP (Lobster (88%), crab (10%) + bottom fish)
+20 - POT/TRAP (lobster(53%) + crab(46%), bottom fish < 1%) - 340,059,280 lbs
 
-20 - POT/TRAP (lobster(53%) + crab(46%), bottom fish < 1%)
+18 - POTS + TRAPS (crab (20%), lobster (20%), whelk (10%), snails (10%), hagfish (20%) + bottom fish (20%)) - 79,123,496 lbs
 
-30 - POTS + TRAPS (crab (>99%))
+30 - POTS + TRAPS (crab (>99%)) - 68,329,188 lbs
 
-18 - POTS + TRAPS (crab (20%), lobster (20%), whelk (10%), snails (10%), hagfish (20%) + bottom fish (20%))
+21 - POT/TRAP (Lobster (88%), crab (10%) + bottom fish) - 39,633,570 lbs
 
-<sup>1</sup> 15 - TRAWL (lobster, 1972,1974)
+<sup>1</sup> 23 - BY HAND (soft clam, crab, quahog, oysters, clam) - 1,665,727 lbs
 
-<sup>1</sup> 53 - TRAMMEL NET (lobster , 2012)
+<sup>1</sup> 33 - DIVING GEAR (lobster, sea urch, squid) - 878,986 lbs
 
-<sup>1</sup> 19 - POTS + TRAPS (rock crab, lobster, 2004)
+<sup>1</sup> 15 - TRAWL (lobster, 1972,1974) - 21,437 lbs
 
-<sup>1</sup> 23 - BY HAND (soft clam, crab, quahog, oysters, clam)
+<sup>1</sup> 19 - POTS + TRAPS (rock crab, lobster, 2004) - 23525 lbs
 
-<sup>1</sup> 33 - DIVING GEAR (lobster, sea urch, squid)
+<sup>1</sup> 53 - TRAMMEL NET (lobster , 2012) - 193 lbs
 

### 2021-03-24 docs/SimData.Rmd
more fixed hydra inputs and array shape outline

@@
 # sexratio fixed 50:50 for all species
-# sexratio <- vector(0.5, )
-# rbind(species = sort.default(sppsubset$Name),
-#                   ratio = rep(0.5))
-# colnames(sexratio) <- sexratio[1,]
-# sexratio <- as.data.frame(sexratio[-1,])
+sexratio <- matrix(0.5, 1, singletons$Nspecies,
+                   dimnames = list("ratio", c(sort.default(sppsubset$Name))))
 
@@
 write_csv(dummyeffort, paste0(o.name,"observation_effort_NOBA.csv"))
+write.csv(sexratio, paste0(o.name,"sexratio_NOBA.csv"))
 
@@
 
-Need to fit growth and maturity parameters from Atlantis output for input into hydra.
-Possibly also recruitment?
-Maturity parameters are needed for SSB calculation and SSB-based recruitment.
-If we fit only to average recruitment with annual deviations, and don't need SSB, maturity and recruitment inputs could be dummy parameters for now.
+Need to fit growth and maturity parameters from Atlantis output for input into hydra. Fit growth using code below.
+
+Generating length composition input data requires mapping numbers at length in the atlantisom 1 cm bin length output to our selected length bins, then summing the numbers at length in those bins for all species.
+The input required is a 4d array, indexed by survey, species, year, and sizebin.
+So each survey can be a separate csv with #species at the top of year rows and sizebin columns with n at length the values, scrolling down for each species.
 
@@
 
+# VBGF packages all take age and length pairs, reformat from length comp at agecl
+
 #multiple surveys named in list object, use as filename
 for(s in names(len_comp_data)){
-  #arrange into wide format: year, Species1, Species2 ... and write csv
-  svlen <- len_comp_data[[s]][[1]] %>%
+  #arrange into long format: year, Species, agecl, length 
+  svlenage <- len_comp_data[[s]][[1]] %>%
     dplyr::mutate(year = ceiling(time/stepperyr)) %>%
@@
 
-    tidyr::spread(species, atoutput) %>%
+}
+
+# rearrange into length bins and write length comp data csvs
+#multiple surveys named in list object, use as filename
+for(s in names(len_comp_data)){
+  #arrange into long format: year, Species, agecl, length 
+  svlenbin <- len_comp_data[[s]][[1]] %>%
+    dplyr::mutate(year = ceiling(time/stepperyr)) %>%
+    dplyr::select(species, agecl, year, upper.bins, atoutput) %>%
+
+    tidyr::spread(lenbins, atoutput) %>%
     write_csv(paste0(o.name,"observation_lengths_NOBA_",s,".csv"))
@@
 # need 2-par maturity curve based on length, atlantis maturity ogive based on age group
+
+```
+
+Stomach weights are needed as input to hydra.
+At present each species has a stomach weight for each size bin that is repeated for each year.
+We have total consumed weight for each predator age class at each timestep in the detailed diet atlantis output, so would need to map age classes to length bins, sum consumed weight by length bin and divide by total numbers in that length bin to get stomach weight.
+
+```{r stomwt}
+
+```
+
+Do we need to fit recruitment?
+May just need average recriutment starting values, which I can get from atlantis.
+Maturity parameters are needed for SSB calculation and SSB-based recruitment.
+If we fit only to average recruitment with annual deviations, and don't need SSB, maturity and recruitment inputs could be dummy parameters for now.
+
+```{r}
 

### 2021-03-25 docs/SimData.Rmd
age length pairs continued

@@
 
+# need to shift agecl to represent midpoint of actual age groups
+# using part of the wtage interpolation function calc_avgwtstage2age()
+annage_comp_data <- read_savedsurvs(d.name, 'survAnnAge')
+
+library(FSA)
 # VBGF packages all take age and length pairs, reformat from length comp at agecl
+# break into 3 chunks of time, "era" 48 years each more like our survey
+# subsample after reformatting with uncount because we wouldn't have this many ages
 
@@
 for(s in names(len_comp_data)){
+
+  annages <- annage_comp_data[[s]][[1]] %>%
+    group_by(species, time, agecl) %>%
+    rename(trueage = agecl) %>%
+    summarise(truenatage = sum(atoutput))
+
+  # Figure out the groups that have multiple ages classes in each stage (or
+  # cohort)
+  multiple_ages <- sp_age[sp_age$NumAgeClassSize>1,]
+
+  # finds the average age of each agecl each timestep based on truenatage
+  svlen_avgage <- annages %>%
+    dplyr::semi_join(len_comp_data[[s]][[1]], by = c("species", "time")) %>%
+    dplyr::inner_join(multiple_ages, by = c("species" = "Name")) %>%
+    dplyr::mutate(agecl = as.integer(ceiling(trueage/NumAgeClassSize))) %>%
+    dplyr::group_by(species, time, agecl) %>%
+    dplyr::mutate(avgage = weighted.mean(trueage, truenatage))
+
   #arrange into long format: year, Species, agecl, length 
   svlenage <- len_comp_data[[s]][[1]] %>%
+    dplyr::left_join(svlen_avgage) %>%
     dplyr::mutate(year = ceiling(time/stepperyr)) %>%
-    dplyr::select(species, agecl, year, upper.bins, atoutput) %>%
+    dplyr::select(species, agecl, avgage, year, upper.bins, atoutput) %>%
+    dplyr::mutate(avgage = coalesce(avgage,agecl)) %>%
+    dplyr::rename(length = upper.bins) %>%
+    dplyr::mutate(era = as.numeric(cut_number(year,3))) %>%
+    tidyr::uncount(atoutput) %>%
+    dplyr::group_by(species, avgage, year) %>%
+    dplyr::sample_frac(0.01)
+
+  svvbgf <- svlenage %>%
+    dplyr::group_by(species, era) %>%
+    group_map(~nls(length~vb1(avgage, Linf, K, t0), data = .,
+                  start = vbStarts(length~avgage,data = .)))
 

### 2021-03-26 docs/SimData.Rmd
add test fit results

@@
 
+sp_age <- NOBAom_ms$funct.group_ss[, c("Name", "NumCohorts", "NumAgeClassSize")]
+sp_age <- sp_age %>%
+  mutate(n_annages = NumCohorts * NumAgeClassSize)
+
+stepperyr <- if(NOBAom_ms$runpar$outputstepunit=="days") 365/NOBAom_ms$runpar$toutinc
+
 # need growth, vonB t0, K, and Linf and simple growth function
@@
 library(FSA)
+vb1 <- vbFuns()
 # VBGF packages all take age and length pairs, reformat from length comp at agecl
@@
 
-  svvbgf <- svlenage %>%
-    dplyr::group_by(species, era) %>%
-    group_map(~nls(length~vb1(avgage, Linf, K, t0), data = .,
-                  start = vbStarts(length~avgage,data = .)))
+  saveRDS(svlenage, file.path(d.name, paste0(scenario.name, "_",
+                                                       s, "agelensamp.rds")))
+
+  capelinvbgf <- svlenage %>%
+    filter(species == "Capelin") %>%
+    group_by(era) %>%
+    group_map(try(~nls(length~vb1(avgage, Linf, K, t0), data = .,
+                  start = vbStarts(length~avgage,data = .))))
+
+  # herringvbgf <- svlenage %>%
+  #   filter(species == "Norweigan_ssh", era==1) %>%
+  #   group_map(try(~nls(length~vb1(avgage, Linf, K, t0), data = .,
+  #                 start = vbStarts(length~avgage,data = .))))
+
+#Error in nls(length ~ vb1(avgage, Linf, K, t0), data = ., start = vbStarts(length ~  : 
+  #singular gradient with group_map, era 2 and 3
+  codvbgf <- svlenage %>%
+    filter(species == "North_atl_cod") 
+    #group_by(era) %>%
+    nls(length~vb1(avgage, Linf, K, t0), data = codvbgf,
+                  start = vbStarts(length~avgage,data = codvbgf))
+#  era==1   Nonlinear regression model
+#   model: length ~ vb1(avgage, Linf, K, t0)
+#    data: codvbgf
+#     Linf        K       t0 
+# 104.7120   0.4260   0.2176 
+#  residual sum-of-squares: 4109723
+# 
+# Number of iterations to convergence: 7 
+# Achieved convergence tolerance: 2.169e-06
+
+#  all years Nonlinear regression model
+#   model: length ~ vb1(avgage, Linf, K, t0)
+#    data: codvbgf
+#     Linf        K       t0 
+# 112.4694   0.3912   0.2097 
+#  residual sum-of-squares: 7354906
+# 
+# Number of iterations to convergence: 11 
+# Achieved convergence tolerance: 9.964e-06
+
+  hadvbgf <- svlenage %>%
+    filter(species == "Haddock") 
+    #group_by(era) %>%
+    nls(length~vb1(avgage, Linf, K, t0), data = hadvbgf,
+                  start = vbStarts(length~avgage,data = hadvbgf))
+# all years
+#     Nonlinear regression model
+#   model: length ~ vb1(avgage, Linf, K, t0)
+#    data: hadvbgf
+#    Linf       K      t0 
+# 77.0277  0.3601 -0.3546 
+#  residual sum-of-squares: 15637500
+# 
+# Number of iterations to convergence: 11 
+# Achieved convergence tolerance: 4.187e-06
+
+  redvbgf <- svlenage %>%
+    filter(species == "Redfish") 
+    #group_by(era) %>%
+    nls(length~vb1(avgage, Linf, K, t0), data = redvbgf,
+                  start = vbStarts(length~avgage,data = redvbgf))
+#all years
+# Nonlinear regression model
+#   model: length ~ vb1(avgage, Linf, K, t0)
+#    data: redvbgf
+#    Linf       K      t0 
+# 40.8016  0.1165 -1.6688 
+#  residual sum-of-squares: 3197990
+# 
+# Number of iterations to convergence: 13 
+# Achieved convergence tolerance: 8.316e-06
+
+    #dplyr::group_by(species, era) %>%
+    #group_map(try(~nls(length~vb1(avgage, Linf, K, t0), data = .,
+    #              start = vbStarts(length~avgage,data = .))))
 

### 2021-03-26 docs/SimData.Rmd
some vbgf fits work but many are strange check data

@@
 
-    #dplyr::group_by(species, era) %>%
-    #group_map(try(~nls(length~vb1(avgage, Linf, K, t0), data = .,
-    #              start = vbStarts(length~avgage,data = .))))
+    saithe1 <- svlenage %>%
+      filter(species=="Saithe",
+             era==1)
+
+    nls(length~vb1(avgage, Linf, K, t0), data = saithe1,
+                  start = vbStarts(length~avgage,data = saithe1))
+#Nonlinear regression model
+#   model: length ~ vb1(avgage, Linf, K, t0)
+#    data: saithe1
+#      Linf         K        t0 
+# 388.18797   0.01975  -3.94798 
+#  residual sum-of-squares: 5691213
+# 
+# Number of iterations to convergence: 5 
+# Achieved convergence tolerance: 5.511e-06
+
+ svvbgf <- svlenage %>%   
+    dplyr::group_by(species) %>%
+    group_map(~try(nls(length~vb1(avgage, Linf, K, t0), data = .,
+                  start = vbStarts(length~avgage,data = .))))
+
+ svvbgf2 <- svlenage %>%   
+    dplyr::group_by(species) %>%
+    group_modify(~broom::tidy(possibly(nls(length~vb1(avgage, Linf, K, t0), data = .x,
+                  start = vbStarts(length~avgage,data = .x)), otherwise = NA_real_)))
+ svvbgf3 <- svlenage %>%   
+    dplyr::group_by(species, era) %>%
+    group_map(~try(nls(length~vb1(avgage, Linf, K, t0), data = .,
+                  start = vbStarts(length~avgage,data = .))))
 

### 2021-03-30 docs/SimData.Rmd
try different start values vbgf

--- 2021-03-25 docs/SimData.Rmd
@@
 
-  svvbgf <- svlenage %>%
-    dplyr::group_by(species, era) %>%
-    group_map(~nls(length~vb1(avgage, Linf, K, t0), data = .,
-                  start = vbStarts(length~avgage,data = .)))
+  saveRDS(svlenage, file.path(d.name, paste0(scenario.name, "_",
+                                                       s, "agelensamp.rds")))
+  #plot len-age data and start values, some are bad
+  svlenage_pl <- svlenage %>%
+    group_by(species, era) %>%
+    group_map(~vbStarts(length~avgage,data=cur_group(),plot=TRUE))
 
@@
 # need 2-par maturity curve based on length, atlantis maturity ogive based on age group
+
+```
+
+```{r vbfits-scratch, eval=FALSE}
+
+  capelinvbgf <- svlenage %>%
+    filter(species == "Capelin") %>%
+    group_by(era) %>%
+    group_map(try(~nls(length~vb1(avgage, Linf, K, t0), data = .,
+                  start = vbStarts(length~avgage,data = .))))
+
+  capelinall <- svlenage %>% filter(species == "Capelin")
+  plot(length~avgage,data=capelinall,pch=19,col=rgb(0,0,0,1/5))
+  curve(vb1(x,Linf=coef(capelinvbgf[[1]])),from=0,to=5,col="red",lwd=2,add=TRUE)
+  curve(vb1(x,Linf=coef(capelinvbgf[[2]])),from=0,to=5,col="blue",lwd=2,add=TRUE)
+  curve(vb1(x,Linf=coef(capelinvbgf[[3]])),from=0,to=5,col="green",lwd=2,add=TRUE)
+
+  # herringvbgf <- svlenage %>%
+  #   filter(species == "Norweigan_ssh", era==1) %>%
+  #   group_map(try(~nls(length~vb1(avgage, Linf, K, t0), data = .,
+  #                 start = vbStarts(length~avgage,data = .))))
+
+#Error in nls(length ~ vb1(avgage, Linf, K, t0), data = ., start = vbStarts(length ~  : 
+  #singular gradient with group_map, era 2 and 3
+  codvbgf <- svlenage %>%
+    filter(species == "North_atl_cod") 
+    #group_by(era) %>%
+    codfit <- nls(length~vb1(avgage, Linf, K, t0), data = codvbgf,
+                  start = vbStarts(length~avgage,data = codvbgf))
+
+    plot(length~avgage,data=codvbgf,pch=19,col=rgb(0,0,0,1/5))
+    curve(vb1(x,Linf=coef(codfit)),from=0,to=20,col="red",lwd=2,add=TRUE)
+#  era==1   Nonlinear regression model
+#   model: length ~ vb1(avgage, Linf, K, t0)
+#    data: codvbgf
+#     Linf        K       t0 
+# 104.7120   0.4260   0.2176 
+#  residual sum-of-squares: 4109723
+# 
+# Number of iterations to convergence: 7 
+# Achieved convergence tolerance: 2.169e-06
+
+#  all years Nonlinear regression model
+#   model: length ~ vb1(avgage, Linf, K, t0)
+#    data: codvbgf
+#     Linf        K       t0 
+# 112.4694   0.3912   0.2097 
+#  residual sum-of-squares: 7354906
+# 
+# Number of iterations to convergence: 11 
+# Achieved convergence tolerance: 9.964e-06
+
+  hadvbgf <- svlenage %>%
+    filter(species == "Haddock") 
+    #group_by(era) %>%
+  hadfit <- nls(length~vb1(avgage, Linf, K, t0), data = hadvbgf,
+                  start = vbStarts(length~avgage,data = hadvbgf))
+
+  plot(length~avgage,data=hadvbgf,pch=19,col=rgb(0,0,0,1/5))
+  curve(vb1(x,Linf=coef(hadfit)),from=0,to=20,col="red",lwd=2,add=TRUE)
+
+# all years
+#     Nonlinear regression model
+#   model: length ~ vb1(avgage, Linf, K, t0)
+#    data: hadvbgf
+#    Linf       K      t0 
+# 77.0277  0.3601 -0.3546 
+#  residual sum-of-squares: 15637500
+# 
+# Number of iterations to convergence: 11 
+# Achieved convergence tolerance: 4.187e-06
+
+  redvbgf <- svlenage %>%
+    filter(species == "Redfish") 
+    #group_by(era) %>%
+
+  redfit <- nls(length~vb1(avgage, Linf, K, t0), data = redvbgf,
+                  start = vbStarts(length~avgage,data = redvbgf))
+
+  plot(length~avgage,data=redvbgf,pch=19,col=rgb(0,0,0,1/5))
+  curve(vb1(x,Linf=coef(redfit)),from=0,to=40,col="red",lwd=2,add=TRUE)
+
+#all years
+# Nonlinear regression model
+#   model: length ~ vb1(avgage, Linf, K, t0)
+#    data: redvbgf
+#    Linf       K      t0 
+# 40.8016  0.1165 -1.6688 
+#  residual sum-of-squares: 3197990
+# 
+# Number of iterations to convergence: 13 
+# Achieved convergence tolerance: 8.316e-06
+
+    saithe1 <- svlenage %>%
+      filter(species=="Saithe",
+             era==1)
+
+    saithe1fit <- nls(length~vb1(avgage, Linf, K, t0), data = saithe1,
+                  start = vbStarts(length~avgage,data = saithe1))
+
+    plot(length~avgage,data=saithe1,pch=19,col=rgb(0,0,0,1/5))
+    curve(vb1(x,Linf=coef(saithe1fit)),from=0,to=20,col="red",lwd=2,add=TRUE)
+#Nonlinear regression model
+#   model: length ~ vb1(avgage, Linf, K, t0)
+#    data: saithe1
+#      Linf         K        t0 
+# 388.18797   0.01975  -3.94798 
+#  residual sum-of-squares: 5691213
+# 
+# Number of iterations to convergence: 5 
+# Achieved convergence tolerance: 5.511e-06
+
+### full time series fits with different start points
+
+ svvbgf <- svlenage %>%   
+    dplyr::group_by(species) %>%
+    group_map(~try(nls(length~vb1(avgage, Linf, K, t0), data = .,
+                  start = vbStarts(length~avgage,data = .))))
+
+ spk <- svlenage %>%
+   group_by(species) %>%
+   group_keys()
+
+vballout <- data.frame() 
+for(i in 1:length(svvbgf)) {
+  if(is.list(svvbgf[[i]])) {
+    sp <- cbind(as.data.frame(spk)[i,], t(coef(svvbgf[[i]])))
+    vballout <- rbind(vballout, sp)
+  }
+}
+
+svvbgf.1 <- svlenage %>%   
+    dplyr::group_by(species) %>%
+    group_map(~try(nls(length~vb1(avgage, Linf, K, t0), data = .,
+                  start = vbStarts(length~avgage,data = .,methLinf="oldAge"))))
+
+vballout.1 <- data.frame() 
+for(i in 1:length(svvbgf.1)) {
+  if(is.list(svvbgf.1[[i]])) {
+    sp <- cbind(as.data.frame(spk)[i,], t(coef(svvbgf.1[[i]])))
+    vballout.1 <- rbind(vballout.1, sp)
+  }
+}
+
+svvbgf.2 <- svlenage %>%   
+    dplyr::group_by(species) %>%
+    group_map(~try(nls(length~vb1(avgage, Linf, K, t0), data = .,
+                  start = vbStarts(length~avgage,data = .,methLinf="longFish"))))
+
+vballout.2 <- data.frame() 
+for(i in 1:length(svvbgf.2)) {
+  if(is.list(svvbgf.2[[i]])) {
+    sp <- cbind(as.data.frame(spk)[i,], t(coef(svvbgf.2[[i]])))
+    vballout.2 <- rbind(vballout.2, sp)
+  }
+}
+
+################## by era
+
+ gk <- svlenage %>%
+   group_by(species, era) %>%
+   group_keys()
+
+ svvbgf3 <- svlenage %>%   
+    dplyr::group_by(species, era) %>%
+    group_map(~try(nls(length~vb1(avgage, Linf, K, t0), data = .,
+                  start = vbStarts(length~avgage,data = .))))
+
+vbout <- data.frame() 
+for(i in 1:length(svvbgf3)) {
+  if(is.list(svvbgf3[[i]])) {
+    spe <- cbind(as.data.frame(gk)[i,], t(coef(svvbgf3[[i]])))
+    vbout <- rbind(vbout, spe)
+  }
+}
+
+# try different start points to see if we can get more species fit
+ svvbgf3.1 <- svlenage %>%   
+    dplyr::group_by(species, era) %>%
+    group_map(~try(nls(length~vb1(avgage, Linf, K, t0), data = .,
+                  start = vbStarts(length~avgage,data = .,methLinf="oldAge"))))
+
+vbout.1 <- data.frame() 
+for(i in 1:length(svvbgf3.1)) {
+  if(is.list(svvbgf3.1[[i]])) {
+    spe <- cbind(as.data.frame(gk)[i,], t(coef(svvbgf3.1[[i]])))
+    vbout.1 <- rbind(vbout.1, spe)
+  }
+}
+
+# try different start points to see if we can get more species fit
+# this one goes to 11!
+svvbgf3.2 <- svlenage %>%   
+    dplyr::group_by(species, era) %>%
+    group_map(~try(nls(length~vb1(avgage, Linf, K, t0), data = .,
+                  start = vbStarts(length~avgage,data = .,methLinf="longFish"))))
+
+vbout.2 <- data.frame() 
+for(i in 1:length(svvbgf3.2)) {
+  if(is.list(svvbgf3.2[[i]])) {
+    spe <- cbind(as.data.frame(gk)[i,], t(coef(svvbgf3.2[[i]])))
+    vbout.2 <- rbind(vbout.2, spe)
+  }
+}
 

### 2021-03-30 docs/SimData.Rmd
almost writing growth pars csv

@@
 
-Generating length composition input data requires mapping numbers at length in the atlantisom 1 cm bin length output to our selected length bins, then summing the numbers at length in those bins for all species.
-The input required is a 4d array, indexed by survey, species, year, and sizebin.
-So each survey can be a separate csv with #species at the top of year rows and sizebin columns with n at length the values, scrolling down for each species.
-
 ```{r fitpars, eval=FALSE}
@@
                                                        s, "agelensamp.rds")))
-  #plot len-age data and start values, some are bad
-  svlenage_pl <- svlenage %>%
+
+  # try different start points to see if we can get more species fit
+  # this one goes to 11!
+  # need because group_map doesn't keep group names
+  gk <- svlenage %>%
     group_by(species, era) %>%
-    group_map(~vbStarts(length~avgage,data=cur_group(),plot=TRUE))
-
-}
-
+    group_keys()
+
+  svvbgf3.2 <- svlenage %>%   
+    dplyr::group_by(species, era) %>%
+    group_map(~try(nls(length~vb1(avgage, Linf, K, t0), data = .,
+                       start = vbStarts(length~avgage,data = .,methLinf="longFish"))))
+
+  #this gets exponential growth pars, probably won't use this growth mod though
+  svexp3 <- svlenage %>%
+    dplyr::group_by(species, era) %>%
+    group_modify(~ broom::tidy(lm(log(length) ~ log(avgage), data = .x))) %>%
+    select(species, era, term, estimate) %>%
+    pivot_wider(names_from = "term", values_from = "estimate") %>%
+    mutate(psi = exp(`(Intercept)`)) %>%
+    rename(kappa = `log(avgage)`) %>%
+    select(species, era, kappa, psi)
+
+  vbout.2 <- data.frame() 
+  for(i in 1:length(svvbgf3.2)) {
+    if(is.list(svvbgf3.2[[i]])) {
+      spe <- cbind(as.data.frame(gk)[i,], as.data.frame(t(coef(svvbgf3.2[[i]]))))
+      vbout.2 <- rbind(vbout.2, spe)
+    }
+  }
+
+  growthpars <- left_join(svexp3, vbout.2)
+
+  maxera <- growthpars %>% 
+    filter(!is.na(Linf)) %>% 
+    group_by(era) %>% 
+    tally() %>%
+    filter(n==11) %>%
+    select(era)
+
+  growthspp <- growthpars %>%
+    filter(era==maxera) %>%
+    select(species, psi, kappa, Linf, K) %>%
+    t() %>%
+    as.data.frame()
+
+  write_csv(paste0(o.name,"growth_species_NOBA_",s,".csv"))
+}
+```
+
+Generating length composition input data requires mapping numbers at length in the atlantisom 1 cm bin length output to our selected length bins, then summing the numbers at length in those bins for all species.
+The input required is a 4d array, indexed by survey, species, year, and sizebin.
+So each survey can be a separate csv with #species at the top of year rows and sizebin columns with n at length the values, scrolling down for each species.
+
+Arrange lengths into appropriate bins:
+
+```{r lendat}
 # rearrange into length bins and write length comp data csvs
@@
 # Achieved convergence tolerance: 5.511e-06
+
+#plot len-age data and start values, some are bad
+  svlenage_pl <- svlenage %>%
+    group_by(species, era) %>%
+    group_map(~vbStarts(length~avgage,data=cur_group(),plot=TRUE))
 

### 2021-03-31 docs/SimData.Rmd
update sim data 1st of 4 survey growth pars

@@
 ### Write to model input files
+
+#### Multispecies surplus production model (MSSPM)
 
@@
 ```
+
+#### Length-structured multispecies model (Hydra)
 
@@
 
-Need to fit growth and maturity parameters from Atlantis output for input into hydra. Fit growth using code below.
+Need to fit growth and maturity parameters from Atlantis output for input into hydra.
+Fit growth using code below, based on the package FSA [@ogle-fsa-2021].
+The package provides multiple ways to get starting values for the R `nls` function for fitting nonlinear models.
+It was necessary to try several different methods for starting values to get at least one von Bertalannfy growth parameter set for each species.
+A scratch code block shows some of the experiments to get the fitting working that resulted in the function to fit for all surveys.
 
@@
 
-  growthpars <- left_join(svexp3, vbout.2)
+  growthpars <- left_join(svexp3, vbout.2) %>%
+    mutate(growthType = 4) # use only vonB for now, can change later
 
@@
     filter(era==maxera) %>%
-    select(species, psi, kappa, Linf, K) %>%
-    t() %>%
-    as.data.frame()
-
-  write_csv(paste0(o.name,"growth_species_NOBA_",s,".csv"))
-}
-```
-
-Generating length composition input data requires mapping numbers at length in the atlantisom 1 cm bin length output to our selected length bins, then summing the numbers at length in those bins for all species.
-The input required is a 4d array, indexed by survey, species, year, and sizebin.
-So each survey can be a separate csv with #species at the top of year rows and sizebin columns with n at length the values, scrolling down for each species.
-
-Arrange lengths into appropriate bins:
-
-```{r lendat}
-# rearrange into length bins and write length comp data csvs
-#multiple surveys named in list object, use as filename
-for(s in names(len_comp_data)){
-  #arrange into long format: year, Species, agecl, length 
-  svlenbin <- len_comp_data[[s]][[1]] %>%
-    dplyr::mutate(year = ceiling(time/stepperyr)) %>%
-    dplyr::select(species, agecl, year, upper.bins, atoutput) %>%
-
-    tidyr::spread(lenbins, atoutput) %>%
-    write_csv(paste0(o.name,"observation_lengths_NOBA_",s,".csv"))
-}
-
-# need 2-par maturity curve based on length, atlantis maturity ogive based on age group
-
-```
+    select(species, psi, kappa, Linf, K, growthType) %>%
+    pivot_longer(psi:growthType, names_to = "par") %>%
+    pivot_wider(names_from = species, 
+                values_from = value) %>%
+    ungroup() %>%
+    select(-era) %>%
+    remove_rownames %>% column_to_rownames(var="par") %>%
+    write.csv(paste0(o.name,"growth_species_NOBA_era",maxera,"_",s,".csv"))
+}
+
+```
+
+This code block has experiments with different fits and different ways to splice the length at age data.
 
@@
 
+Generating length composition input data requires mapping numbers at length in the atlantisom 1 cm bin length output to our selected length bins, then summing the numbers at length in those bins for all species.
+The input required is a 4d array, indexed by survey, species, year, and sizebin.
+So each survey can be a separate csv with #species at the top of year rows and sizebin columns with n at length the values, scrolling down for each species.
+
+Arrange lengths into appropriate bins:
+
+```{r lendat, eval=FALSE}
+# rearrange into length bins and write length comp data csvs
+#multiple surveys named in list object, use as filename
+for(s in names(len_comp_data)){
+  #arrange into long format: year, Species, agecl, length 
+  svlenbin <- len_comp_data[[s]][[1]] %>%
+    dplyr::mutate(year = ceiling(time/stepperyr)) %>%
+    dplyr::select(species, agecl, year, upper.bins, atoutput) %>%
+
+    tidyr::spread(lenbins, atoutput) %>%
+    write_csv(paste0(o.name,"observation_lengths_NOBA_",s,".csv"))
+}
+
+# need 2-par maturity curve based on length, atlantis maturity ogive based on age group
+
+```
+
 Stomach weights are needed as input to hydra.
@@
 
-```{r stomwt}
+```{r stomwt, eval=FALSE}
 
@@
 We can get surface and bottom temp from Atlantis using the layers from each polygon, with layer 0 being the bottom and the maximum layer in the polygon being the surface (according to the [Atlantis wiki/Atlantis output files/Atlantis NetCDF structure](https://confluence.csiro.au/display/Atlantis/Atlantis+NetCDF+Structure).
+
 ```{r surveytemp, eval=FALSE}
@@
 
-## Generating a survey dataset
+#### Age structured multispecies statistical catch at age model (MSSCAA)
+
+## References

### 2021-03-31 docs/SimData.Rmd
write grow par files by era

@@
 
-  maxera <- growthpars %>% 
-    filter(!is.na(Linf)) %>% 
-    group_by(era) %>% 
-    tally() %>%
-    filter(n==11) %>%
-    select(era)
+  # sadly doesn't work so just make a csv for each era and use the one with most spp
+  # maxera <- growthpars %>% 
+  #   filter(!is.na(Linf)) %>% 
+  #   group_by(era) %>% 
+  #   tally() %>%
+  #   filter(n==1) %>%
+  #   select(era)
 
   growthspp <- growthpars %>%
-    filter(era==maxera) %>%
     select(species, psi, kappa, Linf, K, growthType) %>%
@@
     pivot_wider(names_from = species, 
-                values_from = value) %>%
-    ungroup() %>%
+                values_from = value)
+
+  growthspp.era <- split(growthspp, list(growthspp$era))
+
+  for(era in names(growthspp.era)) {
+    gpar <- growthspp.era[[era]] %>%
+      ungroup()%>%
       select(-era) %>%
-    remove_rownames %>% column_to_rownames(var="par") %>%
-    write.csv(paste0(o.name,"growth_species_NOBA_era",maxera,"_",s,".csv"))
+      remove_rownames() %>% 
+      column_to_rownames(var="par") %>%
+      write.csv(paste0(o.name,"growth_species_NOBA_era",era,"_",s,".csv"))
+
+  }
+
 }

### 2021-04-01 docs/SimData.Rmd
update growth file output

@@
                                                        s, "agelensamp.rds")))
+}
+
+  # read saved sample from above
+  survs <- list.files(path=d.name, pattern = "agelensamp.rds", full.names = TRUE)
+
+  survname <-  str_match(survs, paste0(scenario.name,"_\\s*(.*?)\\s","*agelensamp.rds"))[,2]
+
+  agelensamp_data <- lapply(survs, readRDS)
+
+  names(agelensamp_data) <- survname
+
+for(s in names(agelensamp_data)){
 
@@
   growthpars <- left_join(svexp3, vbout.2) %>%
-    mutate(growthType = 4) # use only vonB for now, can change later
+    mutate(growthType = case_when(
+      is.na(Linf) ~ 2,
+      TRUE ~ 4)) # vonB default, exponential if no vonB
 

### 2021-04-06 docs/SimData.Rmd
error on reading back in?

--- 2021-03-31 docs/SimData.Rmd
@@
                                                        s, "agelensamp.rds")))
+# }
+# 
+#   # read saved sample from above
+#   survs <- list.files(path=d.name, pattern = "agelensamp.rds", full.names = TRUE)
+# 
+#   survname <-  str_match(survs, paste0(scenario.name,"_\\s*(.*?)\\s","*agelensamp.rds"))[,2]
+# 
+#   agelensamp_data <- lapply(survs, readRDS)
+# 
+#   names(agelensamp_data) <- survname
+# 
+# 
+# for(s in names(agelensamp_data)){
 

### 2021-04-06 docs/GBFleetDefinitions.Rmd
url typo for FleetDecisions

@@
 
-## Bottom trawls (small/large mesh)
+## Bottom trawls (small/large mesh) {.tabset .tabset-pills}
 
@@
 Since the change in regulations dictate a change in the composition of species caught over time, this gear type needs to be investigated separately.
+The data base holds information based on the mesh size of the otter trawl nets for NEGEAR code = 5 (from 1982 onward).
+
+Below are plots of species caught in the otter trawl (making up 96% of the total catch from 1964 to 2020) by mesh size (in inches).
+Note that the figure begins in 1982.
+Any trip without a mesh size attributed is not displayed, this includes all trips prior to 1982.
+A lack of mesh size accounts for approximately 20% of landing post 1982 and approximately 50% of all landings
+
+### plotted on same scale
+```{r fixed,  echo=F, eval =T}
+knitr::include_graphics("figures/demersalfixed.png")
+```
+
+### plotted on different scale
+```{r free,  echo=F, eval =T}
+knitr::include_graphics("figures/demersalfree_y.png")
+```
 
@@
 
-This exploratory study resulted in the following designation of [Fleets](https://noaa-edab.github.io/ms-keyrun/GBFleetDecision.html)
+This exploratory study resulted in the following designation of [Fleets](https://noaa-edab.github.io/ms-keyrun/GBFleetDecisions.html)
 

### 2021-04-06 docs/SimData.Rmd
initial length comps in model bins

@@
 
+Note that spring surveys don't get mackerel at all (not in the model domain then).
+A full set of von B parameters could only be estimated for the first era of both versions of the fall survey tested here.
+
 ```{r fitpars, eval=FALSE}
@@
 # rearrange into length bins and write length comp data csvs
+# modbins from sizebins above, differ by species so need modbin.min and modbin.max
+# pcumsum is an FSA function
+
+modbins <- sizebins %>%
+  rename(species = `commentField-notused`) %>%
+  pivot_longer(!species, names_to = "sizebin", values_to = "binwidth") %>%
+  group_by(species) %>%
+  arrange(sizebin) %>%
+  mutate(modbin.min = pcumsum(binwidth),
+         modbin.max = cumsum(binwidth)) %>%
+  arrange(species, sizebin)
+
 #multiple surveys named in list object, use as filename
@@
     dplyr::select(species, agecl, year, upper.bins, atoutput) %>%
-
-    tidyr::spread(lenbins, atoutput) %>%
+    dplyr::left_join(modbins) %>%
+    dplyr::filter(modbin.min < upper.bins & upper.bins <= modbin.max) %>%
+    dplyr::group_by(species, year, sizebin) %>%
+    dplyr::summarise(sumlen = sum(atoutput)) %>%
+    tidyr::spread(sizebin, sumlen) %>%
     write_csv(paste0(o.name,"observation_lengths_NOBA_",s,".csv"))

### 2021-04-07 docs/SimData.Rmd
start binwidth automation

@@
 
-*Consider a function to do this using species max size, the vonB k and the desired number of bins for more consistency and to easily change Nsizebins.*
-```{r lbinfunction, eval=FALSE}
-# see e.g., https://www.sciencedirect.com/science/article/abs/pii/S0165783619303297
-# estimated optimal bin size for VBGF estimation is 0.23 x Lmax ^0.6
-
-binwidths <- function(Lmax, vonBk, Nsizebins){
-  #optimal bin size is lower end
-  binopt <- 0.23*Lmax^0.6
-  #vonBk tells you how many fine bins for fast growth
-
-  #aggregate bins in multiples of optimal for bigger fish/slower growth to get Nsizebins
-
-}
-
-```
-
 Initially, we'll directly transfer bin widths for GB and NOBA common species:
@@
 
-# need 2-par maturity curve based on length, atlantis maturity ogive based on age group
+```
+
+The first pass through with the current bins results in very few lengths in the smallest bin for most species.
+This suggests we need to change the binwidths and possibly the number of bins to optimize length fitting.
+Perhaps a larger first bin or don't start it at 0?
+
+*Consider a function to do this using species max size, the vonB k and the desired number of bins for more consistency and to easily change Nsizebins.* Do we need t0 as well? Haven't saved it.
+
+```{r lbinfunction, eval=FALSE}
+# see e.g., https://www.sciencedirect.com/science/article/abs/pii/S0165783619303297
+# estimated optimal bin size for VBGF estimation is 0.23 x Lmax ^0.6
+
+# see Ogle citing Schnute and Fornier on K as a growth increment
+# https://derekogle.com/NCNRS349/modules/Growth/BKG
+# K exponential rate of approach to asymptotic size
+# exp(-K) fraction by which growth increment is multiplied each year
+# at what length is increment*exp(-K) small? group those lengths into bigger bins
+
+# pull observed min and max size from data to optimize start and end bins
+minmaxlen <- list()
+for(s in names(len_comp_data)){
+  minmaxlen[[s]] <- len_comp_data[[s]][[1]] %>%
+    dplyr::group_by(species) %>%
+    dplyr::summarise(minlen = min(lower.bins, na.rm = TRUE),
+                     maxlen = max(upper.bins, na.rm = TRUE)) 
+}
+
+# note!
+# Green_halibut and Saithe have maxlen at 200 in at least one survey, 
+# may need higher maxlength when running atlantisom comps wrapper, fix survey config files
+
+binwidths <- function(Lmax, vonBk, Nsizebins){
+  #optimal bin size is lower end
+  binopt <- 0.23*Lmax^0.6
+  #vonBk tells you how many fine bins for fast growth
+
+  #aggregate bins in multiples of optimal for bigger fish/slower growth to get Nsizebins
+
+}
 
@@
 ```{r}
+
+# need 2-par maturity curve based on length, atlantis maturity ogive based on age group
+# or just translate age at 50% maturity to length and call it good?
 

### 2021-04-07 docs/SimData.Rmd
all libraries at top

@@
 library(ggthemes)
+library(FSA)
 
@@
 
-library(FSA)
 vb1 <- vbFuns()

### 2021-04-08 docs/SimData.Rmd
add vonBK pars

@@
 minmaxlen <- list()
+# read in saved vonB pars--see if seasons are different and will need a replacement if K wasn't estimated
+era <- 1
+vonBK <- list()
+minmaxvonBK <- list()
 for(s in names(len_comp_data)){
-  minmaxlen[[s]] <- len_comp_data[[s]][[1]] %>%
+  minmaxlendf <- len_comp_data[[s]][[1]] %>%
     dplyr::group_by(species) %>%
@@
                      maxlen = max(upper.bins, na.rm = TRUE))
+
+  vonBKdf <- read.csv(paste0(o.name,"growth_species_NOBA_era",era,"_",s,".csv")) %>%
+    rename(par = X) %>%
+    pivot_longer(!par, names_to = "species") %>%
+    pivot_wider(names_from = par, 
+                values_from = value) %>%
+    select(species, K)
+
+  minmaxvonBKdf <- left_join(minmaxlendf,vonBKdf)
+
+  minmaxlen[[s]] <- minmaxlendf
+  vonBK[[s]] <- vonBKdf
+  minmaxvonBK[[s]] <- minmaxvonBKdf
+
 }
@@
 
-binwidths <- function(Lmax, vonBk, Nsizebins){
+binwidths <- function(Lmin, Lmax, vonBk, Nsizebins){
   #optimal bin size is lower end

### 2021-04-08 docs/SimData.Rmd
more length stuff

--- 2021-04-07 docs/SimData.Rmd
@@
 minmaxlen <- list()
+# read in saved vonB pars--see if seasons are different and will need a replacement if K wasn't estimated
+#era <- 1
+vonBK <- list()
+minmaxvonBK <- list()
 for(s in names(len_comp_data)){
-  minmaxlen[[s]] <- len_comp_data[[s]][[1]] %>%
+  minmaxlendf <- len_comp_data[[s]][[1]] %>%
     dplyr::group_by(species) %>%
     dplyr::summarise(minlen = min(lower.bins, na.rm = TRUE),
-                     maxlen = max(upper.bins, na.rm = TRUE)) 
+                     maxlen = max(upper.bins, na.rm = TRUE),
+                     range = maxlen-minlen)
+
+  for(era in 1:3){
+  vonBKer <- read.csv(paste0(o.name,"growth_species_NOBA_era",era,"_",s,".csv")) %>%
+    rename(par = X) %>%
+    pivot_longer(!par, names_to = "species") %>%
+    pivot_wider(names_from = par,
+                names_prefix = era,
+                values_from = value) %>%
+    select(species, paste0(era,"K"))
+
+  ifelse(era==1,
+         vonBKdf <- vonBKer,
+         vonBKdf <- left_join(vonBKdf, vonBKer))
+  }
+
+  minmaxvonBKdf <- left_join(minmaxlendf,vonBKdf)
+
+  minmaxlen[[s]] <- minmaxlendf
+  vonBK[[s]] <- vonBKdf
+  minmaxvonBK[[s]] <- minmaxvonBKdf
+
 }
@@
   #vonBk tells you how many fine bins for fast growth
+  #low vonBK is doesn't asympote, divide length range equally into nbins?
+
+  #high vonBK asympotes quickly,
 

### 2021-04-09 docs/SimData.Rmd
binwidth function dev

@@
   binopt <- 0.23*Lmax^0.6
+
+  #range gives initial default binwidth, all equal (range/Nbins)
+  bindef <- (Lmax-Lmin)/Nsizebins
+
   #vonBk tells you how many fine bins for fast growth
+  # threshold cumprod of 0.05 to define last bin? tiny growth increment
+  # use binopt up to this threshold?
+  # plot(cumprod(rep(1, 160)*exp(-0.0386)), ylim=c(0,1), xlim=c(0,180)) #saithe
+  # lines(cumprod(rep(1, 20)*exp(-1.4)), ylim=c(0,1), xlim=c(0,180)) #capelin
+  # lines(cumprod(rep(1, 40)*exp(-0.163)), ylim=c(0,1), xlim=c(0,180)) #redfish
+  # lines(cumprod(rep(1, 130)*exp(-0.344)), ylim=c(0,1), xlim=c(0,180)) #cod
   #low vonBK is doesn't asympote, divide length range equally into nbins?

### 2021-04-09 docs/SimData.Rmd
id threshold lengths based on vonBk

@@
   # lines(cumprod(rep(1, 130)*exp(-0.344)), ylim=c(0,1), xlim=c(0,180)) #cod
+
+  thresh <- 0.05
+  which(cumprod(rep(1, Lmax)*exp(-vonBk)) < thresh) #gives lengths of fast growth
+
   #low vonBK is doesn't asympote, divide length range equally into nbins?

### 2021-04-09 docs/SimData.Rmd
maybe equal bins are best?

@@
   #range gives initial default binwidth, all equal (range/Nbins)
+  #or should it be max/nbins?
+  #can use if no K estimated
   bindef <- (Lmax-Lmin)/Nsizebins
+
+  #minmaxvonBK$BTS_fall_allbox_effic1 %>% group_by(species) %>% mutate(bindef = (maxlen)/5)
 
@@
   thresh <- 0.05
-  which(cumprod(rep(1, Lmax)*exp(-vonBk)) < thresh) #gives lengths of fast growth
+  which(cumprod(rep(1, Lmax)*exp(-vonBk)) > thresh) #gives 1 cm increments of fast growth?
+
+  #
 

### 2021-04-09 docs/SimData.Rmd
almost a function

@@
 
-binwidths <- function(Lmin, Lmax, vonBk, Nsizebins){
+estbinwidths <- function(Lmax, vonBk, Nsizebins, thresh = 0.05){
   #optimal bin size is lower end
@@
   #can use if no K estimated
-  bindef <- (Lmax-Lmin)/Nsizebins
-
-  #minmaxvonBK$BTS_fall_allbox_effic1 %>% group_by(species) %>% mutate(bindef = (maxlen)/5)
+  bindef <- max(1,ceiling(Lmax/Nsizebins))
+
+  #minmaxvonBK$BTS_fall_allbox_effic1 %>% group_by(species) %>% mutate(bindef = ceiling(maxlen/5))
 
   #vonBk tells you how many fine bins for fast growth
-  # threshold cumprod of 0.05 to define last bin? tiny growth increment
+  #low vonBK is doesn't asympote, divide length range equally into nbins?
+  #high vonBK asympotes quickly, 
+  # threshold cumprod of 0.05 or 0.1 to define last bin? tiny growth increment
   # use binopt up to this threshold? 
@@
 
-  thresh <- 0.05
-  which(cumprod(rep(1, Lmax)*exp(-vonBk)) > thresh) #gives 1 cm increments of fast growth?
-
-  #
-
-  #low vonBK is doesn't asympote, divide length range equally into nbins?
-
-  #high vonBK asympotes quickly,
+  lenthresh <- max(which(cumprod(rep(1, Lmax)*exp(-vonBk)) > thresh)) #gives 1 cm increments of fast growth
+
+  #put all lengths above this into one big bin, 
+  binwidth.max <- Lmax-lenthresh
+
+  #divide lengths below this by thresholds? for min 5 bins, use .5, .3, .2, .1, .05?
+  # or divide equally into nbins-1 bins? equal may be best
+  # hightest growth bin will be very small and we won't have fish in it
+  # dont go smaller than binopt
+  binwidth.gro <- max(ceiling(binopt), ceiling(lenthresh/(Nsizebins-1)))
 
   #aggregate bins in multiples of optimal for bigger fish/slower growth to get Nsizebins
+  ifelse(binwidth.max > bindef, #& binwidth.max < 0.5*Lmax,
+         binwidths <- c(rep(binwidth.gro, Nsizebins-1), binwidth.max),
+         binwidths <- rep(bindef, Nsizebins))
+
+  return(binwidths)
 

### 2021-04-13 docs/SimData.Rmd
new equal binwidths and lengths

@@
 
-*Consider a function to do this using species max size, the vonB k and the desired number of bins for more consistency and to easily change Nsizebins.* Do we need t0 as well? Haven't saved it.
+*Consider a function to do this using species max size, the vonB k and the desired number of bins for more consistency and to easily change Nsizebins.*
+
+This function is below, but with only 5 bins I don't likw how it works, allocating more smaller bins to the smallest sizes that we don't observe and aggregating many lengths into one large bin.
+Therefore the simplest approach is to have equal sized bins (max observed length/desired number of bins).
+We will try this for the length comps.
 
@@
   #aggregate bins in multiples of optimal for bigger fish/slower growth to get Nsizebins
-  ifelse(binwidth.max > bindef, #& binwidth.max < 0.5*Lmax,
+  ifelse(binwidth.max > bindef & binwidth.max < 0.5*Lmax,
          binwidths <- c(rep(binwidth.gro, Nsizebins-1), binwidth.max),
@@
 }
+
+equalbinwidths <- function(Lmax, Nsizebins){
+  bindef <- max(1,ceiling(Lmax/Nsizebins))
+  binwidths <- rep(bindef, Nsizebins)
+  return(binwidths)
+}
+
+# for fixed bins across surveys find min and max observed lengths
+
+maxLrange <- bind_rows(minmaxvonBK) %>%
+  group_by(species) %>%
+  summarise(minminL = min(minlen),
+            maxmaxL = max(maxlen))
+
+Nsizebins  <-  5
+
+newbins <- maxLrange %>%
+  group_by(species) %>%
+  group_modify(~broom::tidy(equalbinwidths(.$maxmaxL, Nsizebins))) %>%
+  mutate(lb = paste0("sizebin",seq(1:Nsizebins))) %>%
+  pivot_wider(names_from = lb, values_from = x)
+
+sizebins <- as.data.frame(newbins) %>%
+  rename('commentField-notused' = species) %>%
+  relocate('commentField-notused', .after = last_col())
+
+write_csv(sizebins, paste0(o.name,"length_sizebins_NOBA.csv"))
 

### 2021-04-13 docs/SimData.Rmd
WIP alternate length data format

@@
 
+Reran the lendat code block above to make the survey-specific length input files with the new equal sized bins.
+
+Now write input files in Gavin's format to have surveys together in one file:
+
+```{r altinput, eval=FALSE}
+
+modbins <- sizebins %>%
+  rename(species = `commentField-notused`) %>%
+  pivot_longer(!species, names_to = "sizebin", values_to = "binwidth") %>%
+  group_by(species) %>%
+  arrange(sizebin) %>%
+  mutate(modbin.min = pcumsum(binwidth),
+         modbin.max = cumsum(binwidth)) %>%
+  arrange(species, sizebin)
+
+#multiple surveys named in list object, use as filename
+for(s in names(len_comp_data)){
+  #arrange into long format: year, Species, agecl, length 
+  svlenbin <- len_comp_data[[s]][[1]] %>%
+    dplyr::mutate(year = ceiling(time/stepperyr)) %>%
+    dplyr::select(species, agecl, year, upper.bins, atoutput) %>%
+    dplyr::left_join(modbins) %>%
+    dplyr::filter(modbin.min < upper.bins & upper.bins <= modbin.max) %>%
+    dplyr::group_by(species, year, sizebin) %>%
+    dplyr::summarise(sumlen = sum(atoutput)) %>%
+    tidyr::spread(sizebin, sumlen) %>%
+    ungroup() %>%
+    dplyr::mutate(inpN = rowSums(.[,-(1:2)], na.rm = TRUE)) %>%
+    dplyr::mutate(type = 0) %>%
+    dplyr::mutate(survey = s) %>%
+    #dplyr::mutate(vars(matches("sizebin"))/inpN)
+
+}
+write_csv(paste0(o.name,"observation_lengths_NOBA_",s,".csv"))
+
+```
+
 Stomach weights are needed as input to hydra.

### 2021-04-29 docs/SimData.Rmd
length files in Gavin's format

--- 2021-04-13 docs/SimData.Rmd
@@
 
+Reran the lendat code block above to make the survey-specific length input files with the new equal sized bins.
+
+Now write input files in Gavin's format to have surveys together in one file:
+
+```{r altinput, eval=FALSE}
+
+modbins <- sizebins %>%
+  rename(species = `commentField-notused`) %>%
+  pivot_longer(!species, names_to = "sizebin", values_to = "binwidth") %>%
+  group_by(species) %>%
+  arrange(sizebin) %>%
+  mutate(modbin.min = pcumsum(binwidth),
+         modbin.max = cumsum(binwidth)) %>%
+  arrange(species, sizebin)
+
+allsvlenbin <- tibble()
+
+#multiple surveys named in list object, use as filename
+for(s in names(len_comp_data)){
+  #arrange into long format: year, Species, agecl, length 
+  svlenbin <- len_comp_data[[s]][[1]] %>%
+    dplyr::mutate(year = ceiling(time/stepperyr)) %>%
+    dplyr::select(species, agecl, year, upper.bins, atoutput) %>%
+    dplyr::left_join(modbins) %>%
+    dplyr::filter(modbin.min < upper.bins & upper.bins <= modbin.max) %>%
+    dplyr::group_by(species, year, sizebin) %>%
+    dplyr::summarise(sumlen = sum(atoutput)) %>%
+    tidyr::spread(sizebin, sumlen) %>%
+    ungroup() %>%
+    dplyr::mutate(inpN = rowSums(.[,-(1:2)], na.rm = TRUE)) %>%
+    dplyr::mutate(type = 0) %>%
+    dplyr::mutate(survey = s) %>%
+    dplyr::mutate_at(vars(contains("sizebin")), ~./inpN) %>%
+    dplyr::select(survey, year, species, type, inpN, everything())
+
+  allsvlenbin <- bind_rows(allsvlenbin, svlenbin)
+}
+write_csv(allsvlenbin, paste0(o.name,"observation_lengths_NOBA_allsurvs.csv"))
+
+```
+
 Stomach weights are needed as input to hydra.

### 2021-04-29 docs/SimData.Rmd
WIP survey and fishery dat in Gavin's format

@@
 
-This function is below, but with only 5 bins I don't likw how it works, allocating more smaller bins to the smallest sizes that we don't observe and aggregating many lengths into one large bin.
+This function is below, but with only 5 bins I don't like how it works, allocating more smaller bins to the smallest sizes that we don't observe and aggregating many lengths into one large bin.
 Therefore the simplest approach is to have equal sized bins (max observed length/desired number of bins).
@@
 
+Survey and catch datasets in Gavin's suggested format (all surveys together, all fleets together):
+
+```{r svcatchalt, eval=FALSE}
+#read in survey biomass data
+survObsBiom <- read_savedsurvs(d.name, 'survB') #reads in all surveys
+
+# get cvs from config files
+svcon <- list.files(path=here("/simulated-data/config/"), pattern = "*survey*", full.names = TRUE)
+fishcon <- list.files(path=here("/simulated-data/config/"), pattern = "*fishery*", full.names = TRUE)
+
+omlist_ss <- NOBAom_ms
+source(here("/simulated-data/config/omdimensions.R"))
+
+svcvlook <- tibble()
+for(c in 1:length(svcon)){
+  source(svcon[c])
+  surv_cv_n <- surv_cv %>% 
+    mutate(survey=survey.name)
+  svcvlook <- bind_rows(svcvlook, surv_cv_n)
+}
+
+fcvlook <- tibble()
+for(c in 1:length(fishcon)){
+  source(fishcon[c])
+  fish_cv_n <- fish_cv %>%
+    mutate(fishery=fishery.name)
+  fcvlook <- bind_rows(fcvlook, fish_cv_n)
+}
+
+allsvbio <- tibble()
+
+#multiple surveys named in list object, use as filename
+for(s in names(survObsBiom)){
+  #arrange into wide format: year, Species1, Species2 ... and write csv
+  svbio <- survObsBiom[[s]][[1]] %>%
+    dplyr::mutate(year = ceiling(time/stepperyr)) %>%
+    dplyr::select(species, year, atoutput) %>%
+    dplyr::rename(biomass = atoutput) %>%
+    dplyr::mutate(survey = s,
+                  CV = 0.2) %>% #placeholder, can get actual value used to generate from config files
+    dplyr::select(survey, year, species, everything()) %>%
+    dplyr::arrange(survey, species, year)
+
+  allsvbio <- bind_rows(allsvbio, svbio)
+}
+write_csv(allsvbio, paste0(o.name,"observation_biomass_NOBA_allsurvs.csv"))
+
+```
+
 Output temperature data from surveys using "Temp" variable in .nc file and the `atlantisom::load_nc_physics` function.

### 2021-04-30 docs/SimData.Rmd
add catch and catch-length comps in Gavin's format

@@
 
-Now write input files in Gavin's format to have surveys together in one file:
+Now write input files in Gavin's format to have surveys together in one file.
+Note that this is just a test.
+The folder has two different survey and two different fishery configurations; they are not independent.
+Also, fishery catch is not currently separated into fleets.
+We wouldn't use all of these in an input file.
+Also add fishery length comps:
 
@@
 
-#multiple surveys named in list object, use as filename
+#multiple surveys named in list object, add to file
 for(s in names(len_comp_data)){
@@
 write_csv(allsvlenbin, paste0(o.name,"observation_lengths_NOBA_allsurvs.csv"))
+
+#requires function below
+catchlen_ss <- read_savedfisheries(d.name, "catchLen")
+
+allfishlenbin <- tibble()
+for(f in names(catchlen_ss)){
+  #arrange into long format: year, Species, agecl, length 
+  fishlenbin <- catchlen_ss[[f]][[1]] %>%
+    dplyr::mutate(year = ceiling(time/stepperyr)) %>%
+    dplyr::select(species, agecl, year, upper.bins, atoutput) %>%
+    dplyr::left_join(modbins) %>%
+    dplyr::filter(modbin.min < upper.bins & upper.bins <= modbin.max) %>%
+    dplyr::group_by(species, year, sizebin) %>%
+    dplyr::summarise(sumlen = sum(atoutput)) %>%
+    tidyr::spread(sizebin, sumlen) %>%
+    ungroup() %>%
+    dplyr::mutate(inpN = rowSums(.[,-(1:2)], na.rm = TRUE)) %>%
+    dplyr::mutate(type = 0) %>%
+    dplyr::mutate(fishery = f,
+                  area = 1) %>%
+    dplyr::mutate_at(vars(contains("sizebin")), ~./inpN) %>%
+    dplyr::select(fishery, area, year, species, type, inpN, everything()) %>%
+    dplyr::arrange(fishery, area, species, year)
+
+  allfishlenbin <- bind_rows(allfishlenbin, fishlenbin)
+}
+write_csv(allfishlenbin, paste0(o.name,"observation_lengths_NOBA_allfisheries.csv"))
 
@@
 
-Survey and catch datasets in Gavin's suggested format (all surveys together, all fleets together):
+Survey and catch datasets in Gavin's suggested format (all surveys together, all fleets together).
+As above, note that this is just a test.
+The folder has two different survey and two different fishery configurations; they are not independent.
+Also, fishery catch is not currently separated into fleets.
+We wouldn't use all of these in an input file.
 
@@
     dplyr::rename(biomass = atoutput) %>%
-    dplyr::mutate(survey = s,
-                  CV = 0.2) %>% #placeholder, can get actual value used to generate from config files
+    dplyr::mutate(survey = s) %>%
+    dplyr::left_join(svcvlook) %>%
     dplyr::select(survey, year, species, everything()) %>%
@@
 
+allcatch <- tibble()
+
+for(f in names(catchbio_ss)){
+  catchbio <- catchbio_ss[[f]][[1]] %>%
+    dplyr::mutate(year = time/365) %>%
+    dplyr::select(species, year, atoutput) %>%
+    dplyr::rename(catch = atoutput) %>%
+    dplyr::mutate(fishery = f) %>%
+    dplyr::mutate(area = 1) %>%
+    dplyr::left_join(fcvlook) %>%
+    dplyr::select(fishery, area, year, species, everything()) %>%
+    dplyr::arrange(fishery, species, year)
+
+  allcatch <- bind_rows(allcatch, catchbio)  
+}
+
+write_csv(allcatch, paste0(o.name,"observation_catch_NOBA_allfisheries.csv"))
 ```

### 2021-04-30 docs/SimData.Rmd
WIP stomach weight inputs

@@
 
-#requires function below
+#requires read_savedfisheries function below, add to atlantisom
 catchlen_ss <- read_savedfisheries(d.name, "catchLen")
@@
 
+What we have saved so far is a diet comp in proportion that has had bias and observation error added to proportions.
+However, we need the stomach weight portion of the equation, which has the survey timing/area/efficiency/selectivity applied to true consumed weight.
+
 ```{r stomwt, eval=FALSE}
+
+# apply surveys to true total consumption (snipped from atlantosom om_diet.R and om_comps.R)
+detaileddiet <- readRDS(file.path(d.name, paste0(scenario.name, "detaileddiet.rds")))
+
+usersurvey <- c(here("simulated-data/config/mssurvey_spring.R"),
+                here("simulated-data/config/mssurvey_fall.R"),
+                here("simulated-data/config/mssurvey_spring_01.R"),
+                here("simulated-data/config/mssurvey_fall_01.R"))
+
+survObsStomWt <- list()
+
+for (s in usersurvey)
+{
+  source(s, local = TRUE)
+
+  # survtime doesn't match units of time.days in detaileddiet
+  survtime <- survey_sample_full*omlist_ss$runpar$outputstep
+
+  # apply survey design to detailed diet
+  survey_cons <- create_survey_diet(dat = detaileddiet,
+                                    time = survtime,
+                                    species = survspp,
+                                    boxes = survboxes,
+                                    effic = surveffic,
+                                    selex = survselex)
+
+  # get numbers at ageclass for same survey design
+  survey_N <- atlantisom::create_survey(dat = omlist_ss$truenums_ss,
+                                        time = survtime,
+                                        species = survspp,
+                                        boxes = survboxes,
+                                        effic = surveffic,
+                                        selex = survselex.agecl)
+
+  # divide surv_cons by survey numbers at agecl to get per capita consumption at agecl
+
+  #save survey consumption, takes a long time to generate with lots of reps/species
+  #if(save){
+  saveRDS(survey_cons, file.path(d.name, paste0(scenario.name, "_",
+                                                survey.name, "surveycons.rds")))
+  #}
+
+  survObsStomWt[[survey.name]] <- survey_cons
+
+}
+
+# divide by survey numbers at agecl to get per capita consumption at agecl
+
+# map to length bins using surey numbers at agecl in length bins
+
+# #multiple surveys named in list object, use as filename
+# for(s in names(all_diets)){
+#   #arrange into wide format: year, Species1, Species2 ... and write csv
+#   svdiet <- all_diets[[s]][[1]] %>%
+#     #calc_timestep2time(d.name, ., run.prm.file) %>% #this gives an arbitrary start year of 1970
+#     dplyr::mutate(time = time.days/NOBAom_ms$runpar$toutinc) %>%
+#     dplyr::left_join(., trueBagecl) %>%
+#     dplyr::mutate(dietBwt = dietSamp*trueBatage) %>%
+#     dplyr::group_by(species,time,prey,trueB) %>%
+#     dplyr::summarise(totdietBwt = sum(dietBwt)) %>%
+#     dplyr::mutate(totdietComp = totdietBwt/trueB) %>%
+#     dplyr::mutate(year = ceiling(time/stepperyr)) %>%
+#     dplyr::ungroup() %>%
+#     dplyr::select(year, species, prey, totdietComp) %>%
+#     write_csv(paste0(o.name,"diet_",s,".csv"))
+# }
+
+```
+
+We also plan to fit to time series of prey proportions for each predator size class. Gavin suggests an input format similar to those for survey lengths.
+
+```{r preyatsize, eval=FALSE}
+
+# same file as for MSSPM, diet proportion by agecl
+all_diets <- read_savedsurvs(d.name, 'survDiet')
+
+#
 

### 2021-05-13 docs/SimData.Rmd
daily stomach weight attempt 1, looks wrong

@@
 However, we need the stomach weight portion of the equation, which has the survey timing/area/efficiency/selectivity applied to true consumed weight.
+The hydra input is mean daily stomach weight in grams.
+From the Atlantis manual, "DetailedDietCheck.txt returns the total consumed biomass since the last output given for each cell (box and layer) of each age group of each functional group." So daily per capita consumed biomass is this output, summed over prey for total consumption, divided by the numbers from the same survey design, divided by the output step `omlist_ss$runpar$outputstep`.
 
@@
                 here("simulated-data/config/mssurvey_fall_01.R"))
+
+omlist_ss <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
+
+source(here("simulated-data/config/omdimensions.R"))
 
@@
   # get numbers at ageclass for same survey design
+  # note different time units!
   survey_N <- atlantisom::create_survey(dat = omlist_ss$truenums_ss,
-                                        time = survtime,
+                                        time = survey_sample_full,
                                         species = survspp,
@@
 
-  # divide surv_cons by survey numbers at agecl to get per capita consumption at agecl
+  # convert survey N times to cons times in days
+  survey_N$time <- survey_N$time*omlist_ss$runpar$outputstep
+
+  survey_totN <- survey_N %>%
+    group_by(species, agecl, time) %>%
+    summarise(totN = sum(atoutput)) %>%
+    ungroup()
+
+  # sum over prey to get total consumption in t, divide by N and timestep, convert to g
+  survey_totcons <- survey_cons %>%
+    group_by(species, agecl, time) %>%
+    summarise(totcons = sum(atoutput)) %>%
+    ungroup() %>%
+    left_join(survey_totN) %>%
+    mutate(percap_cons = totcons/totN) %>%
+    mutate(daily_percap_g = percap_cons/omlist_ss$runpar$outputstep*1000000)
 

### 2021-05-21 docs/SimData.Rmd
partial mapping agecl to lengthbin for diet

@@
 
+  # get rid of polygon
   survey_totN <- survey_N %>%
@@
     mutate(percap_cons = totcons/totN) %>%
-    mutate(daily_percap_g = percap_cons/omlist_ss$runpar$outputstep*1000000)
+    #mutate(daily_percap_g = percap_cons/omlist_ss$runpar$outputstep*1000000)
+    mutate(daily_percap_g = percap_cons*1000000) #try assuming cons is snapshot not cumulative since last timestep
 
@@
 }
-
-# divide by survey numbers at agecl to get per capita consumption at agecl
 
@@
 
-We also plan to fit to time series of prey proportions for each predator size class. Gavin suggests an input format similar to those for survey lengths.
+We also plan to fit to time series of prey proportions for each predator size class. Gavin suggests an input format similar to those for survey lengths. Diet output from surveys is by predator agecl
 
@@
 all_diets <- read_savedsurvs(d.name, 'survDiet')
+
+# get size bins for predators
+sizebins <- read_csv(paste0(o.name,"length_sizebins_NOBA.csv"))
+
+modbins <- sizebins %>%
+  rename(species = `commentField-notused`) %>%
+  pivot_longer(!species, names_to = "sizebin", values_to = "binwidth") %>%
+  group_by(species) %>%
+  arrange(sizebin) %>%
+  mutate(modbin.min = pcumsum(binwidth),
+         modbin.max = cumsum(binwidth)) %>%
+  arrange(species, sizebin)
+
+# diet is by predator agecl, need diet by predator sizebin
+# map predator agecl to lengthbin using length_comp_data from same survey
+
+allsvdietprop <- tibble()
+
+#multiple surveys named in list object, add to file
+for(s in names(all_diets)){
+  #arrange into long format: year, Species, agecl, length 
+  svagelenbin <- len_comp_data[[s]][[1]] %>%
+    dplyr::mutate(year = ceiling(time/stepperyr)) %>%
+    dplyr::select(species, agecl, year, upper.bins, atoutput) %>%
+    dplyr::left_join(modbins) %>%
+    dplyr::filter(modbin.min < upper.bins & upper.bins <= modbin.max) %>%
+    dplyr::group_by(species, year, agecl, sizebin) %>%
+    dplyr::summarise(sumlen = sum(atoutput)) %>%
+    dplyr::group_by(species, year, sizebin) %>%
+    dplyr::mutate(propage = sumlen/sum(sumlen)) #proportion of each agecl contributing to lengthbin
+
+  svdietprop <- all_diets[[s]][[1]] %>%
+    dplyr::mutate(year = ceiling(time.days/365)) %>%
+    # join with agecl to length lookup
+    dplyr::select(species, agecl, year, prey, dietSamp) %>%
+    dplyr::left_join(svagelenbin) %>%
+
+    dplyr::filter(modbin.min < upper.bins & upper.bins <= modbin.max) %>%
+    dplyr::group_by(species, year, sizebin) %>%
+    dplyr::summarise(sumlen = sum(atoutput)) %>%
+    tidyr::spread(sizebin, sumlen) %>%
+    ungroup() %>%
+    dplyr::mutate(inpN = rowSums(.[,-(1:2)], na.rm = TRUE)) %>%
+    dplyr::mutate(type = 0) %>%
+    dplyr::mutate(survey = s) %>%
+    dplyr::mutate_at(vars(contains("sizebin")), ~./inpN) %>%
+    dplyr::select(survey, year, species, type, inpN, everything())
+
+  allsvlenbin <- bind_rows(allsvlenbin, svlenbin)
+}
+write_csv(allsvlenbin, paste0(o.name,"observation_lengths_NOBA_allsurvs.csv"))
 

### 2021-05-24 docs/SimData.Rmd
first try observation diets in Gavin's format

@@
 
+# get Dirichlet alphamults from config files
+svcon <- list.files(path=here("/simulated-data/config/"), pattern = "*survey*", full.names = TRUE)
+
+svalphamultlook <- tibble()
+for(c in 1:length(svcon)){
+  source(svcon[c])
+  svalp <- tibble(
+    surv_alphamult_n = alphamult, 
+    survey = survey.name
+  )
+  svalphamultlook <- bind_rows(svalphamultlook, svalp)
+}
+
 # diet is by predator agecl, need diet by predator sizebin
@@
 
-#multiple surveys named in list object, add to file
+#multiple surveys named in list object
 for(s in names(all_diets)){
-  #arrange into long format: year, Species, agecl, length 
+  #arrange into lookup for agecl to lengthbin in numbers of fish 
   svagelenbin <- len_comp_data[[s]][[1]] %>%
@@
     dplyr::left_join(svagelenbin) %>%
-
-    dplyr::filter(modbin.min < upper.bins & upper.bins <= modbin.max) %>%
-    dplyr::group_by(species, year, sizebin) %>%
-    dplyr::summarise(sumlen = sum(atoutput)) %>%
-    tidyr::spread(sizebin, sumlen) %>%
+    dplyr::mutate(dietpropage = dietSamp*propage) %>% #reweight diets for lengthbins
+    dplyr::group_by(species, year, sizebin, prey) %>%
+    dplyr::summarise(dietsize = sum(dietpropage)) %>%
+    dplyr::filter(prey %in% unique(modbins$species)) %>% #drops predators that don't eat our modeled species
+    tidyr::spread(prey, dietsize) %>%
     ungroup() %>%
-    dplyr::mutate(inpN = rowSums(.[,-(1:2)], na.rm = TRUE)) %>%
-    dplyr::mutate(type = 0) %>%
+    dplyr::filter(!is.na(sizebin)) %>% 
+    dplyr::mutate(allotherprey = 1-rowSums(.[,-(1:3)], na.rm = TRUE)) %>%
     dplyr::mutate(survey = s) %>%
-    dplyr::mutate_at(vars(contains("sizebin")), ~./inpN) %>%
-    dplyr::select(survey, year, species, type, inpN, everything())
-
-  allsvlenbin <- bind_rows(allsvlenbin, svlenbin)
-}
-write_csv(allsvlenbin, paste0(o.name,"observation_lengths_NOBA_allsurvs.csv"))
+    dplyr::left_join(svalphamultlook) %>% #what is effective sample size for Dirichlet?
+    dplyr::rename(inpN = surv_alphamult_n) %>% #use this input for now
+    dplyr::select(survey, year, species, sizebin, inpN, everything())
+
+  allsvdietprop <- bind_rows(allsvdietprop, svdietprop)
+}
+write_csv(allsvdietprop, paste0(o.name,"observation_diets_NOBA_allsurvs.csv"))
 

### 2021-05-24 docs/SimData.Rmd
fix readlines code blocks, update html

@@
 
-```{r, code = readLines(here("/simulated-data/config/NOBA2Config.R")), eval=F}
+```{r, code = readLines(here("simulated-data/config/NOBA2Config.R")), eval=F}
 ```
@@
 
-```{r, code = readLines(here("/simulated-data/config/omdimensions.R")), eval=F}
+```{r, code = readLines(here("simulated-data/config/omdimensions.R")), eval=F}
 ```
@@
 
-```{r, code = readLines(here("/simulated-data/config/mssurvey_spring.R")), eval=F}
-```
-
-```{r, code = readLines(here("/simulated-data/config/mssurvey_fall.R")), eval=F}
+```{r, code = readLines(here("simulated-data/config/mssurvey_spring.R")), eval=F}
+```
+
+```{r, code = readLines(here("simulated-data/config/mssurvey_fall.R")), eval=F}
 ```
@@
 
-```{r, code = readLines(here("/simulated-data/config/msfishery.R")), eval=F}
+```{r, code = readLines(here("simulated-data/config/msfishery.R")), eval=F}
 ```

### 2021-06-18 docs/GBLandingsByLength.Rmd
add R files to read and process length and landings data -rmd to document process

---
title: "Landings by Length for Georges Bank"
author: Andy Beet
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    code_fold: hide
    number_sections: false
csl: apa-annotated-bibliography.csl
bibliography: references.bib
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
options(warn=-1)
```

## Goal

The ultimate goal is to create a data set that contains total annual biomass divided into size classes based on length for each of the [focal species](https://noaa-edab.github.io/ms-keyrun/SpeciesEnvDat.html#focalSpecies) and for each fleet that targets them. The resulting table will have a format similar to:

```{r tabformat, echo=F,eval=T}

df <- data.frame(Fleet = c(1,1,1,1,1,1),
           Area=c(1,1,1,1,1,1),
           Year =c(1964,1965,1966,1964,1965,1966),
           Species=c(1,1,1,2,2,2),
           Biomass=c(222,345,343,2137,678,346),
           sizeclass1 = c(0,0,0,0,0,0),
           sizeclass2 = c(0.05,0.05,0.05,0.05,0.05,0.05),
           sizeclass3 = c(0.25,0.25,0.25,0.25,.25,.25),
           sizeclass4 = c(0.5,0.5,0.5,0.5,.5,.5),
           sizeclass5 = c(.25,.25,.25,.25,.25,.25))

 df %>%  kableExtra::kbl() %>%
  kableExtra::kable_styling("hover", full_width = F)

```

Columns with the format `sizeclassx` represent the xth size class. A size class is defined by a range of lengths (cm) and its width is dependent on the species. Size classes were based on life history traits. (Expand this). For the focal species, the widths of each size class are defined in the following table:

```{r hydrabins, echo=F,eval=T,width = "50%", }
tab <- hydradata::hydraDataList$binwidth
row.names(tab) <- c("SPINY DOGFISH","WINTER SKATE","ATLANTIC HERRING", "ATLANTIC COD","HADDOCK","YELLOWTAIL FLOUNDER","WINTER FLOUNDER","ATLANTIC MACKEREL", "SILVER HAKE","GOOSEFISH")
tab %>%  kableExtra::kbl() %>%
  kableExtra::kable_styling("hover", full_width = F)

```

For example, the first size class for SPINY DOGFISH is [0-20] cm, the second class is (20-40] cm, ..., the 5 size class is (80-110] cm.

## Data

Commercial fisheries landings were pulled using R package [`comlandr`](https://github.com/NOAA-EDAB/comlandr)
```{r comlandLandings,eval=F,echo=T}
comlandData <- comlandr::get_comland_data(channel = channel, filterByYear = c(1964:2019),
    useLanded = T, removeParts = T, useHerringMaine = T, useForeign = F, 
    refYear = NA, refMonth = NA, disagSkatesHakes = T)
```

The length data were pulled from the `stockeff.mv_cf_len` table
```{r comlandlengths,eval=F,echo=T}
channel <- dbutils::connect_to_database("server","user") 
lengthData <- hydra_pull_GB_lengths(channel, area=c(cfdbs::EPUs$data$GB,537))

```

For each species of interest, the length data were then aggregated to provide the number of samples of a given fish LENGTH for each YEAR, NEGEAR combination. The landings data were also aggregated to provide total landings (metric tons) by YEAR and NEGEAR. These two aggregated data sets were then used to [expand the total landings](#expand) in a given YEAR, NEGEAR based on the sampled length distribution.

```{r processLandLen, echo=T,eval=F}
process_hydra_GB_comland(channel,comlandData,lengthData)
```

## Method to expand landings  {#expand}

For each unique category (YEAR, NEGEAR) we have both total landings (metric tons) and a sample of fish lengths. The sampled fish lengths are converted to weights using the [weight-length relationship](#wl). The resulting sample fish weights are mean weights (metric tons). The weight distribution of sampled fish is then scaled to the weight of total landings. This scaling assumes that the landed (commercial) fish have the same length distribution as the sampled fish.

For each unique category we define

$$expansion \; factor = \frac{total \; landings}{\sum sampled \; fish \; weigths} $$
Therefore the expanded weights of the sample (to meet the total landings)

$$expanded \; sample\; weights \; = expansion \; factor \;*\; sampled \; fish \; weights$$

## Weight-length relationship {#wl}

The length-weight data were obtained from SVDBS and a length-weight relationship was fit.

$$ W_i = \alpha L_i ^ \beta e^{z_i}$$

for fish i, where $\alpha$ and $\beta$ are intercept and slope parameters (on log scale) and $z_i \sim N(0,\sigma^2)$

### Note:

This model can be extended to :

$$ W_{ji} = \alpha L_{ji} ^ {\beta_j} e^{z_i} $$
for season j, j= 1, ..., 4 (spring, summer, fall, winter) or sex j, j = 1, ..., 3 (0, 1, 2 categories) or extended further to include season:sex combinations. These models are all nested and can therefore be tested using standard statistical methods. For the ms-keyrun project we did not consider extended models based on season or sex.

## Size class proportions

For each species the annual landings (by NEGEAR) which are now apportioned to a population of fish lengths can be aggregated over the size class intervals described in table x (link).

## References

### 2021-06-18 R/hydra_format.R

@@ -1,0 +1,194 @@
+#' Format expanded landings for Hydra
+#'
+#'
+#' 1. Consolidate NEGEAR to NEGEAR2 (2 or 3 digit gear code use)
+#' 2. Drop market category
+#' 3. Find total Biomass by fleet and year for each species
+#' 4. Within fleet and year attribute biomass of fish into length bins
+#'   specified by hydra (hydradata::hydraDataList$binwidth)
+#' @noRd
+
+
+
+format_hydra <- function(expandedLandings,fleets,itis) {
+
+  # truncate Gear code to 2 digit gear code
+  new <- expandedLandings %>%
+    dplyr::mutate(NEGEAR2 = base::substr(NEGEAR,start=1,stop=2)) %>%
+    dplyr::select(-NEGEAR)
+
+  dataNew <- new %>% dplyr::left_join(.,fleets,by=("NEGEAR2")) %>%
+    dplyr::group_by(YEAR,species_itis,hydraFleets,LENGTH) %>%
+    dplyr::summarise(totalWeight=sum(weight),.groups="drop")
+
+  hydraSpecies <- data.frame(oldName=c("Acod","Aherring","Amackerel","goosefish","haddock",
+                                       "silverhake","spinydog","winterfl","winterskate","yellowtailfl"),
+                             newName=c("ATLANTIC COD","ATLANTIC HERRING","ATLANTIC MACKEREL","GOOSEFISH","HADDOCK",
+                                       "SILVER HAKE","SPINY DOGFISH","WINTER FLOUNDER","WINTER SKATE","YELLOWTAIL FLOUNDER")) %>%
+    dplyr::left_join(.,mscatch::speciesLookupTable,by=c("newName"="COMMON_NAME.y")) %>%
+    dplyr::select(oldName,newName,SPECIES_ITIS) %>%
+    dplyr::distinct()
+
+  # organizes hydra length bins into long format table, change species name format , add itis
+  modbins <- hydradata::hydraDataList$binwidth %>%
+    dplyr::mutate(species = rownames(.)) %>%
+    tidyr::pivot_longer(!species, names_to = "sizebin", values_to = "binwidth") %>%
+    dplyr::group_by(species) %>%
+    dplyr::arrange(sizebin) %>%
+    dplyr::mutate(min = cumsum(binwidth)-binwidth,
+                  max = cumsum(binwidth)) %>%
+    dplyr::arrange(species, sizebin) %>%
[…]

### 2021-06-18 R/hydra_process_GB_comland.R

@@ -1,0 +1,137 @@
+#' sample Data pull of all species in Hydra model on GB
+#'
+#'Sample data in the format we'll need from proper data pulls using Seans comland script.
+#'Landings pulled using comlandr
+#'
+#'All missing areas need to dealt with prior to pulling species data.
+#'Not the case in this sample. We pull from GB
+#'
+#'@param channel obtained from logging into the sole database.
+#'@param comlandData data from comland pull
+#'
+#'
+
+#channel <- cfdbs::connect_to_database("server","user")
+# need to pull comland data. Missing MARKET_CODE because pull from AA not stockeff
+# comlandData <- readRDS(here::here("data-raw","comland_negear.rds"))
+# lengthData <- readRDS(here::here("data-raw","hydra_lengths.rds")) # see test_data_pull_hydra_GB_lengths.r
+# library(magrittr)
+
+process_hydra_GB_comland <- function(channel,comlandData,lengthData){
+
+  hydraSpecies <- data.frame(oldName=c("Acod","Aherring","Amackerel","goosefish","haddock",
+                                       "silverhake","spinydog","winterfl","winterskate","yellowtailfl"),
+                             newName=c("ATLANTIC COD","ATLANTIC HERRING","ATLANTIC MACKEREL","GOOSEFISH","HADDOCK",
+                                       "SILVER HAKE","SPINY DOGFISH","WINTER FLOUNDER","WINTER SKATE","YELLOWTAIL FLOUNDER")) %>%
+    dplyr::left_join(.,mscatch::speciesLookupTable,by=c("newName"="COMMON_NAME.y")) %>%
+    dplyr::select(oldName,newName,SPECIES_ITIS,SVSPP) %>%
+    dplyr::distinct()
+
+  GBArea <- c(cfdbs::EPUs$data$GB,537)
+
+  ## for test data assume we this is EPU data. To achieve this we just sum over AREAS for now
+  lands <- comlandData$comland %>%
+    dplyr::filter(AREA %in% GBArea) %>%
+    #dplyr::rename(MARKET_CODE = MKTCAT) %>%
+    dplyr::group_by(YEAR, MONTH, NEGEAR, NESPP3) %>%
+    #dplyr::mutate(SPPLNDLB = SPPLIVMT/0.00045359237)  %>%
[…]

### 2021-06-18 R/hydra_pull_GB_lengths.R

@@ -1,0 +1,50 @@
+#' sample Data pull. All hydra species lengths
+#'
+#'Lengths pulled from "mv_cf_len" filtered to GB stat areas
+#'
+#'
+#'@param channel obtained from logging into the sole database.
+#'@param area stat areas to pull lengths from
+#'
+
+
+#channel <- cfdbs::connect_to_database("server","user") eventually remove this
+
+pull_hydra_GB_lengths <- function(channel, area=c(cfdbs::EPUs$data$GB,537)){
+  library(magrittr)
+
+  hydraSpecies <- data.frame(oldName=c("Acod","Aherring","Amackerel","goosefish","haddock",
+                                       "silverhake","spinydog","winterfl","winterskate","yellowtailfl"),
+                             newName=c("ATLANTIC COD","ATLANTIC HERRING","ATLANTIC MACKEREL","GOOSEFISH","HADDOCK",
+                                       "SILVER HAKE","SPINY DOGFISH","WINTER FLOUNDER","WINTER SKATE","YELLOWTAIL FLOUNDER")) %>%
+    dplyr::left_join(.,mscatch::speciesLookupTable,by=c("newName"="COMMON_NAME.y")) %>%
+    dplyr::select(oldName,newName,SPECIES_ITIS) %>%
+    dplyr::distinct()
+
+
+  # pull sample length data and massage it
+  message("Puling length data ...")
+
+  testDataPullLength <- cfdbs::get_landings_length(channel,year="all",area=area,species=hydraSpecies$SPECIES_ITIS,species_itis=T)
+
+  # create unique tripid since NUMSAMP is replicated for each species reported within a trip
+  lengths <- testDataPullLength$data %>%
+    dplyr::mutate(tripid = paste0(PERMIT,YEAR,MONTH,DAY))
+
+  # just extract the lengths and the number at length for the year, qr etc
+  sampleLengths <- lengths %>%
+    dplyr::select(YEAR,QTR,SPECIES_ITIS,NEGEAR,MARKET_CODE,LENGTH,NUMLEN,tripid) %>%
+    dplyr::arrange(SPECIES_ITIS,YEAR,QTR,NEGEAR,MARKET_CODE,LENGTH)
+  sampleLengths$SPECIES_ITIS <- as.integer(sampleLengths$SPECIES_ITIS)
+  sampleLengths$YEAR <- as.integer(sampleLengths$YEAR)
[…]

### 2021-06-18 docs/GBLandingsByLength.Rmd
add html GBLandingsByLength version of rmd for documention

@@
 ```{r processLandLen, echo=T,eval=F}
-process_hydra_GB_comland(channel,comlandData,lengthData)
+expanded <- hydra_process_GB_comland(channel,comlandData,lengthData)
 ```
@@
 
-For each species the annual landings (by NEGEAR) which are now apportioned to a population of fish lengths can be aggregated over the size class intervals described in table x (link).
+For each species the annual landings (by NEGEAR) which are now apportioned to fish lengths can be aggregated over the size class intervals described in table x (link) based on the [fleets](https://noaa-edab.github.io/ms-keyrun/GBFleetDecisions.html) the NEGEAR codes are associated with.
+The Hydra [simulation and estimation](https://github.com/NOAA-EDAB/hydra_sim/wiki) model implements three fleets: demersal, pelagic, and fixed gear.
+For this model the broader [fleet classification](https://noaa-edab.github.io/ms-keyrun/GBFleetDecisions.html) is aggregated further (Table xxx).
+
+Any NEGEAR gear types that do not neatly fall into any of the three defined fleets (i.e.
+the `other` category) are assigned to the fleet dominating landings.
+The proportions of total landings by size class for each fleet can then be calculated.
+
+```{r hydraformat, echo=T, eval= F}
+  speciesList <- unique(expanded$species_itis)
+  for (itis in speciesList) {
+    hydra_format(expanded,mscatch::fleets,itis)
+  }
+```
+
+```{r hydrafleet, echo=T, eval=T}
+mscatch::fleets %>%  kableExtra::kbl() %>%
+  kableExtra::kable_styling("hover", full_width = F)
+```
 

### 2021-06-18 R/hydra_format.R

@@ -61,7 +61,7 @@ format_hydra <- function(expandedLandings,fleets,itis) {
       dplyr::summarize(w = sum(totalWeight),.groups="drop")
 
     # for all fleets calculate the proportions of landings in each size bin
-    # remove other fleet and landings witout length bins.
+    # remove other fleet and landings without length bins.
     # These dealt with separately
     proportionsinit <- spd %>% dplyr::group_by(YEAR,hydraFleets) %>%
       dplyr::filter(hydraFleets != "other") %>%

### 2021-06-22 docs/GBLandingsByLength.Rmd
added proportional landings by sizeclass data file and table in rmd

@@
 speciesList <- unique(expanded$species_itis)
+finalTable <- NULL
 for (itis in speciesList) {
-    hydra_format(expanded,mscatch::fleets,itis)
+  itisData <- hydra_format(expanded,mscatch::fleets,itis)
+  finalTable <- rbind(finalTable,itisData)
 }
+
+write.csv(finalTable,here::here("data-raw","catchAtLengthProportions.csv"))
 ```
+
+The resulting table for the species of interest over 5 size class intervals. totalBiomass is in metric tons.
+
+```{r props, echo = T,eval=T}
+readr::read_csv(here::here("data-raw","catchAtLengthProportions.csv"),col_types = readr::cols()) %>% 
+  dplyr::mutate(across(dplyr::starts_with("sizeclass"),round,5)) %>% 
+  dplyr::mutate(totalBiomass=round(totalBiomass)) %>%
+  DT::datatable(.,rownames = FALSE, 
+          options = list(pageLength = 10, 
+                         order = list(list(0, 'asc'))
+          ))
+
+```
+
+## Assign fleets definition to hydra fleets
+
+The [fleet](https://noaa-edab.github.io/ms-keyrun/GBFleetDecisions.html) definitions are further aggregated to meet the needs of the Hydra model
 

### 2021-06-22 R/hydra_format.R

@@ -10,7 +10,7 @@
 
 
 
-format_hydra <- function(expandedLandings,fleets,itis) {
+hydra_format <- function(expandedLandings,fleets,itis) {
 
   # truncate Gear code to 2 digit gear code
   new <- expandedLandings %>%

### 2021-06-22 R/hydra_process_GB_comland.R

@@ -15,7 +15,7 @@
 # need to pull comland data. Missing MARKET_CODE because pull from AA not stockeff
 # comlandData <- readRDS(here::here("data-raw","comland_negear.rds"))
 # lengthData <- readRDS(here::here("data-raw","hydra_lengths.rds")) # see test_data_pull_hydra_GB_lengths.r
-# library(magrittr)
+ library(magrittr)
 
 process_hydra_GB_comland <- function(channel,comlandData,lengthData){
 
@@ -132,6 +132,6 @@ process_hydra_GB_comland <- function(channel,comlandData,lengthData){
 
   }
 
+  saveRDS(expanded,here::here("data-raw","expandedHydraLandings.rds"))
   return(expanded)
-  #saveRDS(expanded,here::here("data-raw","expandedHydraLandings.rds"))
 }

### 2021-08-02 diagnostics/DiagnosticPlotExamples.Rmd
add initial diagnostic scripts

---
title: "Diagnostic scripts"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_fold: hide
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(here)
library(tidyverse)  
library(atlantisom)
library(ggthemes)
library(FSA)

```

# Plotting functions for comparisons

A collection of functions used previosuly that may be harvested and modified for diagnostics or visualizations in the ms-keyrun project.

## Multispecies output plotting functions

These are also assuming you are using atlantisom output datasets but could be generalized.

My plotting functions: too specialized to include in the package?

```{r plotfuns}

# plot biomass time series facet wrapped by species
plotB <- function(dat, truedat=NULL){

    ggplot() +
    geom_line(data=dat, aes(x=time/stepperyr,y=atoutput, color="Survey Biomass"), 
              alpha = 10/10) +
    {if(!is.null(truedat)) geom_line(data=truedat, aes(x=time/365,y=atoutput, color="True B"), alpha = 3/10)} + 
    theme_tufte() +
    theme(legend.position = "top") +
    xlab("model year") +
    ylab("tons") +
    labs(colour=scenario.name) +
    facet_wrap(~species, scales="free")

}

# make a catch series function that can be split by fleet? this doesnt
# also note different time (days) from model timestep in all other output
plotC <- function(dat, truedat=NULL){

    ggplot() +
    geom_line(data=dat, aes(x=time/365,y=atoutput, color="Catch biomass"), 
              alpha = 10/10) +
    {if(!is.null(truedat)) geom_line(data=truedat, aes(x=time/365,y=atoutput, color="True Catch"), alpha = 3/10)} + 
    theme_tufte() +
    theme(legend.position = "top") +
    xlab("model year") +
    ylab("tons") +
    labs(colour=scenario.name) +
    facet_wrap(~species, scales="free")

}

# note on ggplot default colors, can get the first and second using this
# library(scales)
# show_col(hue_pal()(2))

# plot length frequencies by timestep (one species)
plotlen <- function(dat, effN=1, truedat=NULL){

  cols <- c("Census Lcomp"="#00BFC4","Sample Lcomp"="#F8766D")  
  ggplot(mapping=aes(x=upper.bins)) +
    {if(is.null(truedat)) geom_bar(data=dat, aes(weight = atoutput/effN))} +
    {if(!is.null(truedat)) geom_bar(data=dat, aes(weight = censuslen/totlen, fill="Census Lcomp"), alpha = 5/10)} +
    {if(!is.null(truedat)) geom_bar(data=dat, aes(weight = atoutput/effN, fill="Sample Lcomp"), alpha = 5/10)} +
    theme_tufte() +
    theme(legend.position = "bottom") +
    xlab("length (cm)") +
    {if(is.null(truedat)) ylab("number")} +
    {if(!is.null(truedat)) ylab("proportion")} +
    scale_colour_manual(name="", values=cols) +
    labs(subtitle = paste(scenario.name,
                          dat$species)) +
    facet_wrap(~time, ncol=6, scales="free_y")

}

# plot numbers at age by timestep (one species)
Natageplot <- function(dat, effN=1, truedat=NULL){
  ggplot() +
    geom_point(data=dat, aes(x=agecl, y=atoutput/effN, color="Est Comp")) +
    {if(!is.null(truedat)) geom_line(data=dat, aes(x=agecl, y=numAtAge/totN, color="True Comp"))} + 
    theme_tufte() +
    theme(legend.position = "bottom") +    
    xlab("age/agecl") +
    {if(is.null(truedat)) ylab("number")} +
    {if(!is.null(truedat)) ylab("proportion")} +
    labs(subtitle = paste(scenario.name,
                          dat$species)) + 
    facet_wrap(~time, ncol=6, scales="free_y")
}

# plot weight at age time series facet wrapped by species
wageplot <- function(dat, truedat=NULL){
  ggplot(dat, aes(time/stepperyr, atoutput)) +
    geom_line(aes(colour = factor(agecl))) +
    theme_tufte() +
    theme(legend.position = "bottom") +
    xlab("model year") +
    ylab("average individual weight (g)") +
    labs(subtitle = paste0(scenario.name)) +
    facet_wrap(c("species"), scales="free_y")
}

# compare N at age and C at age between standard and ANNAGE outputs
totNageplot <- function(dat, anndat){

    ggplot() +
    geom_line(data=dat, aes(x=time/stepperyr,y=totN, color="Tot N cohorts"), 
              alpha = 5/10) +
    geom_line(data=anndat, aes(x=time/stepperyr,y=totN, color="Tot N annage"), alpha = 5/10) + 
    theme_tufte() +
    theme(legend.position = "top") +
    xlab("model year") +
    ylab("N") +
    labs(colour=scenario.name) +
    facet_wrap(~species, scales="free")

}

```

Usage examples (from msSurveysTest.Rmd)

### Visualize census surveys {.tabset #census}

We are using the full time series for biomass and weight at age, and a subsample of years 30-53 for compositions. In this test we see the effect of surveying in one season, with migratory species missed or only partially represented in some surveys, and with the overall surveyed biomass being at the lower or higher end of the range depending on the time of year. (Observation error is included in the survey biomass time series.) Samples for length and age assume a total of 100000 fish were randomly sampled for each species in each survey (surveffN defined in the survey config files).

#### Survey biomass
```{r, results="asis"}

# compare with true output (all timesteps)
for(s in names(survObsBiom)){
  cat("  \n##### ",  s,"  \n")
  print(plotB(survObsBiom[[s]][[1]], omlist_ss$truetotbio_ss))
  cat("  \n")
}

# plots survey only
# for(s in names(survObsBiom)){
#   print(plotB(survObsBiom[[s]][[1]]))
# }

```

#### Survey length composition
```{r, results="asis"}

for(s in names(len_comp_data)){
  cat("  \n##### ",  s,"  \n")
  lcompsub <- as.data.frame(len_comp_data[[s]][[1]]) %>% filter(time %in% c(150:270)) %>%
    group_by(species) %>%
    group_map(~ plotlen(.x), keep = TRUE)

  for(i in 1:length(lcompsub)) {
    print(lcompsub[[i]])
  }
  cat("  \n")
}

```

#### Survey age class composition (Atlantis age class)
```{r, results="asis", message=FALSE, warning=FALSE}

for(s in names(age_comp_data)){
  cat("  \n##### ",  s,"  \n")
  acompsub <- as.data.frame(age_comp_data[[s]][[1]]) %>% filter(time %in% c(150:270)) %>%
    group_by(species) %>%
    left_join(., trueNagecl) %>%
    #group_map(~ Natageplot(.x), keep = TRUE) # plots only sampled age comp
    group_map(~ Natageplot(.x, effN = 100000, truedat = 1), keep = TRUE) # plots merged true age comp

  for(i in 1:length(acompsub)) {
    print(acompsub[[i]])
  }
  cat("  \n")
}

```

#### Survey age composition (annual ages)
```{r plot-annagecomp, results="asis", message=FALSE, warning=FALSE}

for(s in names(annage_comp_data)){
  cat("  \n##### ",  s,"  \n")
  acompsub <- as.data.frame(annage_comp_data[[s]][[1]]) %>% filter(time %in% c(150:270)) %>%
    group_by(species) %>%
    left_join(., trueNage) %>%
    #group_map(~ Natageplot(.x), keep = TRUE) # plots only sampled age comp
    group_map(~ Natageplot(.x, effN = 100000, truedat = 1), keep = TRUE) # plots merged true age comp

  for(i in 1:length(acompsub)) {
    print(acompsub[[i]])
  }
  cat("  \n")
}
```

#### Survey weight at age (Atlantis age class)
```{r, results='asis'}

for(s in names(wtage)){
  cat("  \n##### ",  s,"  \n")
  print(wageplot(wtage[[s]][[1]]))
  cat("  \n")
}

```

#### Survey iterpolated weight at age (annual ages) 
```{r plot-annwtage, fig.height=10, results='asis'}

# annage_wtage read in above 
# annage_wtage <- annage_wtage[[1]] #this still has second list component, diagnostic plot
# annage_wtage <- annage_wtage[[1]] #still in this old version but just removed from function.

for(s in names(annage_wtage)){
  cat("  \n##### ",  s,"  \n")
  print(wageplot(annage_wtage[[s]][[1]][[1]]))
  cat("  \n")
}
```

### {-}

## Diet data (simulated)

Plotting functions and colors for everyone

```{r color-plotfun}

# length(unique(samp2$prey)) #39 prey categories max 
# 35 categories in samp5 without unid length(unique) #http://medialab.github.io/iwanthue/
preycol <- c("#7b7927",
             "#746dd8",
             "#a6bc3a",
             "#4f3587",
             "#86af42",
             "#c666c7",
             "#42c87f",
             "#cb417f",
             "#6fb95c",
             "#882e7b",
             "#67b271",
             "#c83c63",
             "#43c8ac",
             "#da4953",
             "#3ba7e5",
             "#caa432",
             "#5081db",
             "#c8802f",
             "#6c79c2",
             "#c7b961",
             "#bb86d6",
             "#3b7125",
             "#d672b9",
             "#b57e43",
             "#872957",
             "#dd6741",
             "#db75a2",
             "#9e3d14",
             "#db6f81",
             "#7b3015",
             "#e08262",
             "#8d2a41",
             "#d56467",
             "#8c242e",
             "#a73830")
names(preycol) <- as.factor(sort(unique(samp5$prey)))

# going for more greyscale for unident categories, same website
unidcol <- c("#b8b8b2",
             "#302a1d",
             "#6b7069",
             "#1e3430")
names(unidcol) <- as.factor(c("Unid", "Unid_Fish",  "Unid_Invert", "Unid_Plankton"))

col <- c(preycol, unidcol)

# plot diet comp over time at age by species
plotdiet <- function(dat, compdat=NULL, namedat, namecomp=NULL){

  dat <- dat %>% add_column(run = namedat)
  if(!is.null(compdat)) compdat <- compdat %>% add_column(run = namecomp)

    ggplot() +
    geom_bar(data=dat, aes(time.days/365, dietSamp, fill=prey), stat = "identity") +
    {if(!is.null(compdat)) geom_bar(data=compdat, aes(time.days/365, dietSamp, fill=prey), stat = "identity")} + 
    theme_tufte() +
    theme(legend.position = "bottom") +
    xlab("year") +
    ylab("diet proportion") +
    facet_grid(agecl~run) + 
    scale_fill_manual(values=col) + 
    ggtitle(dat$species)

}

# method for a single species diet, no comparisons
# plist = lapply(split(ms_diet, ms_diet$species), function(d) {
#   ggplot(d, aes(time.days/365, atoutput, fill=prey)) + 
#     geom_bar(stat = "identity") +
#     facet_wrap(species~agecl) +
#     xlab("year") +
#     ylab("diet proportion") +
#     theme_tufte() +
#     theme(legend.position="bottom")
# })

```

Usage example, assuming you have an atlantis survey diet output in the `d.name` folder:

### Compare spring and fall surveys {.tabset}

The diet was sampled with alphamult = 10 and unidprey = 0.3 using the survey configuration files.

```{r, fig.width=9, fig.height=12, results='asis'}

NOBAom_ms_diet <- atlantisom::read_savedsurvs(d.name, 'survDiet')

preds <- unique(unique(NOBAom_ms_diet[[2]][[1]]$species))

fall <- as.data.frame(NOBAom_ms_diet$BTS_fall_nearbox_qmix_selmix[[1]])
spring <- as.data.frame(NOBAom_ms_diet$BTS_spring_nearbox_qmix_selmix[[1]])

for(i in 1:length(preds)) {
  cat("  \n####",  as.character(preds[i]),"  \n")
  print(plotdiet(dat = filter(spring, species %in% preds[i]), namedat = "Spring survey diet", compdat = filter(fall, species %in% preds[i]), namecomp = "Fall survey diet")) 
  cat("  \n")
}

```

### {-}

### 2021-08-02 diagnostics/DiagnosticPlotExamples.Rmd
cross reference full knitted docs

@@
 
-Usage examples (from msSurveysTest.Rmd)
+Usage examples (from https://github.com/sgaichas/poseidon-dev/blob/master/msSurveysTest.Rmd, see [visualization](https://sgaichas.github.io/poseidon-dev/msSurveysTest.html))
 
@@
 ## Diet data (simulated)
+
+From https://github.com/sgaichas/poseidon-dev/blob/master/SurveyDietCompTest.Rmd, see [visualizations](https://sgaichas.github.io/poseidon-dev/SurveyDietCompTest.html)
 

### 2021-08-27 data-raw/R/hydra_format.R
Added data folder. - updated catch at length code and rmd

@@ -1,11 +1,16 @@
 #' Format expanded landings for Hydra
 #'
 #'
+#'@param expandedLandings output from \code{hydra_process_GB_comland}
+#'@param fleets fleets to assign landings to (see \code{mscatch::fleets}
+#'@param itis itis code for species in which to process
+#'
 #' 1. Consolidate NEGEAR to NEGEAR2 (2 or 3 digit gear code use)
 #' 2. Drop market category
 #' 3. Find total Biomass by fleet and year for each species
 #' 4. Within fleet and year attribute biomass of fish into length bins
 #'   specified by hydra (hydradata::hydraDataList$binwidth)
+#'   
 #' @noRd
 
 

### 2021-08-27 data-raw/R/hydra_main.R

@@ -1,0 +1,40 @@
+#' create hydra data for package in rda format 
+#'
+#' As outlined in GBLandingsByLength.Rmd
+#'
+#'
+#'
+#'
+#'
+#'
+#'
+
+channel <- dbutils::connect_to_database("server","user")
+
+source(here::here("data-raw/R","hydra_pull_GB_lengths.R"))
+source(here::here("data-raw/R","hydra_process_GB_comland.R"))
+source(here::here("data-raw/R","hydra_format.R"))
+
+hydra_main <- function(channel){
+  
+  #pull comland data
+  comlandData <- readRDS(here::here("data-raw/data","comland_negear.rds"))
+  comlandData <- comlandData$comland
+  # get lengths in GB from cfdbs 
+  lengthData <- hydra_pull_GB_lengths(channel, area=c(cfdbs::EPUs$data$GB,537))
+  
+  # expand lengths as in mscatch
+  expanded <- hydra_process_GB_comland(channel,comlandData,lengthData)
+  
+  # for each species format the result
+  speciesList <- unique(expanded$species_itis)
+  catchAtLengthProportions <- NULL
+  for (itis in speciesList) {
+    itisData <- hydra_format(expanded,mscatch::fleets,itis)
+    catchAtLengthProportions <- rbind(catchAtLengthProportions,itisData)
+  }
+  
+  # save final table
+  usethis::use_data(catchAtLengthProportions)
+  
+}

### 2021-08-27 data-raw/R/hydra_process_GB_comland.R

@@ -8,6 +8,7 @@
 #'
 #'@param channel obtained from logging into the sole database.
 #'@param comlandData data from comland pull
+#'@param lengthData result of \code{hydra_pull_GB_lengths}
 #'
 #'
 
@@ -17,7 +18,7 @@
 # lengthData <- readRDS(here::here("data-raw","hydra_lengths.rds")) # see test_data_pull_hydra_GB_lengths.r
  library(magrittr)
 
-process_hydra_GB_comland <- function(channel,comlandData,lengthData){
+hydra_process_GB_comland <- function(channel,comlandData,lengthData){
 
   hydraSpecies <- data.frame(oldName=c("Acod","Aherring","Amackerel","goosefish","haddock",
                                        "silverhake","spinydog","winterfl","winterskate","yellowtailfl"),
@@ -30,7 +31,7 @@ process_hydra_GB_comland <- function(channel,comlandData,lengthData){
   GBArea <- c(cfdbs::EPUs$data$GB,537)
 
   ## for test data assume we this is EPU data. To achieve this we just sum over AREAS for now
-  lands <- comlandData$comland %>%
+  lands <- comlandData %>%
     dplyr::filter(AREA %in% GBArea) %>%
     #dplyr::rename(MARKET_CODE = MKTCAT) %>%
     dplyr::group_by(YEAR, MONTH, NEGEAR, NESPP3) %>%

### 2021-08-27 data-raw/R/hydra_pull_GB_lengths.R

@@ -10,7 +10,7 @@
 
 #channel <- cfdbs::connect_to_database("server","user") eventually remove this
 
-pull_hydra_GB_lengths <- function(channel, area=c(cfdbs::EPUs$data$GB,537)){
+hydra_pull_GB_lengths <- function(channel, area=c(cfdbs::EPUs$data$GB,537)) {
   library(magrittr)
 
   hydraSpecies <- data.frame(oldName=c("Acod","Aherring","Amackerel","goosefish","haddock",
@@ -23,7 +23,7 @@ pull_hydra_GB_lengths <- function(channel, area=c(cfdbs::EPUs$data$GB,537)){
 
 
   # pull sample length data and massage it
-  message("Puling length data ...")
+  message("Pulling length data ...")
 
   testDataPullLength <- cfdbs::get_landings_length(channel,year="all",area=area,species=hydraSpecies$SPECIES_ITIS,species_itis=T)
 

### 2021-08-30 data-raw/R/get_focal_species.R
export focalSpecies.rda to data folder

@@ -4,9 +4,9 @@
 #'
 #'@param channel Object inherited from \link[DBI]{DBIConnection-class}. This object is used to connect
 #' to communicate with the database engine. (see \code{\link{connect_to_database}})
-#'@param saveToRDS Character string. File path to where output should be saved (Default = NULL, not saved)
+#'@param saveToData Boolean. Export to data folder (Default = T)
 #'
-#'@return A tibble
+#'@return A tibble (Also written to \code{data} folder)
 #'\item{NESPP3}{code for secies found in commercial fishing database}
 #'\item{SVSPP}{code for species found in bottom trawl survey database}
 #'\item{NAFOSPP}{code for species in nafo database}
@@ -20,7 +20,7 @@
 
 library(magrittr)
 
-get_focal_species <- function(channel,saveToRDS=NULL) {
+get_focal_species <- function(channel,saveToData=T) {
   
   # create lookup table for species of interest
   lookup <- dbutils::create_species_lookup(channel,species=c(32,73,197,74,15,106,105,121,72,23,75),speciesType = "SVSPP")
@@ -29,8 +29,9 @@ get_focal_species <- function(channel,saveToRDS=NULL) {
     dplyr::rename(SVSPP=SVSPPsv) %>%
     dplyr::distinct(.)
   
-  if (!is.null(saveToRDS)) {
-    saveRDS(focalSpecies,saveToRDS)
+  if (saveToData) {
+    #saveRDS(focalSpecies,saveToRDS)
+    usethis::use_data(focalSpecies)
   }
   
   return(focalSpecies)

### 2021-08-30 data-raw/R/cluster_by_focal_species.R
update links to data-raw data

@@ -16,14 +16,14 @@
 
 source(here::here("R","cluster_gears.r"))
 library(magrittr)
-focalSpecies <- readRDS(here::here("data","focalspecies.rds"))
+#focalSpecies <- readRDS(here::here("data","focalspecies.rds"))
 
-cluster_by_focal_species <- function(focalSpecies){
+cluster_by_focal_species <- function(focalSpecies=mskeyrun::focalSpecies){
 
   # read in landings data by gear/species aggregated over time
-  allGearData <- readRDS(here::here("data","timeSeriesSpeciesByGear.rds"))
+  allGearData <- readRDS(here::here("data-raw/data","timeSeriesSpeciesByGear.rds"))
   # read in all gear codes
-  gearCodes <- readRDS(here::here("data","gearCodeTable.rds"))
+  gearCodes <- readRDS(here::here("data-raw/data","gearCodeTable.rds"))
   
   ## PROCESS GEAR CODES -----------------------------------
   # split long names by comma and select first name

### 2021-08-30 data-raw/R/plot_landings.R
catch proportion work - moved R and rds to data-raw/R data-raw/data

@@ -23,7 +23,7 @@ library(magrittr)
 
 plot_landings <- function() {
   # REVENUEFILE is a df where each record contains landings for a subtrip
-  REVENUEFILE <- readRDS(here::here("data-raw","Landings_VTR_Geret_Data_summarized.rds")) # This stores a variable called REVENUEFILE
+  REVENUEFILE <- readRDS(here::here("data-raw/data","Landings_VTR_Geret_Data_summarized.rds")) # This stores a variable called REVENUEFILE
 
   data <- REVENUEFILE %>% 
     dplyr::group_by(YEAR,AREA,NESPP3,COMMON_NAME) %>%

### 2021-08-30 data-raw/R/plot_landings_proportions.R

@@ -26,7 +26,7 @@ calc_prop <- function(landings,inout) {
 
 plot_landings_proportions <- function(species="cols") {
    # REVENUEFILE is a df where each record contains landings for a subtrip
-  REVENUEFILE <- readRDS(here::here("data-raw","Landings_VTR_Geret_Data_summarized.rds"))
+  REVENUEFILE <- readRDS(here::here("data-raw/data","Landings_VTR_Geret_Data_summarized.rds"))
   
   # calculate proportion of landings inside GB section of stat area
   data <- REVENUEFILE %>% 

### 2021-08-30 data-raw/R/hydra_main.R
changed name of rds file in script - comland_negear to hydra_comland_gear

@@ -18,7 +18,7 @@ source(here::here("data-raw/R","hydra_format.R"))
 hydra_main <- function(channel){
   
   #pull comland data
-  comlandData <- readRDS(here::here("data-raw/data","comland_negear.rds"))
+  comlandData <- readRDS(here::here("data-raw/data","hydra_comland_negear.rds"))
   comlandData <- comlandData$comland
   # get lengths in GB from cfdbs 
   lengthData <- hydra_pull_GB_lengths(channel, area=c(cfdbs::EPUs$data$GB,537))

### 2021-08-30 data-raw/R/calc_all_species_proportions.R
path chages to data location in cluster analysis scripts

@@ -28,7 +28,7 @@ calc_prop <- function(landings,inout) {
 calc_all_species_proportions <- function() {
   
   # REVENUEFILE is a df where each record contains landings for a subtrip
-  REVENUEFILE <- readRDS(here::here("data-raw","Landings_VTR_Geret_Data.rds")) # This stores a variable called REVENUEFILE
+  REVENUEFILE <- readRDS(here::here("data-raw/data","Landings_VTR_Geret_Data.rds")) # This stores a variable called REVENUEFILE
   
   # split the are column to define inside and out, sum landings by year, spp, area 
   data <- REVENUEFILE %>%
@@ -45,6 +45,6 @@ calc_all_species_proportions <- function() {
       dplyr::group_by(YEAR,AREA,NESPP3) %>% 
       dplyr::summarise(PROP_GB = calc_prop(TOTALLANDINGS,InOut),.groups = "drop")
  
-  saveRDS(props,file = here::here("data-raw","Landings_VTR_proportions_all_species.rds"))
+  saveRDS(props,file = here::here("data-raw/data","Landings_VTR_proportions_all_species.rds"))
   
 }

### 2021-08-30 data-raw/R/cluster_analysis_all_gears.R

@@ -19,9 +19,9 @@ library(magrittr)
 cluster_analysis_all_gears <- function(){
 
   # read in landings data by gear/species aggregated over time
-  allGearData <- readRDS(here::here("data","gearLandingsBySpecies.rds"))
+  allGearData <- readRDS(here::here("data-raw/data","gearLandingsBySpecies.rds"))
   # read in all gear codes
-  gearCodes <- readRDS(here::here("data","gearCodeTable.rds"))
+  gearCodes <- readRDS(here::here("data-raw/data","gearCodeTable.rds"))
   
   # split long names by comma and select first name
   gearCodes <- gearCodes$data %>% 

### 2021-08-30 data-raw/R/cluster_by_focal_species.R

@@ -14,7 +14,7 @@
 #' 
 #' 
 
-source(here::here("R","cluster_gears.r"))
+source(here::here("data-raw/R","cluster_gears.r"))
 library(magrittr)
 #focalSpecies <- readRDS(here::here("data","focalspecies.rds"))
 

### 2021-08-30 data-raw/R/All_Species_Landings_Proportions.R
removed All_species_proportions.rds and created lazy data. updated r script to crete this

@@ -1,7 +1,7 @@
 #Calculate proportions of commercial landings on Georges Bank
 library(here); library(data.table); library(magrittr)
 
-vtr <- as.data.table(readRDS(here('/data-raw/Landings_VTR_Geret_Data.rds')))
+vtr <- as.data.table(readRDS(here('/data-raw/data/Landings_VTR_Geret_Data.rds')))
   
 # split the are column to define inside and out, sum landings by year, spp, area and 
 # select just 12 species that make up 90% of landings
@@ -29,4 +29,7 @@ props <- unique(props, by = c('AREA', 'NESPP3', 'InOut'))
 #Drop extra columns
 props[, c('YEAR', 'TOTALLANDINGS', 'AllLandings', 'Prop') := NULL]
 
-saveRDS(props, file = here::here('data-raw/All_Species_Proportions.rds'))
+AllSpeciesProportions <- props
+
+#saveRDS(props, file = here::here('data-raw/data/All_Species_Proportions.rds'))
+usethis::use_data(AllSpeciesProportions)

### 2021-08-30 data-raw/R/hydra_process_GB_comland.R
changed name and location of hydra intermediate length and expanded landings rds files

@@ -133,6 +133,6 @@ hydra_process_GB_comland <- function(channel,comlandData,lengthData){
 
   }
 
-  saveRDS(expanded,here::here("data-raw","expandedHydraLandings.rds"))
+  saveRDS(expanded,here::here("data-raw/data","GB_hydraExpandedLandings.rds"))
   return(expanded)
 }

### 2021-08-30 data-raw/R/hydra_pull_GB_lengths.R

@@ -43,7 +43,7 @@ hydra_pull_GB_lengths <- function(channel, area=c(cfdbs::EPUs$data$GB,537)) {
 
   sampleLengths <- dplyr::as_tibble(sampleLengths)
 
-  saveRDS(sampleLengths,here::here("data-raw","hydra_lengths.rds"))
+  saveRDS(sampleLengths,here::here("data-raw/data","GB_hydra_lengths.rds"))
   
   return(sampleLengths)
 

### 2021-09-01 data-raw/R/create_focal_species.R
rename r files to start wwith "create" if they are used to create data

@@ -20,7 +20,7 @@
 
 library(magrittr)
 
-get_focal_species <- function(channel,saveToData=T) {
+create_focal_species <- function(channel,saveToData=T) {
   
   # create lookup table for species of interest
   lookup <- dbutils::create_species_lookup(channel,species=c(32,73,197,74,15,106,105,121,72,23,75),speciesType = "SVSPP")

### 2021-09-01 data-raw/R/create_hydra_main.R

@@ -15,7 +15,7 @@ source(here::here("data-raw/R","hydra_pull_GB_lengths.R"))
 source(here::here("data-raw/R","hydra_process_GB_comland.R"))
 source(here::here("data-raw/R","hydra_format.R"))
 
-hydra_main <- function(channel){
+create_hydra_main <- function(channel){
   
   #pull comland data
   comlandData <- readRDS(here::here("data-raw/data","hydra_comland_negear.rds"))

### 2021-09-01 R/catchAtLengthProportions.r
documentation for initial data files

@@ -1,0 +1,23 @@
+#' catch by size class, area, fleet
+#'
+#' For the species found in the Hydra model, the [proportion](https://noaa-edab.github.io/ms-keyrun/GBLandingsByLength.html) of total landings are reported by [fleet](https://noaa-edab.github.io/ms-keyrun/GBFleetDefinitions.html), year, and by [sizeclass](https://noaa-edab.github.io/ms-keyrun/GBLandingsByLength.html)
+#'
+#' @format A data frame with 3 columns
+#'
+#' \describe{
+#'   \item{fleet}{Name of [fleet](https://noaa-edab.github.io/ms-keyrun/GBFleetDefinitions.html)}
+#'   \item{area}{Area in which landings are reported}
+#'   \item{year}{year in which landings reported}
+#'   \item{species_itis}{itis code}
+#'   \item{type}{type of data. 0 = landings}
+#'   \item{totalBiomass}{Total Biomass}
+#'   \item{sizeclass1}{proportion of total biomass attributed to size class 1 }
+#'   \item{sizeclass2}{proportion of total biomass attributed to size class 1}
+#'   \item{sizeclass3}{proportion of total biomass attributed to size class 1}
+#'   \item{sizeclass4}{proportion of total biomass attributed to size class 1}
+#'   \item{sizeclass5}{proportion of total biomass attributed to size class 1}
+#'
+#' }
+#'
+#'
+"catchAtLengthProportions"

### 2021-09-01 R/fleets.r

@@ -1,0 +1,20 @@
+#' Fleet designations based on NEGEAR2 codes
+#'
+#' NEGEAR2 codes are assigned to fleets based on which species are landed. 
+#' The methodology and the decisions made to create this data set can be found below
+#'
+#' @format A data frame with 3 columns
+#'
+#' \describe{
+#'   \item{fleetName}{Name of fleet}
+#'   \item{NEGEAR2}{Gear codes associated with each fleet}
+#'   \item{hydraFleets}{Fleet aggregation to satify Hydra multispecies model}
+#'
+#' }
+#'
+#' @section Methodology:
+#'
+#' This data set was created based on clustering [methods](https://noaa-edab.github.io/ms-keyrun/GBFleetDefinitions.html).
+#'
+#'
+"fleets"

### 2021-09-01 R/focalSpecies.r

@@ -1,0 +1,19 @@
+#' The focal species in mskeyrun project
+#'
+#' Species name, latin name, and various related species codes
+#'
+#' @format A data frame
+#'
+#' \describe{
+#' \item{NESPP3}{species code found in commercial fishing database}
+#' \item{SVSPP}{species code found in bottom trawl survey database}
+#' \item{NAFOSPP}{species code in nafo database}
+#' \item{SPECIES_ITIS}{itis code}
+#' \item{COMMON_NAME}{common name from commercial fishing database}
+#' \item{COMNAME}{common name from bottom trawl survey fishing database}
+#' \item{SCIENTIFIC_NAME}{scientific name of species}
+#'
+#' }
+#'
+#'
+"focalSpecies"

### 2022-08-24 data-raw/R/create_lw_table.R
length weight parameter estimation - RMD describing method, and results (plots) - create_lw_table.r to assemble parameter estimates for all species - clean_lw_data.r to remove spurious len/weight pairs

@@ -1,0 +1,98 @@
+#'Read in survey length weight data and create a table
+#'
+#'
+#'
+
+library(magrittr)
+#channel <- dbutils::connect_to_database("sole","abeet")
+source(here::here("data-raw/R/clean_lw_data.r"))
+
+
+create_lw_table <- function(overwrite=F){
+
+  # filter its
+  itis <- mskeyrun::focalSpecies %>% 
+    dplyr::select(SPECIES_ITIS,SVSPP,LongName) %>% 
+    dplyr::filter(!(LongName == "Pollock")) %>%
+    dplyr::distinct()
+  
+  lengthWeightAgeData <- clean_lw_data()
+  
+  # format
+  lengthWeightAge <- lengthWeightAgeData %>%
+    dplyr::mutate(YEAR = as.numeric(substring(CRUISE6,1,4)),
+                  SVSPP = as.integer(SVSPP)) 
+  
+  # strata for Georges Bank
+  GB  <- c(1090, 1130:1210, 1230, 1250, 3460, 3480, 3490, 3520:3550)
+  plist <- list()
+  paramTable <- NULL
+  for (i in 1:length(itis$SVSPP)) {
+    isp <- itis$SVSPP[i]
+    speciesName <- itis$LongName[i]
+
+    # filter by species and weight
+    # may need to filter anomylous observations
+    lengthWeight <- lengthWeightAge %>% 
+      dplyr::filter(SVSPP==isp,
+                    INDWT > 0) %>% 
+      dplyr::mutate(INDWT = INDWT * 1000) %>% # convert to grams
+      dplyr::mutate(SEX = as.factor(SEX)) 
+    
+    # GB stocks use length weight data from GB only
+    if (any(isp %in% c(73,106,105))) { # cod, ytflounder wflounder
+      lengthWeight %>% 
+        dplyr::mutate(STRATUM = as.integer(STRATUM)) %>%
+        dplyr::filter(STRATUM %in% GB) 
+    }
+    #dplyr::filter(LENGTH > 80 & INDWT < 3)
+    
+    # fit length weight relationships
+    fits <- mscatch::fit_length_weight(lengthWeight,speciesName,speciesRules=NULL,suppressMessages=T)
+    # estimated variance parameter
+    sigma <- sqrt(fits$params$SINGLE$var)
+    palpha <- as.numeric(exp(fits$params$SINGLE$logAlpha))
+    pbeta <- as.numeric(fits$params$SINGLE$betas)
+    
+    # predicted weight
[…]

### 2022-08-24 data-raw/R/create_lw_table.R
added sigma as a column in the realBiolPar data file

@@ -83,7 +83,7 @@ create_lw_table <- function(overwrite=F){
                     Name = modelName,
                     WLa = alpha,
                     WLb = beta) %>% 
-      dplyr::select(ModSim,Code,Name,WLa,WLb) %>%
+      dplyr::select(ModSim,Code,Name,WLa,WLb,sigma) %>%
       dplyr::distinct() 
                           
     realBiolPar <- paramTable

### 2022-08-24 R/realBiolPar.R
add data r description for realBiolPar data

@@ -1,0 +1,22 @@
+#' The focal species biological parameters in mskeyrun project
+#'
+#' Length (cm) weight (g) parameters based on bottom trawl survey data. See details below
+#' 
+#' Model: W = aL^b e^Z
+#' where Z ~ IIDN(0,sigma^2)
+#' 
+#' E(W) = aL^b e^((sigma^2)/2)
+#' 
+#' @format A data frame
+#'
+#' \describe{
+#'\item{ModSim}{Atlantis model name and simulation id}
+#'\item{Code}{Itis code for species}
+#'\item{Name}{Common name for species}
+#'\item{WLa}{Weight-Length equation parameter a, W = aL^b}
+#'\item{WLb}{Weight-Length equation parameter b, W = aL^b}
+#'\item{sigma}{Estimated standard deviation of the Weight-Length equation parameter W = aL^b}
+#' }
+#'
+#'
+"realBiolPar"

### 2022-08-24 R/realSurveyLencomp.r
added survey biomass and numbers by length summed over tows for a given year and season - realSurveyLennumcomp.rda (renamed) - renamed script to create this to realSurveyLennumcomp.r

@@ -1,23 +1,0 @@
-#' Survey length composition for the mskeyrun project based on bottom trawl survey data
-#'
-#' Data pulled from the NEFSC bottom trawl survey database. Data pulled using 
-#'  \code{\link{survdat}} package.
-#' 
-#' @format A data frame
-#'
-#' \describe{
-#'\item{ModSim}{Atlantis model name and simulation id, or Actual data}
-#'\item{year}{year in which survey was undertaken}
-#'\item{season}{season in which survey was undertaken}
-#'\item{Code}{Species itis code}
-#'\item{Name}{Common name for species}
-#'\item{fishery}{fishery name}
-#'\item{lenbin}{recorded 1 cm length bin}
-#'\item{variable}{metric recorded, biomass}
-#'\item{value}{value of the variable}
-#'\item{units}{units of the variable}
-#'
-#' }
-#'
-#'
-"realSurveyLencomp"

### 2022-08-24 R/realSurveyLennumcomp.r

@@ -1,0 +1,23 @@
+#' Survey length composition for the mskeyrun project based on bottom trawl survey data
+#'
+#' Data pulled from the NEFSC bottom trawl survey database. Data pulled using 
+#'  \code{\link{survdat}} package.
+#' 
+#' @format A data frame
+#'
+#' \describe{
+#'\item{ModSim}{Atlantis model name and simulation id, or Actual data}
+#'\item{year}{year in which survey was undertaken}
+#'\item{season}{season in which survey was undertaken}
+#'\item{Code}{Species itis code}
+#'\item{Name}{Common name for species}
+#'\item{fishery}{fishery name}
+#'\item{lenbin}{recorded 1 cm length bin}
+#'\item{variable}{metric recorded, biomass & numbers}
+#'\item{value}{value of the variable, summed over all tows}
+#'\item{units}{units of the variable}
+#'
+#' }
+#'
+#'
+"realSurveyLennumcomp"

### 2022-08-24 data-raw/R/create_real_survey_lennumcomp.R
added numbers variable to data set of biomass at length. renamed the file

@@ -8,9 +8,9 @@
 
 library(magrittr)
 
-create_real_survey_lencomp <- function(convertKGtoMT=F, overwrite=F) {
+create_real_survey_lennumcomp <- function(convertKGtoMT=F, overwrite=F) {
   
-  if (convertKGtoMT  == T) {
+  if (convertKGtoMT) {
     scalar <- 1000
     unitsLabel <- "metric tons"
   } else {
@@ -50,7 +50,7 @@ create_real_survey_lencomp <- function(convertKGtoMT=F, overwrite=F) {
   # aggregate sampled data to find total weight of subsampled species per tow
   # This is the total weight of all lengthed fish per tow
   sampledTow <- survdatLWpull$survdat %>% 
-    dplyr::select(SVSPP,YEAR,SEASON,CRUISE6,STATION,STRATUM,TOW,LENGTH,WGTLEN) %>%
+    dplyr::select(SVSPP,YEAR,SEASON,CRUISE6,STATION,STRATUM,TOW,LENGTH,NUMLEN,WGTLEN) %>%
     dplyr::filter(SVSPP %in% svspps,
                   LENGTH >0) %>%
     dplyr::arrange(SVSPP,YEAR,SEASON,CRUISE6,STATION,STRATUM,TOW) %>%
@@ -75,28 +75,37 @@ create_real_survey_lencomp <- function(convertKGtoMT=F, overwrite=F) {
     dplyr::mutate(expansionFactor = towWeight/towSampleWeight)
   
   # apply this expansion factor to the weights of lengthed fish to "scale" up/down the biomass per length class
-  realSurveyLencomp <- dplyr::left_join(db,towlw,by = c("SVSPP", "YEAR", "SEASON", "CRUISE6", "STATION", "STRATUM", "TOW")) %>%
+  # also apply expansion factor to numbers at length
+  realSurveyLennumcomp <- dplyr::left_join(db,towlw,by = c("SVSPP", "YEAR", "SEASON", "CRUISE6", "STATION", "STRATUM", "TOW")) %>%
+    dplyr::mutate(expNum = expansionFactor * NUMLEN) %>% 
     dplyr::mutate(weight = expansionFactor * WGTLEN) %>% 
     dplyr::left_join(.,species, by = "SVSPP") %>%
     dplyr::mutate(ModSim = "Actual",
-                  fishery = "demersal",
-                  variable = "biomass",
-                  units = unitsLabel) %>%
[…]

### 2022-08-24 data-raw/R/create_sim_startpars.R
WIP starting N @len and true age, avg rec

@@ -1,0 +1,166 @@
+#' Read in initial age and recruitment and save as rda
+#' 
+#' atlantosom output is accessed and true values pulled for par inputs
+#' 
+#'@param atlmod configuration file specifying Atlantis simulation model filenames 
+#'and locations  
+#'@param saveToData Boolean. Export to data folder (Default = T)
+#'
+#'@return A tibble (Also written to \code{data} folder)
+#'\item{ModSim}{Atlantis model name and simulation id}
+#'\item{year}{year simulated survey conducted}
+#'\item{Code}{Atlantis model three letter code for functional group}
+#'\item{Name}{Atlantis model common name for functional group}
+#'\item{agecl}{age class of Atlantis functional group}
+#'\item{lenbin}{1 cm length bin, lower limit}
+#'\item{variable}{number at length (Natlen)}
+#'\item{value}{value of the variable}
+#'\item{units}{units of the variable}
+#'
+
+library(magrittr)
+
+create_sim_startpars <- function(atlmod,fitstart=NULL,fitend=NULL,saveToData=T) {
+
+  # input is path to model config file for atlantisom
+  source(atlmod)
+  
+  # path for survey and fishery config files
+  cfgpath <- stringr::str_extract(atlmod, ".*config")
+  
+  #works because atlantis directory named for model and simulation
+  modpath <- stringr::str_split(d.name, "/", simplify = TRUE)
+  modsim <- modpath[length(modpath)]
+  
+  # read true list with run and biol pars, etc
+  omlist_ss <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
+  
+  # model timesteps, etc from omdimensions script
+  source(paste0(cfgpath,"/omdimensions.R"), local = TRUE)
+  
+  #Number of years
+  nyears <- omlist_ss$runpar$nyears
+  total_sample <- omlist_ss$runpar$tstop/omlist_ss$runpar$outputstep
+  
+  # user specified fit start and times if different from full run
+  fitstartyr <- ifelse(!is.null(fitstart), fitstart-1, 0)
+  fitendyr <- ifelse(!is.null(fitend), fitend, total_sample)
+  
+  atlantis_full <- c(1:total_sample)  
+  mod_burnin <- fitstartyr*stepperyr+1
+  fit_nyears <- fitendyr-fitstartyr
[…]

### 2022-08-24 data-raw/build_simdata.R

@@ -45,5 +45,5 @@ create_sim_survey_agelen(atlmod, fitstart=40, fitend=120)
 
 create_sim_percapconsumption(atlmod, fitstart=40, fitend=120)
 
-
+create_sim_startpars(atlmod, fitstart=40, fitend=120)
 

### 2022-09-15 data-raw/R/create_real_dietcomp.R
WIP GB diet data for mskeyrun

@@ -1,0 +1,553 @@
+#' Create real diet datasets for mskeyrun
+#' 
+#' Diet dataset built by Brian Smith on github is accessed, mean wt and proportion 
+#' of prey in diet by species with variance is calculated
+#' 
+#' Modified \code{get_diet} function by Sean Hardison, original here
+#' https://github.com/NOAA-EDAB/ECSA/blob/master/R/get_diet.R
+#' Creates a dataset of annual and seasonal weighted diet composition with sample sizes.
+#' Datasets are for a selected species and set of seasonal survey strata.
+#' Need to clearly define what each data column is, render to tidy data.
+#' This is a first draft that will be revised with dplyr functions later.
+#' n.b., Original file from Brian Smith, August 23 2018 was allwt_nstoms.R, renamed as get_diet.R
+#'
+#' 
+#'@param focalspp mskeyrun dataset, output of \code\link{{create_focal_species}} or vector of ALL CAPS predator scientific names
+#'@param survstrat list of NEFSC trawl survey strata
+#'@param saveToData Boolean. Export to data folder (Default = T)
+#'
+#'@return A tibble (Also written to \code{data} folder)
+#'dataframe with variables svspp, year, season, meansw, num_tows, variance, cv, prey, totwt, relmsw, ci, relci, nstom
+#'\item{ModSim}{Atlantis model name and simulation id}
+#'\item{year}{year diet data collected}
+#'\item{Code}{SPECIES_ITIS code for predator species}
+#'\item{Name}{Common name for predator species}
+#'\item{season}{season diet data collected}
+#'\item{prey}{Common name for prey item}
+#'\item{variable}{mean prey item weight in year/species/season (meansw),
+#'                variance of meansw (variance),
+#'                cv of meansw (cv),
+#'                confidence interval for mean prey weight in diet (ci),
+#'                total prey weight in year/species/season (totwt),
+#'                percent of prey item in diet in year/species/season (relmsw),
+#'                confidence interval for percent prey in diet (relci),
+#'                number of tows in year/species/season (num_tows),
+#'                number of stomachs in year/species/season (nstom)}
+#'\item{value}{value of the variable}
+#'\item{units}{units of the variable; grams for meansw, variance, ci, totwt,
[…]

### 2022-09-15 data-raw/R/create_real_lendietcomp.R

@@ -1,0 +1,566 @@
+#' Create real diet datasets for mskeyrun-hydra
+#' 
+#' Diet dataset built by Brian Smith on github is accessed, mean wt and proportion 
+#' of prey in diet by predator species and length with variance is calculated
+#' 
+#' Modified \code{get_diet} function by Sean Hardison, original here
+#' https://github.com/NOAA-EDAB/ECSA/blob/master/R/get_diet.R
+#' Creates a dataset of annual and seasonal weighted diet composition with sample sizes.
+#' Datasets are for a selected species and set of seasonal survey strata.
+#' Need to clearly define what each data column is, render to tidy data.
+#' This is a first draft that will be revised with dplyr functions later.
+#' n.b., Original file from Brian Smith, August 23 2018 was allwt_nstoms.R, renamed as get_diet.R
+#'
+#' 
+#'@param focalspp mskeyrun dataset, output of \code\link{{create_focal_species}} or vector of ALL CAPS predator scientific names
+#'@param survstrat list of NEFSC trawl survey strata
+#'@param saveToData Boolean. Export to data folder (Default = T)
+#'
+#'@return A tibble (Also written to \code{data} folder)
+#'dataframe with variables svspp, year, season, meansw, num_tows, variance, cv, prey, totwt, relmsw, ci, relci, nstom
+#'\item{ModSim}{Atlantis model name and simulation id}
+#'\item{year}{year diet data collected}
+#'\item{Code}{SPECIES_ITIS code for predator species}
+#'\item{Name}{Common name for predator species}
+#'\item{season}{season diet data collected}
+#'\item{pdlen}{predator length bin (1 cm)}
+#'\item{prey}{Common name for prey item}
+#'\item{variable}{mean prey item weight in year/species/season (meansw),
+#'                variance of meansw (variance),
+#'                cv of meansw (cv),
+#'                confidence interval for mean prey weight in diet (ci),
+#'                total prey weight in year/species/season (totwt),
+#'                percent of prey item in diet in year/species/season (relmsw),
+#'                confidence interval for percent prey in diet (relci),
+#'                number of tows in year/species/season (num_tows),
+#'                number of stomachs in year/species/season (nstom)}
+#'\item{value}{value of the variable}
+#'\item{units}{units of the variable; grams for meansw, variance, ci, totwt,
[…]

### 2022-09-15 data-raw/build_dietdata.R

@@ -1,0 +1,17 @@
+# build diet datasets for mskeyrun
+# Sarah's notes
+# source create_real_dietcomp.R and create_real_lendietcomp.R
+# run the following
+
+library(magrittr)
+
+GBsurvstrata  <- c(1090, 1130:1210, 1230, 1250, 3460, 3480, 3490, 3520:3550)
+
+focalspp <- mskeyrun::focalSpecies %>%
+  dplyr::filter(modelName != "Pollock") %>% # not using in these models
+  dplyr::mutate(Name = modelName)
+
+create_real_dietcomp(focalspp, GBsurvstrata)
+
+create_real_lendietcomp(focalspp, GBsurvstrata)
+

### 2022-09-15 R/surveyDietcomp.R
still checking datasets and building vignette

@@ -1,0 +1,29 @@
+#' The survey diet compositions for the focal species within the mskeyrun project 
+#'
+#' Survey diet data from Northeast Fisheries Science Center's Bottom Trawl Survey, 1973-2019.
+#'
+#' @format A data frame
+#' 
+#'  \describe{
+#'\item{ModSim}{Atlantis model name and simulation id}
+#'\item{year}{year diet data collected}
+#'\item{Code}{SPECIES_ITIS code for predator species}
+#'\item{Name}{Common name for predator species}
+#'\item{season}{season diet data collected}
+#'\item{prey}{Common name for prey item}
+#'\item{variable}{mean prey item weight in year/species/season (meansw),
+#'                variance of meansw (variance),
+#'                cv of meansw (cv),
+#'                confidence interval for mean prey weight in diet (ci),
+#'                total prey weight in year/species/season (totwt),
+#'                percent of prey item in diet in year/species/season (relmsw),
+#'                confidence interval for percent prey in diet (relci),
+#'                number of tows in year/species/season (num_tows),
+#'                number of stomachs in year/species/season (nstom)}
+#'\item{value}{value of the variable}
+#'\item{units}{units of the variable; grams for meansw, variance, ci, totwt,
+#'             percent for relmsw, relci, numbers for num_tows, nstom, unitless cv}
+#' }
+#' 
+"surveyDietcomp"
+

### 2022-09-15 R/surveyLenDietcomp.R

@@ -1,0 +1,30 @@
+#' The survey diet compositions for the focal species within the mskeyrun project 
+#'
+#' Survey diet data from Northeast Fisheries Science Center's Bottom Trawl Survey, 1973-2019.
+#'
+#' @format A data frame
+#' 
+#'  \describe{
+#'\item{ModSim}{Atlantis model name and simulation id}
+#'\item{year}{year diet data collected}
+#'\item{Code}{SPECIES_ITIS code for predator species}
+#'\item{Name}{Common name for predator species}
+#'\item{season}{season diet data collected}
+#'\item{pdlen}{predator length bin (1 cm)}
+#'\item{prey}{Common name for prey item}
+#'\item{variable}{mean prey item weight in year/species/season (meansw),
+#'                variance of meansw (variance),
+#'                cv of meansw (cv),
+#'                confidence interval for mean prey weight in diet (ci),
+#'                total prey weight in year/species/season (totwt),
+#'                percent of prey item in diet in year/species/season (relmsw),
+#'                confidence interval for percent prey in diet (relci),
+#'                number of tows in year/species/season (num_tows),
+#'                number of stomachs in year/species/season (nstom)}
+#'\item{value}{value of the variable}
+#'\item{units}{units of the variable; grams for meansw, variance, ci, totwt,
+#'             percent for relmsw, relci, numbers for num_tows, nstom, unitless cv}
+#' }
+#' 
+"surveyLenDietcomp"
+

### 2022-09-15 data-raw/R/create_real_dietcomp.R

@@ -48,7 +48,7 @@ create_real_dietcomp <- function(focalspp, survstrat, saveToData = T){
   
   ## Filter out only good stomachs for strata and species
   allsum_raw <- allfh %>%
-    dplyr::filter(pdscinam %in% focalspp$SCIENTIFIC_NAME,
+    dplyr::filter(pdscinam %in% unique(focalspp$SCIENTIFIC_NAME),
                   stratum %in% survstrat,
                   pynam != 'BLOWN', 
                   pynam != 'PRESERVED',
@@ -541,13 +541,13 @@ create_real_dietcomp <- function(focalspp, survstrat, saveToData = T){
                                     variable %in% c("cv") ~ "unitless",
                                     TRUE ~ as.character("NA")))
   
-  SurveyDietcomp <- survdiet
+  surveyDietcomp <- survdiet
   
   if (saveToData) {
     
-    usethis::use_data(SurveyDietcomp, overwrite = TRUE)
+    usethis::use_data(surveyDietcomp, overwrite = TRUE)
   }
   
-  return(SurveyDietcomp)
+  return(surveyDietcomp)
   
 }

### 2022-09-15 data-raw/R/create_real_lendietcomp.R

@@ -50,7 +50,7 @@ create_real_lendietcomp <- function(focalspp, survstrat, saveToData = T){
   load(url("https://github.com/Laurels1/Condition/raw/master/data/allfh.RData"))
   
   survdietraw <- allfh %>%
-    dplyr::filter(pdscinam %in% focalspp$SCIENTIFIC_NAME,
+    dplyr::filter(pdscinam %in% unique(focalspp$SCIENTIFIC_NAME),
                   stratum %in% survstrat,
                   pynam != 'BLOWN', 
                   pynam != 'PRESERVED',
@@ -555,12 +555,12 @@ create_real_lendietcomp <- function(focalspp, survstrat, saveToData = T){
                                     variable %in% c("cv") ~ "unitless",
                                     TRUE ~ as.character("NA")))
   
-  SurveyLenDietcomp <- survdietlen
+  surveyLenDietcomp <- survdietlen
   
   if (saveToData) {
     
-    usethis::use_data(SurveyLenDietcomp, overwrite = TRUE)
+    usethis::use_data(surveyLenDietcomp, overwrite = TRUE)
   }
   
-  return(SurveyLenDietcomp)  
+  return(surveyLenDietcomp)  
 }

### 2022-09-15 data-raw/build_dietdata.R

@@ -9,7 +9,9 @@ GBsurvstrata  <- c(1090, 1130:1210, 1230, 1250, 3460, 3480, 3490, 3520:3550)
 
 focalspp <- mskeyrun::focalSpecies %>%
   dplyr::filter(modelName != "Pollock") %>% # not using in these models
-  dplyr::mutate(Name = modelName)
+  dplyr::select(-NESPP3) %>%
+  dplyr::mutate(Name = modelName) %>%
+  dplyr::distinct()
 
 create_real_dietcomp(focalspp, GBsurvstrata)
 

### 2022-09-16 R/surveyDietcomp.R
streamline dietcomp data, update documentation

@@ -13,12 +13,12 @@
 #'\item{prey}{Common name for prey item}
 #'\item{variable}{mean prey item weight in year/species/season (meansw),
 #'                variance of meansw (variance),
+#'                number of tows in year/species/season/prey (num_tows),
 #'                cv of meansw (cv),
 #'                confidence interval for mean prey weight in diet (ci),
 #'                total prey weight in year/species/season (totwt),
 #'                percent of prey item in diet in year/species/season (relmsw),
 #'                confidence interval for percent prey in diet (relci),
-#'                number of tows in year/species/season (num_tows),
 #'                number of stomachs in year/species/season (nstom)}
 #'\item{value}{value of the variable}
 #'\item{units}{units of the variable; grams for meansw, variance, ci, totwt,

### 2022-09-16 R/surveyLenDietcomp.R

@@ -12,15 +12,15 @@
 #'\item{season}{season diet data collected}
 #'\item{pdlen}{predator length bin (1 cm)}
 #'\item{prey}{Common name for prey item}
-#'\item{variable}{mean prey item weight in year/species/season (meansw),
+#'\item{variable}{mean prey item weight in year/species/season/pdlen (meansw),
 #'                variance of meansw (variance),
+#'                number of tows in year/species/season/prey/pdlen (num_tows),
 #'                cv of meansw (cv),
 #'                confidence interval for mean prey weight in diet (ci),
-#'                total prey weight in year/species/season (totwt),
-#'                percent of prey item in diet in year/species/season (relmsw),
+#'                total prey weight in year/species/season/pdlen (totwt),
+#'                percent of prey item in diet in year/species/season/pdlen (relmsw),
 #'                confidence interval for percent prey in diet (relci),
-#'                number of tows in year/species/season (num_tows),
-#'                number of stomachs in year/species/season (nstom)}
+#'                number of stomachs in year/species/season/pdlen (nstom)}
 #'\item{value}{value of the variable}
 #'\item{units}{units of the variable; grams for meansw, variance, ci, totwt,
 #'             percent for relmsw, relci, numbers for num_tows, nstom, unitless cv}

### 2022-09-16 data-raw/R/create_real_dietcomp.R

@@ -26,12 +26,12 @@
 #'\item{prey}{Common name for prey item}
 #'\item{variable}{mean prey item weight in year/species/season (meansw),
 #'                variance of meansw (variance),
+#'                number of tows in year/species/season/prey (num_tows),
 #'                cv of meansw (cv),
 #'                confidence interval for mean prey weight in diet (ci),
 #'                total prey weight in year/species/season (totwt),
 #'                percent of prey item in diet in year/species/season (relmsw),
 #'                confidence interval for percent prey in diet (relci),
-#'                number of tows in year/species/season (num_tows),
 #'                number of stomachs in year/species/season (nstom)}
 #'\item{value}{value of the variable}
 #'\item{units}{units of the variable; grams for meansw, variance, ci, totwt,
@@ -537,11 +537,25 @@ create_real_dietcomp <- function(focalspp, survstrat, saveToData = T){
                         values_to = "value") %>%
     dplyr::mutate(units = case_when(variable %in% c("meansw", "variance","totwt","ci") ~ "grams", 
                                     variable %in% c("relmsw", "relci") ~ "percent",
-                                    variable %in% c("num_tows","nstom") ~ "numbers",
+                                    variable %in% c("num_tows","nstom") ~ "number",
                                     variable %in% c("cv") ~ "unitless",
                                     TRUE ~ as.character("NA")))
   
-  surveyDietcomp <- survdiet
+  # reclassify variables that apply to pred not each prey
+  allprey <- survdiet %>%
+     dplyr::filter(variable %in% c("totwt", "nstom")) %>%
+     dplyr::mutate(prey = "ALL") %>%
+     dplyr::distinct()
+  
+  # remove duplicates and bind in reclassified pred variables
+  eachprey <- survdiet %>%
+    filter(variable != "totwt", 
+           variable != "nstom") 
+  
+ 
+  survdietfin <- bind_rows(eachprey, allprey)
+  
+  surveyDietcomp <- survdietfin
   
   if (saveToData) {
     

### 2022-09-16 data-raw/R/create_real_lendietcomp.R

@@ -25,15 +25,15 @@
 #'\item{season}{season diet data collected}
 #'\item{pdlen}{predator length bin (1 cm)}
 #'\item{prey}{Common name for prey item}
-#'\item{variable}{mean prey item weight in year/species/season (meansw),
+#'\item{variable}{mean prey item weight in year/species/season/pdlen (meansw),
 #'                variance of meansw (variance),
+#'                number of tows in year/species/season/prey/pdlen (num_tows),
 #'                cv of meansw (cv),
 #'                confidence interval for mean prey weight in diet (ci),
-#'                total prey weight in year/species/season (totwt),
-#'                percent of prey item in diet in year/species/season (relmsw),
+#'                total prey weight in year/species/season/pdlen (totwt),
+#'                percent of prey item in diet in year/species/season/pdlen (relmsw),
 #'                confidence interval for percent prey in diet (relci),
-#'                number of tows in year/species/season (num_tows),
-#'                number of stomachs in year/species/season (nstom)}
+#'                number of stomachs in year/species/season/pdlen (nstom)}
 #'\item{value}{value of the variable}
 #'\item{units}{units of the variable; grams for meansw, variance, ci, totwt,
 #'             percent for relmsw, relci, numbers for num_tows, nstom, unitless cv}
@@ -551,7 +551,7 @@ create_real_lendietcomp <- function(focalspp, survstrat, saveToData = T){
                         values_to = "value") %>%
     dplyr::mutate(units = case_when(variable %in% c("meansw", "variance","totwt","ci") ~ "grams", 
                                     variable %in% c("relmsw", "relci") ~ "percent",
-                                    variable %in% c("num_tows","nstom") ~ "numbers",
+                                    variable %in% c("num_tows","nstom") ~ "number",
                                     variable %in% c("cv") ~ "unitless",
                                     TRUE ~ as.character("NA")))
   

### 2022-09-16 data-raw/R/create_real_dietcomp.R
bugfix: correct 2 scientific names in diet datasets

@@ -552,6 +552,9 @@ create_real_dietcomp <- function(focalspp, survstrat, saveToData = T){
     filter(variable != "totwt", 
            variable != "nstom") 
   
+  # correct prey scientific names for mskeyrun species
+  eachprey$prey[eachprey$prey == "PLEURONECTES AMERICANUS"] <- "PSEUDOPLEURONECTES AMERICANUS"
+  eachprey$prey[eachprey$prey == "PLEURONECTES FERRUGINUS"] <- "LIMANDA FERRUGINEA"
  
   survdietfin <- bind_rows(eachprey, allprey)
   

### 2022-09-16 data-raw/R/create_real_lendietcomp.R

@@ -555,6 +555,11 @@ create_real_lendietcomp <- function(focalspp, survstrat, saveToData = T){
                                     variable %in% c("cv") ~ "unitless",
                                     TRUE ~ as.character("NA")))
   
+  # correct prey scientific names for mskeyrun species
+  survdietlen$prey[survdietlen$prey == "PLEURONECTES AMERICANUS"] <- "PSEUDOPLEURONECTES AMERICANUS"
+  survdietlen$prey[survdietlen$prey == "PLEURONECTES FERRUGINUS"] <- "LIMANDA FERRUGINEA"
+  
+  
   surveyLenDietcomp <- survdietlen
   
   if (saveToData) {

### 2022-09-19 data-raw/R/get_length_sample_n.R
add actual survey length samp size for hydra

@@ -1,0 +1,44 @@
+#' get numbers of sampled lengths for input into hydra
+#' based on create_real_survey_lennumcomp.R
+#' 
+
+library(magrittr)
+
+get_length_sample_n <- function(overwrite=T){
+  
+  survdatLWpull <- readRDS(here::here("data-raw/data/survdatLW.rds"))
+  
+  # filter GB region
+  GBStrata  <- c(1090, 1130:1210, 1230, 1250, 3460, 3480, 3490, 3520:3550)
+  
+  # get survey codes of focal species in mskeyrun
+  species <- mskeyrun::focalSpecies %>% 
+    dplyr::select(SVSPP,SPECIES_ITIS,LongName) %>% 
+    dplyr::filter(SVSPP != 75) %>% #omit pollock
+    dplyr::mutate(SPECIES_ITIS = as.double(SPECIES_ITIS)) %>%
+    dplyr::distinct()
+  
+  # get survey codes
+  svspps <- species$SVSPP
+  
+  newData <- survdatLWpull$survdat %>% 
+    dplyr::filter(SVSPP %in% svspps) %>%
+    dplyr::filter(STRATUM %in% GBStrata) %>%
+    dplyr::group_by(YEAR, SEASON, SVVESSEL, SVSPP) %>%
+    dplyr::summarise(lensampsize = sum(NUMLEN)) %>%
+    dplyr::left_join(species)
+  
+  surveyLenSampN <- newData
+  
+  if (overwrite) {
+    usethis::use_data(surveyLenSampN,overwrite=overwrite)
+  }
+  
+}
+
+# # visualize
+# library(ggplot2)
+# ggplot2::ggplot(surveyLenSampN, aes(x=YEAR, y=lensampsize, colour = SEASON)) + 
+#   geom_line()+ 
+#   facet_wrap(~LongName, scales = "free_y")+
+#   theme_bw()

### 2022-09-20 data-raw/R/get_fishery_length_sample_n.r
code to create length sample totals for Hydra

@@ -1,0 +1,47 @@
+#' Get fisheries length sample size by year
+#'
+#'
+
+get_fishery_length_sample_n <- function(overwrite=F) {
+
+  # identifies itis codes for all 10 species
+  itis <- mskeyrun::focalSpecies %>%
+    dplyr::distinct(SPECIES_ITIS) %>%
+    dplyr::filter(!(SPECIES_ITIS %in% c(164727,166774))) %>%
+    dplyr::pull()
+
+  maintab <- NULL
+  for(species in itis) {
+
+    # herring and skates data came from maine and survey respectively
+    # These datasets are not exported with mscatch
+    db <- eval(rlang::parse_expr(paste0("mscatch::sampleLengths_",species,"_GB")))
+
+    tab <- mskeyrun::focalSpecies %>%
+      dplyr::select(SPECIES_ITIS,SVSPP,LongName) %>%
+      dplyr::filter(SPECIES_ITIS == species) %>%
+      dplyr::distinct()
+
+
+    sptab <- db %>%
+      dplyr::group_by(YEAR) %>%
+      dplyr::summarise(lensampsize = as.double(sum(NUMLEN)),.groups = "drop") %>%
+      dplyr::mutate(YEAR= as.double(YEAR),
+                    SPECIES_ITIS = as.double(tab$SPECIES_ITIS),
+                    SVSPP = tab$SVSPP,
+                    LongName = tab$LongName) %>%
+      dplyr::select(YEAR,SVSPP,lensampsize,SPECIES_ITIS,LongName)
+
+    maintab <- rbind(maintab,sptab)
+
+  }
+
+  fisheryLenSampN <- maintab
+
+  if (overwrite) {
+    save(fisheryLenSampN,file=here::here("data/fisheryLenSampN.rda"))
+  }
+
+  return(fisheryLenSampN)
+
+}

### 2022-09-20 R/fisheryLenSampN.r
length samples by year and species for fishery

@@ -1,0 +1,16 @@
+#' Fishery sampled number of lengths for input into hydra model
+#'
+#' Number of fish measured by year, species
+#' for mskeyrun focal species
+#'
+#' #' @format A data frame
+#' 
+#'  \describe{
+#'\item{YEAR}{year diet data collected}
+#'\item{SVSPP}{SVSPP code for species}
+#'\item{lensampsize}{number of fish measured for length}
+#'\item{SPECIES_ITIS}{SPECIES_ITIS code for species}
+#'\item{LongName}{Common name for species}
+#'
+#' }
+'fisheryLenSampN'

### 2022-09-21 data-raw/R/format_herring_data.R
vignette to explain where data comes from for mscatch processing - functions to pull sample data (get_herring, get_skate, test_mscatch_data_pull_hydra_GB)

@@ -1,0 +1,103 @@
+#' reads in length data for herring
+#'
+#' Data pulled from maine data tables. use all lengths and apply
+#' to GB landings since a single stock
+#'
+#' Need to convert MT (herring database) to pounds (comlandr)
+#' Need to convert mm (herring lengths) to cm (svdbs)
+#'
+
+#data <- survdat::get_length_weight_data()
+library(magrittr)
+
+
+format_herring_data <- function(overwrite=F) {
+
+  kgTolbs <- 2.2046226218 #(1kg = 2.2046226218 lbs)
+
+  #herring data
+  herring <- readRDS(here::here("data-raw/data","herringData.rds"))
+
+  # length and weight data
+  lengthWeightData <- herring$length %>%
+    dplyr::mutate(tripid = "maine") %>%
+    dplyr::mutate(MARKET_CODE = "UN") %>%
+    dplyr::mutate(YEAR = as.integer(YEAR),
+                  MONTH = as.integer(MONTH)) %>%
+    dplyr::mutate(QTR = dplyr::case_when(MONTH %in% c(1:3) ~ 1,
+                                         MONTH %in% c(4:6) ~ 2 ,
+                                         MONTH %in% c(7:9) ~ 3,
+                                         MONTH %in% c(10:12) ~ 4,
+                                         TRUE ~ 0)) %>%
+    dplyr::mutate(LENGTH = round(SAMPLENGTH/10)) # convert mm to cm
+
+  # format for mscatch
+  lengthData <- lengthWeightData %>%
+    dplyr::select(YEAR,QTR,NEGEAR,MARKET_CODE,LENGTH,tripid) %>%
+    dplyr::group_by(YEAR,NEGEAR,QTR,MARKET_CODE,LENGTH,tripid) %>%
+    dplyr::summarise(NUMLEN = dplyr::n(),.groups="drop") %>%
+    dplyr::mutate(QTR = as.integer(QTR)) %>%
+    dplyr::filter(!is.na(LENGTH)) %>%
+    dplyr::select(YEAR,QTR,NEGEAR,MARKET_CODE,LENGTH,NUMLEN,tripid)
+
+
+  lengthDataSum <- lengthData %>%
+    dplyr::group_by(YEAR, QTR, NEGEAR, MARKET_CODE) %>%
+    dplyr::summarize(len_totalNumLen=sum(as.numeric(NUMLEN)),len_numLengthSamples=length(unique(tripid)),.groups="drop")
+
[…]

### 2022-09-21 data-raw/R/test_mscatch_data_pull_hydra_GB.R

@@ -1,0 +1,42 @@
+#' sample Data pull of all species in Hydra model on GB
+#'
+#'Sample data in the format we'll need from proper data pulls using Seans comland script.
+#'Landings pulled from "MV_CF_Landings", lengths pulled from "mv_cf_len"
+#'
+#'All missing areas need to dealt with prior to pulling species data.
+#'Not the case in this sample. We pull from GB
+#'
+#'@param channel obtained from logging into the sole database.
+#'
+#'@section Other species of interest:
+#'212 - atlantic mackerel - (K.Curti assessment)
+#'081 - cod (mike palmer assessment)
+#'
+
+#channel <- cfdbs::connect_to_database("server","user")
+source(here::here("data-raw","test_mscatch_data_pull.r"))
+
+test_mscatch_data_pull_hydra_GB <- function(channel){
+
+  library(magrittr)
+
+  hydraSpecies <- data.frame(oldName=c("Acod","Aherring","Amackerel","goosefish","haddock",
+                                       "silverhake","spinydog","winterfl","winterskate","yellowtailfl"),
+                             newName=c("ATLANTIC COD","ATLANTIC HERRING","ATLANTIC MACKEREL","GOOSEFISH","HADDOCK",
+                                       "SILVER HAKE","SPINY DOGFISH","WINTER FLOUNDER","WINTER SKATE","YELLOWTAIL FLOUNDER")) %>%
+    # hydraSpecies <- data.frame(oldName=c("Aherring","Amackerel", "silverhake","spinydog","winterfl","winterskate"),
+    #                            newName=c("ATLANTIC HERRING","ATLANTIC MACKEREL","SILVER HAKE","SPINY DOGFISH","WINTER FLOUNDER","WINTER SKATE")) %>%
+    dplyr::left_join(.,mscatch::speciesLookupTable,by=c("newName"="COMMON_NAME.y")) %>%
+    dplyr::select(oldName,newName,SPECIES_ITIS) %>%
+    dplyr::distinct()
+
+  GBArea <- c(cfdbs::EPUs$data$GB,537)
+
+  for (itis in unique(hydraSpecies$SPECIES_ITIS)) {
+    message(paste0("pulling: ",itis))
+    #fileName <- paste0(itis,"_GB")
+    test_mscatch_data_pull(channel,species=itis,species_itis = T, area = GBArea,areaLengths="all", stock="GB")
+  }
+
+
+}

### 2022-09-21 data-raw/R/make_species_datasets_for_mskeyrun.R
add function to run through mscatch process and generate resulting data - make_species_datasets_for_mskeyrun.r

@@ -1,0 +1,164 @@
+#' Creates the data sets for mskeyrun.
+#'
+#' All species have a fixed time interval, QTR, SEMESTER, YEAR.
+#' All species borrow samples.
+#' UNclassified use the same temporal aggregation
+#' All species use length data from survey to fit length weight relationships
+#' Landings from GB statistical areas
+#' Lengths from same area unless shelf wide stock, then lengths from larger area
+#'
+#' This alleviates the immediate need for coding the length expansion to deal
+#' with differing levels of aggregation that would arise from collapsing time
+#' based on sample availability
+#'
+#'
+
+#channel <- dbutils::connect_to_database("sole","abeet")
+
+library(magrittr)
+
+make_species_datasets_for_mskeyrun <- function(channel) {
+
+
+  itis <- mscatch::speciesLookupTable %>%
+    dplyr::distinct(SPECIES_ITIS) %>%
+    dplyr::filter(!(SPECIES_ITIS %in% c(164727,166774))) %>%
+    dplyr::pull()
+
+
+  for (species_itis in itis) {
+
+    speciesRules <- mscatch::get_species_object_mskeyrun(species_itis)
+
+    message(paste0("Processing ",speciesRules$speciesName))
+    #speciesRules$howAggregate <- "borrow"
+    #speciesRules$LengthWeightData <- "survey"
+
+    aggregate_to = toupper(speciesRules$temporalAggregation)
+    stock <- "GB"#speciesRules$stock
+    borrowLengths <- T
+
+    speciesName <- speciesRules$speciesName
+    speciessvspp <- speciesRules$SVSPP
+
+    outputDir <- here::here("output",paste0(speciesName[1],"_",stock,"_tt_",aggregate_to,"_",speciesRules$howAggregate))
+
+    if (!dir.exists(outputDir)){ # create directory to store exploratory/informational plots
+      dir.create(outputDir)
+    }
+
+    # landings from GB, lengths from NEUS wide
+    # All landings are in lbs (from stockeff)
+    landName <- rlang::parse_expr(paste0("mscatch::sampleData_",species_itis,"_GB"))
+    lengthName <- rlang::parse_expr(paste0("mscatch::sampleLengths_",species_itis,"_GB"))
+
+    # filter lengths based on spurious values. Will eventually do this during data pull
+    if (species_itis == 160617) {
+     lengthData <- eval(lengthName) %>% dplyr::filter(LENGTH < 700)
[…]

### 2022-09-23 data-raw/R/get_length_sample_n.R
bugfix: na removal in survey length samp size

@@ -25,7 +25,7 @@ get_length_sample_n <- function(overwrite=T){
     dplyr::filter(SVSPP %in% svspps) %>%
     dplyr::filter(STRATUM %in% GBStrata) %>%
     dplyr::group_by(YEAR, SEASON, SVVESSEL, SVSPP) %>%
-    dplyr::summarise(lensampsize = sum(NUMLEN)) %>%
+    dplyr::summarise(lensampsize = sum(NUMLEN, na.rm = TRUE)) %>%
     dplyr::left_join(species)
   
   surveyLenSampN <- newData

### 2022-09-23 data-raw/R/pull_comland.R
Added discards to data pull

@@ -8,16 +8,25 @@ library(here); library(data.table); library(comlandr)
 channel <- dbutils::connect_to_database(server = 'sole', uid = 'slucey')
 
 #Pull data
-comland <- comlandr::get_comland_data(channel, filterByArea = c(521, 522, 525, 
-                                                                526, 537, 538, 
-                                                                551, 552, 561,
-                                                                562), 
+landings <- comlandr::get_comland_data(channel, filterByYear = 1968:2019,
+                                       filterByArea = c(521, 522, 525, 526, 537, 
+                                                        538, 551, 552, 561, 562), 
                                       aggArea = T, aggGear = T,
                                       unkVar = c('MONTH', 'Fleet', 'EPU'),
                                       knStrata = c('NESPP3', 'YEAR', 'EPU',
                                                    'HY', 'QY', 'MONTH', 
                                                    'Fleet', 'TONCL1'))
-comland$comland <- comland$comland[EPU == 'GB', ]
+
+#Pull discard data
+channel2 <- dbutils::connect_to_database('nova', 'slucey')
+
+discards <- comlandr::get_comdisc_data(channel2, landings, aggArea = T, aggGear = T,
+                                       extendTS = T)
+
+#Filter to just Georges Bank
+landings$comland <- landings$comland[EPU == 'GB', ]
+discards$comdisc <- discards$comdisc[EPU == 'GB', ]
 
 #save data
-save(comland, file = here::here('data-raw', 'data', 'comland.RData'))
+save(landings, file = here::here('data-raw', 'data', 'landings.RData'))
+save(discards, file = here::here('data-raw', 'data', 'discards.RData'))

### 2022-09-23 data-raw/R/create_fishery_index.R
Added discards

@@ -3,14 +3,28 @@
 library(here); library(survdat); library(data.table); library(sf)
 
 #Load landings data
-load(here::here('data-raw', 'data', 'comland.RData'))
+load(here::here('data-raw', 'data', 'landings.RData'))
+
+#Load discards data
+load(here::here('data-raw', 'data', 'discards.RData'))
 
 #Sum over all catch variables for total catch
 spp <- as.numeric(unique(mskeyrun::focalSpecies$NESPP3))
-catchIndex <- comland$comland[NESPP3 %in% spp, .(value = sum(SPPLIVMT)),
+
+#Landings
+landIndex <- landings$comland[NESPP3 %in% spp, .(value = sum(SPPLIVMT)),
                               by = c('NESPP3', 'YEAR')]
-catchIndex[, variable := 'commerical landings']
-catchIndex[, units    := 'metric tons']
+landIndex[, variable := 'commerical landings']
+landIndex[, units    := 'metric tons']
+
+#Discards
+discIndex <- discards$comdisc[NESPP3 %in% spp, .(value = sum(DISMT, na.rm = T)),
+                              by = c('NESPP3', 'YEAR')]
+discIndex[, variable := 'commercial discards']
+discIndex[, units := 'metric tons']
+
+#Combine landings and discards
+catchIndex <- data.table::rbindlist(list(landIndex, discIndex))
 
 #Output to package
 usethis::use_data(catchIndex, overwrite = TRUE)

### 2022-09-23 data-raw/R/create_fishery_index.R
variable name update and documentation

@@ -14,7 +14,7 @@ spp <- as.numeric(unique(mskeyrun::focalSpecies$NESPP3))
 #Landings
 landIndex <- landings$comland[NESPP3 %in% spp, .(value = sum(SPPLIVMT)),
                               by = c('NESPP3', 'YEAR')]
-landIndex[, variable := 'commerical landings']
+landIndex[, variable := 'commercial landings']
 landIndex[, units    := 'metric tons']
 
 #Discards

### 2022-10-04 data-raw/R/create_real_fishery_lencomp.R
updated realFisheryLencomp to include abundance by length in addition to biomass

@@ -1,4 +1,4 @@
-#' Length comps
+#' Length abundance and biomass by length
 #'
 #'Read and format output from mscatch to mskeyrun
 #'
@@ -32,16 +32,19 @@ create_real_fishery_lencomp <- function(convertKGtoMT=T,outPath=NULL) {
 
   for (it in itis) {
     # expandedLandings in kgs
-    fileNm <- here::here("data-raw/data/",paste0("expandedLandings",it,".rds"))
-
-    if(file.exists(fileNm)) {
+    fileNmAge <- here::here("data-raw/data/",paste0("numbersatAge",it,".rds"))
+    fileNmLen <- here::here("data-raw/data/",paste0("expandedLandings",it,".rds"))
+    
+    if(file.exists(fileNmAge)) {
+      noAge <-  F
       message(paste0("Reading itis = ",it))
-      speciesData <- readRDS(fileNm)
+      speciesDataAge <- readRDS(fileNmAge)
     } else {
-      message(paste0("Couldn't find itis = ",it))
-      next
+      noAge <-  T
+      message(paste0("Couldn't find Age data for itis = ",it,". Reading in length data only"))
+      speciesDataLen <- readRDS(fileNmLen)
     }
-
+    
     name <- mscatch::speciesLookupTable %>%
       dplyr::filter(SPECIES_ITIS == it) %>%
       dplyr::pull(COMMON_NAME.y) %>%
@@ -49,20 +52,38 @@ create_real_fishery_lencomp <- function(convertKGtoMT=T,outPath=NULL) {
       abutils::capitalize_first_letter()
     ## process data
     #ModSim        year Code  Name         fishery lenbin variable value units
-    speciesFormat <- speciesData %>%
-      dplyr::group_by(YEAR,NEGEAR,LENGTH,species_itis) %>%
-      dplyr::summarise(value = sum(weight)/scalar,.groups="drop") %>%
-      dplyr::rename(year=YEAR,Code=species_itis,fishery=NEGEAR,lenbin = LENGTH) %>%
-      dplyr::mutate(Name = name,
-                    ModSim = "Actual",
-                    variable = "biomass",
-                    units=unitsLabel) %>%
-      dplyr::select(ModSim,year,Code,Name,fishery,lenbin,variable,value,units)
-
-
+    if (noAge) {
+      speciesFormat <- speciesDataLen %>%
[…]

### 2022-10-04 data-raw/R/get_fishery_length_sample_n.r
updated fisheryLenSampN to consist of length samples just from GB, even for species with a unit stock

@@ -15,8 +15,8 @@ get_fishery_length_sample_n <- function(overwrite=F) {
 
     # herring and skates data came from maine and survey respectively
     # These datasets are not exported with mscatch
-    db <- eval(rlang::parse_expr(paste0("mscatch::sampleLengths_",species,"_GB")))
-
+    db <- eval(rlang::parse_expr(paste0("mscatch::GB_Lengths_Only_",species)))
+    
     tab <- mskeyrun::focalSpecies %>%
       dplyr::select(SPECIES_ITIS,SVSPP,LongName) %>%
       dplyr::filter(SPECIES_ITIS == species) %>%

### 2022-10-05 data-raw/R/create_real_fishery_lencomp.R
fixes missing abundance for elasmobranch in realFisheryLencomp data and fixed to cod and winterflounder age data  caused by mscatch::calc_numbers_at_age

@@ -10,10 +10,10 @@
 library(magrittr)
 
 create_real_fishery_lencomp <- function(convertKGtoMT=T,outPath=NULL) {
-
+  
   # 1kg = 2.2046226218 lbs
-
-
+  
+  
   if (convertKGtoMT  == T) {
     scalar <- 1000
     unitsLabel <- "metric tons"
@@ -21,29 +21,21 @@ create_real_fishery_lencomp <- function(convertKGtoMT=T,outPath=NULL) {
     scalar <- 1
     unitsLabel <- "kilograms"
   }
-
-
+  
+  
   itis <- mscatch::speciesLookupTable %>%
     dplyr::distinct(SPECIES_ITIS) %>%
     dplyr::filter(!(SPECIES_ITIS %in% c(164727,166774))) %>%
     dplyr::pull()
-
+  
   realFisheryLencomp <- NULL
-
+  
   for (it in itis) {
-    # expandedLandings in kgs
-    fileNmAge <- here::here("data-raw/data/",paste0("numbersatAge",it,".rds"))
+    # expandedLandings in kgs. This is referencing mscatch dir structure
     fileNmLen <- here::here("data-raw/data/",paste0("expandedLandings",it,".rds"))
     
-    if(file.exists(fileNmAge)) {
-      noAge <-  F
-      message(paste0("Reading itis = ",it))
-      speciesDataAge <- readRDS(fileNmAge)
-    } else {
-      noAge <-  T
-      message(paste0("Couldn't find Age data for itis = ",it,". Reading in length data only"))
-      speciesDataLen <- readRDS(fileNmLen)
-    }
+    message(paste0("Reading itis = ",it))
+    speciesDataLen <- readRDS(fileNmLen)
     
     name <- mscatch::speciesLookupTable %>%
       dplyr::filter(SPECIES_ITIS == it) %>%
@@ -52,50 +44,35 @@ create_real_fishery_lencomp <- function(convertKGtoMT=T,outPath=NULL) {
       abutils::capitalize_first_letter()
     ## process data
     #ModSim        year Code  Name         fishery lenbin variable value units
-    if (noAge) {
-      speciesFormat <- speciesDataLen %>%
-        dplyr::group_by(YEAR,NEGEAR,LENGTH,species_itis) %>%
-        dplyr::summarise(value = sum(weight)/scalar,.groups="drop") %>%
-        dplyr::rename(year=YEAR,Code=species_itis,fishery=NEGEAR,lenbin = LENGTH) %>%
[…]

### 2023-06-12 data-raw/R/create_sim_fishery_index.R
start updates for new catch input

@@ -56,9 +56,14 @@ create_sim_fishery_index <- function(atlmod,fitstart=NULL,fitend=NULL,saveToData
   fit_nyears <- fitendyr-fitstartyr
   fit_ntimes <- fit_nyears*stepperyr
   fittimes <- atlantis_full[mod_burnin:(mod_burnin+fit_ntimes-1)]
-  #fit_timesteps <- seq(fittimes[stepperyr], max(fittimes), by=stepperyr) #last timestep
+  fit_timesteps <- seq(fittimes[stepperyr], max(fittimes), by=stepperyr) #last timestep
   #fit_years <- unique(floor(fittimes/stepperyr)) #from Christine's new sardine_config.R
   fittimes.days <- if(omlist_ss$runpar$outputstepunit=="days") fittimes*omlist_ss$runpar$outputstep
+  
+  # catch files from CATCH.nc are subannual and by fleet
+  # BUT have codes not species names,
+  # and time in model timesteps not days
+  # SO conversion needed
   
   
   # fishery cv lookup from config files
@@ -72,7 +77,8 @@ create_sim_fishery_index <- function(atlmod,fitstart=NULL,fitend=NULL,saveToData
   
   allcatch <- tibble::tibble()
   
-  for(f in names(catchbio_ss)){
+  # limit catchbio_ss to names in fcvlook
+  for(f in names(catchbio_ss)[names(catchbio_ss) %in% fcvlook$fishery]){
     catchbio <- catchbio_ss[[f]][[1]] %>%
       #dplyr::filter(time>0) %>%
       dplyr::filter(time %in% fittimes.days) %>%

### 2023-06-13 R/simCatchIndexSubannual.r
create subannual simulated fishery data

@@ -2,14 +2,15 @@
 #'
 #' atlantosom output is accessed and fishery data pulled over time
 #' simulated fishery catches are cumulative, not snapshots
-#' fishery total catch for the year is the sum of the months from 
-#' simCatchIndexSubannual
+#' fishery total catch for the year is the sum of the months
+#' fishery catch for the month is all catch reported since the last reporting month
 #'
 #' @format A data frame
 #'
 #' \describe{
 #'\item{ModSim}{Atlantis model name and simulation id}
 #'\item{year}{year simulated fishery conducted}
+#'\item{fishMonth}{month simulated fishery conducted}
 #'\item{Code}{Atlantis model three letter code for functional group}
 #'\item{Name}{Atlantis model common name for functional group}
 #'\item{fishery}{simulated fishery name}
@@ -20,4 +21,4 @@
 #' }
 #'
 #'
-"simCatchIndex"
+"simCatchIndexSubannual"

### 2023-06-13 data-raw/R/create_sim_fishery_index.R

@@ -1,6 +1,11 @@
-#' Read in survey data save as rda
+#' Read in fishery catch data save as rda
 #' 
-#' atlantosom output is accessed and surveys pulled over time
+#' atlantosom output is accessed and fishery data pulled over time
+#' simulated fishery catches are cumulative, not snapshots
+#' fishery total catch for the year is the sum of the months
+#' fishery catch for the month is all catch reported since the last reporting month
+#' 
+#' also makes annual aggregate dataset for backwards compatibility
 #' 
 #'@param atlmod configuration file specifying Atlantis simulation model filenames 
 #'and locations  
@@ -9,6 +14,7 @@
 #'@return A tibble (Also written to \code{data} folder)
 #'\item{ModSim}{Atlantis model name and simulation id}
 #'\item{year}{year simulated fishery conducted}
+#'\item{fishMonth}{month simulated fishery conducted}
 #'\item{Code}{Atlantis model three letter code for functional group}
 #'\item{Name}{Atlantis model common name for functional group}
 #'\item{fishery}{simulated fishery name}
@@ -47,6 +53,9 @@ create_sim_fishery_index <- function(atlmod,fitstart=NULL,fitend=NULL,saveToData
   nyears <- omlist_ss$runpar$nyears
   total_sample <- omlist_ss$runpar$tstop/omlist_ss$runpar$outputstep
   
+  # throw an error if fstepperyr is not equal to stepperyr
+  if(stepperyr != fstepperyr) stop("Error: check Atlantis timestep output for fishery")
+  
   # user specified fit start and times if different from full run
   fitstartyr <- ifelse(!is.null(fitstart), fitstart-1, 0)
   fitendyr <- ifelse(!is.null(fitend), fitend, total_sample)
@@ -65,7 +74,6 @@ create_sim_fishery_index <- function(atlmod,fitstart=NULL,fitend=NULL,saveToData
   # and time in model timesteps not days
   # SO conversion needed
   
-  
   # fishery cv lookup from config files
   fcvlook <- tibble::tibble()
   for(c in 1:length(fishcon)){
@@ -78,36 +86,51 @@ create_sim_fishery_index <- function(atlmod,fitstart=NULL,fitend=NULL,saveToData
   allcatch <- tibble::tibble()
   
   # limit catchbio_ss to names in fcvlook
[…]

### 2023-06-13 data-raw/R/create_sim_fishery_wtage.R
adds check for timestep

@@ -63,6 +63,9 @@ create_sim_fishery_wtage <- function(atlmod,fitstart=NULL,fitend=NULL,saveToData
   nyears <- omlist_ss$runpar$nyears
   total_sample <- omlist_ss$runpar$tstop/omlist_ss$runpar$outputstep
   
+  # throw an error if fstepperyr is not equal to stepperyr
+  if(stepperyr != fstepperyr) stop("Error: check Atlantis timestep output for fishery")
+  
   # user specified fit start and times if different from full run
   fitstartyr <- ifelse(!is.null(fitstart), fitstart-1, 0)
   fitendyr <- ifelse(!is.null(fitend), fitend, total_sample)

### 2023-06-13 data-raw/R/create_sim_fishery_lencomp_subannual.R
adds subannual fishery length comps

@@ -1,0 +1,114 @@
+#' Read in fishery data save as rda
+#' 
+#' atlantosom output is accessed and fishery catch pulled over time
+#' 
+#'@param atlmod configuration file specifying Atlantis simulation model filenames 
+#'and locations  
+#'@param saveToData Boolean. Export to data folder (Default = T)
+#'
+#'@return A tibble (Also written to \code{data} folder)
+#'\item{ModSim}{Atlantis model name and simulation id}
+#'\item{year}{year simulated fishery conducted}
+#'\item{fishMonth}{month simulated fishery conducted}
+#'\item{Code}{Atlantis model three letter code for functional group}
+#'\item{Name}{Atlantis model common name for functional group}
+#'\item{fishery}{simulated fishery name}
+#'\item{lenbin}{1 cm length bin, lower limit}
+#'\item{variable}{number at length (Natlen)}
+#'\item{value}{value of the variable}
+#'\item{units}{units of the variable}
+#'
+
+library(magrittr)
+
+create_sim_fishery_lencomp_subannual <- function(atlmod,fitstart=NULL,fitend=NULL,saveToData=T) {
+
+  # input is path to model config file for atlantisom
+  source(atlmod)
+  
+  # path for survey and fishery config files
+  cfgpath <- stringr::str_extract(atlmod, ".*config")
+  
+  #works because atlantis directory named for model and simulation
+  modpath <- stringr::str_split(d.name, "/", simplify = TRUE)
+  modsim <- modpath[length(modpath)]
+  
+  #read in survey annual age comp data
+  catchlen_ss <- atlantisom::read_savedfisheries(d.name, "catchLen")
+  
+  # get config files -- needed?
+  fvcon <- list.files(path=cfgpath, pattern = "*fishery*", full.names = TRUE)
+  
+  # read true list with run and biol pars, etc
+  omlist_ss <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
+  
+  # model timesteps, etc from omdimensions script
+  source(paste0(cfgpath,"/omdimensions.R"), local = TRUE)
+  
+  #Number of years
+  nyears <- omlist_ss$runpar$nyears
+  total_sample <- omlist_ss$runpar$tstop/omlist_ss$runpar$outputstep
+  
[…]

### 2023-06-13 data-raw/build_simdata.R

@@ -37,6 +37,8 @@ create_sim_survey_lencomp(atlmod, fitstart=40, fitend=120)
 
 create_sim_fishery_lencomp(atlmod, fitstart=40, fitend=120)
 
+create_sim_fishery_lencomp_subannual(atlmod, fitstart=40, fitend=120)
+
 create_sim_survey_dietcomp(atlmod, fitstart=40, fitend=120)
 
 create_sim_survey_bottemp(atlmod, fitstart=40, fitend=120)

### 2023-06-13 data-raw/build_simdata.R
adds subannual fishery wt at age info

@@ -25,7 +25,7 @@ create_sim_survey_info(atlmod)
 
 create_sim_survey_index(atlmod, fitstart=40, fitend=120)
 
-create_sim_fishery_index(atlmod, fitstart=40, fitend=120)
+create_sim_fishery_index(atlmod, fitstart=40, fitend=120) #creates subannual amd aggregate
 
 create_sim_survey_agecomp(atlmod, fitstart=40, fitend=120)
 
@@ -44,6 +44,8 @@ create_sim_survey_dietcomp(atlmod, fitstart=40, fitend=120)
 create_sim_survey_bottemp(atlmod, fitstart=40, fitend=120)
 
 create_sim_fishery_wtage(atlmod, fitstart=40, fitend=120)
+
+create_sim_fishery_wtage_subannual(atlmod, fitstart=40, fitend=120)
 
 create_sim_survey_wtage(atlmod, fitstart=40, fitend=120)
 

### 2023-06-13 R/simFisheryAgeLencomp.r
adds simulated fishery agecl length info

@@ -1,0 +1,22 @@
+#' The Atlantis-simulated fishery length composition for the mskeyrun project
+#'
+#' Simulated fishery length composition data from Atlantis model as sampled by atlantisom
+#'
+#' @format A data frame
+#'
+#' \describe{
+#'\item{ModSim}{Atlantis model name and simulation id}
+#'\item{year}{year simulated survey conducted}
+#'\item{Code}{Atlantis model three letter code for functional group}
+#'\item{Name}{Atlantis model common name for functional group}
+#'\item{fishery}{simulated fishery name}
+#'\item{agecl}{age class of Atlantis functional group}
+#'\item{lenbin}{1 cm length bin, lower limit}
+#'\item{variable}{number at length (Natlen)}
+#'\item{value}{value of the variable}
+#'\item{units}{units of the variable}
+#'
+#' }
+#'
+#'
+"simFisheryAgeLencomp"

### 2023-06-13 data-raw/R/create_sim_fishery_agelen.R

@@ -1,0 +1,113 @@
+#' Read in fishery data save as rda
+#' 
+#' atlantosom output is accessed and fishery catch pulled over time
+#' 
+#'@param atlmod configuration file specifying Atlantis simulation model filenames 
+#'and locations  
+#'@param saveToData Boolean. Export to data folder (Default = T)
+#'
+#'@return A tibble (Also written to \code{data} folder)
+#'\item{ModSim}{Atlantis model name and simulation id}
+#'\item{year}{year simulated survey conducted}
+#'\item{Code}{Atlantis model three letter code for functional group}
+#'\item{Name}{Atlantis model common name for functional group}
+#'\item{fishery}{simulated fishery name}
+#'\item{agecl}{age class of Atlantis functional group}
+#'\item{lenbin}{1 cm length bin, lower limit}
+#'\item{variable}{number at length (Natlen)}
+#'\item{value}{value of the variable}
+#'\item{units}{units of the variable}
+#'
+
+library(magrittr)
+
+create_sim_fishery_agelen <- function(atlmod,fitstart=NULL,fitend=NULL,saveToData=T) {
+
+  # input is path to model config file for atlantisom
+  source(atlmod)
+  
+  # path for survey and fishery config files
+  cfgpath <- stringr::str_extract(atlmod, ".*config")
+  
+  #works because atlantis directory named for model and simulation
+  modpath <- stringr::str_split(d.name, "/", simplify = TRUE)
+  modsim <- modpath[length(modpath)]
+  
+  #read in survey annual age comp data
+  catchlen_ss <- atlantisom::read_savedfisheries(d.name, "catchLen")
+  
+  # get config files -- needed?
+  fvcon <- list.files(path=cfgpath, pattern = "*fishery*", full.names = TRUE)
+  
+  # read true list with run and biol pars, etc
+  omlist_ss <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
+  
+  # model timesteps, etc from omdimensions script
+  source(paste0(cfgpath,"/omdimensions.R"), local = TRUE)
+  
+  #Number of years
+  nyears <- omlist_ss$runpar$nyears
+  total_sample <- omlist_ss$runpar$tstop/omlist_ss$runpar$outputstep
+  
[…]

### 2023-06-13 data-raw/build_simdata.R

@@ -51,6 +51,8 @@ create_sim_survey_wtage(atlmod, fitstart=40, fitend=120)
 
 create_sim_survey_agelen(atlmod, fitstart=40, fitend=120)
 
+create_sim_fishery_agelen(atlmod, fitstart=40, fitend=120)
+
 create_sim_percapconsumption(atlmod, fitstart=40, fitend=120)
 
 create_sim_startpars(atlmod, fitstart=40, fitend=120)

### 2023-06-13 data-raw/simulated-data/config/mssurvey_fall_fw.R
generate simulated food web model survey B and catch

@@ -1,0 +1,78 @@
+# Default survey configuration here has a range of efficiencies and selectivities
+# To emulate a range of species in a single multispecies survey
+# Also now happens in "spring" and "fall"
+# Need to define survey season, area, efficiency, selectivity
+
+# Survey name
+survey.name="BTS_fall_allbox_effic1_fw"
+
+#Atlantis model timestep corresponding to the true output--now from census_spec.R
+timestep <- stepperyr #5
+
+#Which atlantis timestep does the survey run in?--now from census_spec.R
+# with 5 output steps per year, 0 is Jan-Feb-midMar, 1 is midMar-Apr-May, 
+# 2 is June-July-midAug, 3 is midAug-Sept-Oct, 4 is Nov-Dec (ish)
+
+# No, timestep 0 is initial condition and should be ignored to align 
+# snapshots (biomass, numbers) with
+# cumulative outputs (fishery catch, numbers)
+
+# with 5 output steps per (non leap) year:
+# 1 is day 73, or 14 March
+# 2 is day 146, or 26 May
+# 3 is day 219, or 7 August
+# 4 is day 292, or 19 October
+# 5 is day 365, or 31 December
+
+survey_sample_time <- 3 # fall survey
+
+#The last timestep to sample
+total_sample <- noutsteps-1 #495
+
+#Vector of indices of survey times to pull
+survey_sample_full <- seq(survey_sample_time,
+                          total_sample, by=timestep)
+
+survtime <- survey_sample_full
+
+# survey area
+# should return all model areas
+survboxes <- allboxes
+
+# survey efficiency (q)
+# should return a perfectly efficient survey 
+surveffic <- data.frame(species=survspp,
+                     efficiency=rep(1.0,length(survspp)))
+
+# survey selectivity (agecl based)
+# this is by age class, need to change to use with ANNAGEBIO output
+#survselex <- data.frame(species=rep(survspp, each=n_age_classes),
+#                     agecl=rep(c(1:n_age_classes),length(survspp)),
+#                     selex=rep(1.0,length(survspp)*n_age_classes))
+
+# for annage output
+survselex <- data.frame(species=rep(names(annages), n_annages), #  
+                        agecl=unlist(sapply(n_annages,seq)),
+                        selex=rep(1.0,sum(n_annages)))
+
+survselex.agecl <- survselex 
+
+
+# effective sample size needed for sample_fish
+# this effective N is high but not equal to total for numerous groups
[…]

### 2023-06-13 data-raw/R/create_sim_fishery_index.R
correction, dont call mskeyrun here

@@ -116,7 +116,7 @@ create_sim_fishery_index <- function(atlmod,fitstart=NULL,fitend=NULL,saveToData
   simCatchIndexSubannual <- allcatch
   
   #build new annual index from this one
-  simCatchIndex <- mskeyrun::simCatchIndexSubannual %>%
+  simCatchIndex <- allcatch %>%
     dplyr::group_by(ModSim, year, Code, Name, fishery, variable, units) %>%
     dplyr::summarize(value = sum(value, na.rm = TRUE)) %>%
     dplyr::mutate(value = ifelse(variable == "cv", value/stepperyr, value)) %>%

### 2023-06-13 R/simCatchIndexSubannualFW.r
adds simulated catch for food web modeling

@@ -1,0 +1,25 @@
+#' The Atlantis-simulated fishery catch indices for the mskeyrun project
+#'
+#' atlantosom output is accessed and fishery data pulled over time
+#' simulated fishery catches are cumulative, not snapshots
+#' fishery total catch for the year is the sum of the months
+#' fishery catch for the month is all catch reported since the last reporting month
+#' this dataset includes food web and multispecies model simulated species
+#'
+#' @format A data frame
+#'
+#' \describe{
+#'\item{ModSim}{Atlantis model name and simulation id}
+#'\item{year}{year simulated fishery conducted}
+#'\item{fishMonth}{month simulated fishery conducted}
+#'\item{Code}{Atlantis model three letter code for functional group}
+#'\item{Name}{Atlantis model common name for functional group}
+#'\item{fishery}{simulated fishery name}
+#'\item{variable}{catch or coefficient of variation (cv) of biomass}
+#'\item{value}{value of the variable}
+#'\item{units}{units of the variable}
+#'
+#' }
+#'
+#'
+"simCatchIndexSubannualFW"

### 2023-06-13 data-raw/R/create_sim_fishery_index_fw.R

@@ -1,0 +1,146 @@
+#' Read in fishery catch data save as rda
+#' 
+#' atlantosom output is accessed and fishery data pulled over time
+#' simulated fishery catches are cumulative, not snapshots
+#' fishery total catch for the year is the sum of the months
+#' fishery catch for the month is all catch reported since the last reporting month
+#' 
+#' also makes annual aggregate dataset for backwards compatibility
+#' 
+#'@param atlmod configuration file specifying Atlantis simulation model filenames 
+#'and locations  
+#'@param saveToData Boolean. Export to data folder (Default = T)
+#'
+#'@return A tibble (Also written to \code{data} folder)
+#'\item{ModSim}{Atlantis model name and simulation id}
+#'\item{year}{year simulated fishery conducted}
+#'\item{fishMonth}{month simulated fishery conducted}
+#'\item{Code}{Atlantis model three letter code for functional group}
+#'\item{Name}{Atlantis model common name for functional group}
+#'\item{fishery}{simulated fishery name}
+#'\item{variable}{catch or coefficient of variation (cv) of biomass}
+#'\item{value}{value of the variable}
+#'\item{units}{units of the variable}
+#'
+
+library(magrittr)
+
+create_sim_fishery_index_fw <- function(atlmod,fitstart=NULL,fitend=NULL,saveToData=T) {
+
+  # input is path to model config file for atlantisom
+  source(atlmod)
+  
+  # path for survey and fishery config files
+  cfgpath <- stringr::str_extract(atlmod, ".*config")
+  
+  #works because atlantis directory named for model and simulation
+  modpath <- stringr::str_split(d.name, "/", simplify = TRUE)
+  modsim <- modpath[length(modpath)]
+  
+  #read in survey biomass data
+  catchbio_ss <- atlantisom::read_savedfisheries(d.name, 'Catch')
+  
+  # get config files for fishery cv
+  fishcon <- list.files(path=cfgpath, pattern = "*fishery*", full.names = TRUE)
+  
+  # read true list with run and biol pars, etc
+  omlist_ss <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_fw.rds")))
+  
[…]

### 2023-06-13 data-raw/build_simdata.R

@@ -57,3 +57,9 @@ create_sim_percapconsumption(atlmod, fitstart=40, fitend=120)
 
 create_sim_startpars(atlmod, fitstart=40, fitend=120)
 
+# food web model specific datasets add other species
+
+create_sim_survey_index_fw(atlmod, fitstart=40, fitend=120)
+
+create_sim_fishery_index_fw(atlmod, fitstart=40, fitend=120)
+

### 2023-06-13 data-raw/R/create_sim_survey_index_fw.R
start food web survey, missing biomass pools

@@ -1,0 +1,107 @@
+#' Read in survey data save as rda
+#' 
+#' atlantosom output is accessed and surveys pulled over time
+#' 
+#'@param atlmod configuration file specifying Atlantis simulation model filenames 
+#'and locations  
+#'@param saveToData Boolean. Export to data folder (Default = T)
+#'
+#'@return A tibble (Also written to \code{data} folder)
+#'\item{ModSim}{Atlantis model name and simulation id}
+#'\item{year}{year simulated survey conducted}
+#'\item{Code}{Atlantis model three letter code for functional group}
+#'\item{Name}{Atlantis model common name for functional group}
+#'\item{survey}{simulated survey name}
+#'\item{variable}{biomass or coefficient of variation (cv) of biomass}
+#'\item{value}{value of the variable}
+#'\item{units}{units of the variable}
+#'
+
+library(magrittr)
+
+create_sim_survey_index_fw <- function(atlmod,fitstart=NULL,fitend=NULL,saveToData=T) {
+
+  # input is path to model config file for atlantisom
+  source(atlmod)
+  
+  # path for survey and fishery config files
+  cfgpath <- stringr::str_extract(atlmod, ".*config")
+  
+  #works because atlantis directory named for model and simulation
+  modpath <- stringr::str_split(d.name, "/", simplify = TRUE)
+  modsim <- modpath[length(modpath)]
+  
+  #read in survey biomass data
+  survObsBiom <- atlantisom::read_savedsurvs(d.name, 'survB') #reads in all surveys
+  
+  # get config files for survey cv
+  svcon <- list.files(path=cfgpath, pattern = "*survey*", full.names = TRUE)
+  
+  # read true list with run and biol pars, etc--specific to food web
+  omlist_ss <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_fw.rds")))
+  
+  # model timesteps, etc from omdimensions script
+  source(paste0(cfgpath,"/omdimensions.R"), local = TRUE)
+  
+  #Number of years
+  nyears <- omlist_ss$runpar$nyears
+  total_sample <- omlist_ss$runpar$tstop/omlist_ss$runpar$outputstep
+  
+  # user specified fit start and times if different from full run
[…]

### 2023-08-25 data-raw/R/create_real_survey_lennumcomp.R
reworked survey lengths. replaced survdat with similar brute force method

@@ -7,9 +7,11 @@
 #' 
 
 library(magrittr)
+library(sf)
 
-
-create_real_survey_lennumcomp <- function(convertKGtoMT=F, overwrite=F) {
+create_real_survey_lennumcomp <- function(convertKGtoMT=F,
+                                          displayTable=F,
+                                          overwrite=F) {
   
   if (convertKGtoMT) {
     scalar <- 1000
@@ -44,60 +46,149 @@ create_real_survey_lennumcomp <- function(convertKGtoMT=F, overwrite=F) {
   # get survey codes
   svspps <- species$SVSPP
   
+  # IS THIS CORRECT. WHY DID I USE WGTLEN AS ABUNDANCE
+  # newData <- survdatLWpull$survdat %>% 
+  #   dplyr::filter(SVSPP %in% svspps) %>% 
+  #   dplyr::mutate(SVSPPLEN = paste0(SVSPP,"-",LENGTH)) %>% 
+  #   dplyr::select(-BIOMASS,-ABUNDANCE) %>%
+  #   dplyr::mutate(BIOMASS = PREDWT,
+  #                 ABUNDANCE = WGTLEN)
+  # 
+  
+  
+  # I THINK THIS IS BETTER.
+  # filter all GB Strata
   newData <- survdatLWpull$survdat %>% 
-    dplyr::filter(SVSPP %in% svspps) %>% 
-    dplyr::mutate(SVSPPLEN = paste0(SVSPP,"-",LENGTH)) %>% 
-    dplyr::select(-BIOMASS,-ABUNDANCE) %>%
-    dplyr::mutate(BIOMASS = PREDWT,
-                  ABUNDANCE = WGTLEN) 
+    dplyr::filter(STRATUM %in% GBStrata) %>% 
+    dplyr::mutate(SVSPPLEN = paste0(SVSPP,"-",LENGTH))
   
   
-  # fall: calculate the swept area biomass for each species length combination 
-  sweptAreaFall <- survdat::calc_swept_area(newData,
-                                            filterBySeason = "FALL",
-                                            filterByArea = GBStrata,
-                                            groupDescription = "SVSPPLEN") %>%
-    dplyr::select(YEAR,SVSPPLEN,tot.biomass,tot.abundance) %>%
-    tidyr::separate(.,SVSPPLEN,c("SVSPP","LENGTH"),sep = "-") %>% 
-    dplyr::mutate(SEASON = "FALL")
+  realSurveyLennumcomp <- NULL
[…]

### 2023-10-07 R/simSurveyDietcompFW.r
add food web diets and documentation

@@ -1,0 +1,23 @@
+#' The Atlantis-simulated survey diet composition for the mskeyrun project
+#'
+#' Simulated survey diet composition data from Atlantis model as sampled by atlantisom: 
+#' Full food web beyond focal species for Rpath
+#'
+#' @format A data frame
+#'
+#' \describe{
+#'\item{ModSim}{Atlantis model name and simulation id}
+#'\item{year}{year simulated survey conducted}
+#'\item{Code}{Atlantis model three letter code for predator group}
+#'\item{Name}{Atlantis model common name for predator group}
+#'\item{survey}{simulated survey name}
+#'\item{agecl}{age class of Atlantis functional group}
+#'\item{prey}{Atlantis model common name for prey group}
+#'\item{variable}{proportion in diet (dietprop)}
+#'\item{value}{value of the variable}
+#'\item{units}{units of the variable}
+#'
+#' }
+#'
+#'
+"simSurveyDietcompFW"

### 2023-10-07 R/simSurveyIndexFW.r

@@ -1,7 +1,7 @@
 #' The Atlantis-simulated survey biomass indices for the mskeyrun project
 #'
 #' Simulated survey index data from Atlantis model as sampled by atlantisom
-#' Includes all structured groups as separate surveys
+#' Includes all structured groups as separate surveys for Rpath
 #'
 #' @format A data frame
 #'

### 2023-10-07 data-raw/R/create_sim_survey_dietcomp_fw.R

@@ -1,0 +1,108 @@
+#' Read in survey data save as rda
+#' 
+#' atlantosom output is accessed and surveys pulled over time
+#' 
+#'@param atlmod configuration file specifying Atlantis simulation model filenames 
+#'and locations  
+#'@param saveToData Boolean. Export to data folder (Default = T)
+#'
+#'@return A tibble (Also written to \code{data} folder)
+#'\item{ModSim}{Atlantis model name and simulation id}
+#'\item{year}{year simulated survey conducted}
+#'\item{Code}{Atlantis model three letter code for predator group}
+#'\item{Name}{Atlantis model common name for predator group}
+#'\item{survey}{simulated survey name}
+#'\item{agecl}{age class of Atlantis functional group}
+#'\item{prey}{Atlantis model common name for prey group}
+#'\item{variable}{proportion in diet (dietprop)}
+#'\item{value}{value of the variable}
+#'\item{units}{units of the variable}
+#'
+
+library(magrittr)
+
+create_sim_survey_dietcomp_fw <- function(atlmod,fitstart=NULL,fitend=NULL,saveToData=T) {
+
+  # input is path to model config file for atlantisom
+  source(atlmod)
+  
+  # path for survey and fishery config files
+  cfgpath <- stringr::str_extract(atlmod, ".*config")
+  
+  #works because atlantis directory named for model and simulation
+  modpath <- stringr::str_split(d.name, "/", simplify = TRUE)
+  modsim <- modpath[length(modpath)]
+  
+  #read in survey annual age comp data
+  all_diets <- atlantisom::read_savedsurvs(d.name, 'survDiet')
+  
+  # get config files -- needed?
+  svcon <- list.files(path=cfgpath, pattern = "*survey*", full.names = TRUE)
+  
+  # read true list with run and biol pars, etc
+  omlist_ss <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_fw.rds")))
+  
+  # model timesteps, etc from omdimensions script
+  source(paste0(cfgpath,"/omdimensions.R"), local = TRUE)
+  
+  #Number of years
+  nyears <- omlist_ss$runpar$nyears
+  total_sample <- omlist_ss$runpar$tstop/omlist_ss$runpar$outputstep
+  
[…]

### 2023-10-07 data-raw/build_simdata.R

@@ -63,6 +63,8 @@ create_sim_survey_index_fw(atlmod, fitstart=40, fitend=120)
 
 create_sim_fishery_index_fw(atlmod, fitstart=40, fitend=120)
 
+create_sim_survey_dietcomp_fw(atlmod, fitstart=40, fitend=120)
+
 # below combines already loaded mskeyrun datasets,  
 # outputs of create_sim_survey_agelen and create_sim_survey_dietcomp
 # ensure that these are up to date before running
